% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
% DOCUMENT CLASS
% =============================================
\documentclass[twoside,10pt]{book}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\author{}
\date{}

\usepackage{fontspec}
  \setmainfont[Ligatures=TeX]{Charter}
  \setsansfont{Lucida Grande}
  \setmonofont{American Typewriter}
  \newfontfamily\chapfont{Charter Black}
  \newfontfamily\secfont{Charter Black}
  \newfontfamily\handfont{Bradley Hand}

\setlength{\paperwidth}{6.125in}
\setlength{\paperheight}{9.25in}
\pdfpagewidth=\paperwidth
\pdfpageheight=\paperheight

\setlength{\textwidth}{4in}
\setlength{\textheight}{7in}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0.125in}
\setlength{\topmargin}{-0.25in}


\renewcommand{\baselinestretch}{1.1}
\renewcommand{\arraystretch}{1.2}
\setlength{\tabcolsep}{2pt}

\raggedbottom

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{} % clear all header and footer fields
\renewcommand{\headrulewidth}{0pt} % remove the header rule
\fancyfoot[C]{\small\thepage} % small page number in the center of the footer
\setlength{\footskip}{36pt}

\usepackage{tikz}
\usepackage{xcolor}
\definecolor{avocadogreen}{rgb}{0.34, 0.51, 0.01} % An example of avocado green
\definecolor{vintageorange}{rgb}{0.95, 0.65, 0.0} % An example of vintage orange
\definecolor{darkavocadogreen}{rgb}{0.20, 0.30, 0.01} % A darker shade of avocado green
\definecolor{darkvintageorange}{rgb}{0.70, 0.40, 0.0} % A darker shade of vintage orange




% TITLE
% ======================================================================
\begin{document}
\pagestyle{empty}
\cleardoublepage
\mbox{ }
\cleardoublepage
\frontmatter
\begin{center}
  \vspace*{48pt}
  \hfill {\fontsize{86}{52}\sffamily\bfseries poems}
  \\[12pt]
  \hrule height 1pt
  \hfill {\fontsize{24}{52}\sffamily carol carpenter}
  \vfill
  \hfill {\sffamily colloquial publishing}
  \\
  \hfill {\sffamily new york}
\end{center}
\color{black} % Sets the text color to white

\clearpage
\mbox{}
\\[36pt]
Carol Carpenter (1943--2016).~\textit{Poems}.~Colloquial Books.~New
York. 
\\[12pt]
\begin{center}
Copyright \copyright\, Carol Carpenter
\\[4pt]
All rights reserved.
\end{center}
\vfill
\begin{quote}
\emph{This is a work of fiction. Any resemblance to actual events or
persons, living or dead, is entirely coincidental.}
\end{quote}
\vfill
\vfill
\null

\clearpage
\pagestyle{fancy}
\noindent
{\Large\bf Previous Publications}

``Big Pink'' appeared in \emph{Snake Nation Review,} Issue 19\\
(Winter, 2005)

``Blonde Showgirls'' appeared in \emph{The Mochila Review}, Volume 11
(2009)

``The Catcher'' appeared in \emph{Living with Adolescent Literature: The
Connecticut English Journal''}

``Chicken Bones'' appeared in \emph{Real}, Volume 34.2 (2010)

``Chroma: Gloria Colors Her World'' appeared in \emph{Underground
Window}, Volume 2, Issue 1 (November, 2005)

``Cross Currents'' appeared in \emph{The Anthology of New England
Writers}, 2006.

``Cupcake Cathedral'' appeared in \emph{The Rockford Review}, Winter
2005.

``Daylight Savings Time'' appeared in \emph{LitRag} Eighteen (2004)

``Death of the Bird Child'' appeared in the 3\textsuperscript{rd} Annual
Creekfeet\\
Poetry Contest, 2008.

``The Diet'' appeared in Pteranodon, Vol. 1, No. 2.

``Distances'' appeared in \emph{Nit \& Wit.}

\textbf{\hfill\break
}



\cleardoublepage
\tableofcontents


\mainmatter
\cleardoublepage

\section{Aborigine}

We rode this island bareback\\
curled desert sand under toes,\\
dug for honey ants. We looked\\
back at Uluru whose red rock\\
spirits sang our home to life,\\
hummed us toward valleys,\\
banks of mud. We rolled\\
our bodies in brown soil\\
wet with water, snug\\
as second skin dried in heat.\\
We masked our human smell,\\
tracked downwind the sounds\\
of night where ancestors dwelled.

We peeled back from trees\\
in Arnhem Land, painted\\
charcoal stories of our clan\\
who sprung from red clay\\
along with our lizards, turtles,\\
sun who all renewed our lives\\
during daily rituals

when we streaked our bodies ochre,\\
chanted our history, our circle\\
of feet moving fast, our galloping\\
bones shaking a tune that lured\\
women to men, moon to earth\\
where we became one and no one\\
could own this land, this black sky\\
pricked with stars during Dreamtime\\
when spirits created light for us\\
who came first.

\clearpage
\section{Above the Liquor Store}

Social Services should be notified\\
about those two youngsters, babies\\
really, not more\\
than three or four\\
years to their names\\
which the woman in the white Escort\\
didn't know

when she saw these two round faces,\\
blurred by yellow moon, flickering on and off\\
in code above the liquor store. Alone\\
up there in that Detroit apartment,\\
those babes teetered on the edge\\
of the window ledge as if they drank\\
vodka straight during rush our exhaust

and bathed in perfumed city dusk\\
muffled, thick as lust, a touch\\
of dust that woman wiped\\
from her dashboard, from her lips,\\
dry and chapped. She watched\\
those toddlers rake the air\\
with fingers spread\\
down toward motor heat,

down toward her. She could,\\
she knew, report those children\\
or report herself as the woman below\\
who braked on red, and waited\\
for green, her signal to go.

\clearpage
\section{An Agoraphobic Explains Fear}

On television, you can watch,\\
as I have more than once,\\
some stranger's brain\\
cupped in a surgeon's hands\\
as she expounds on normal functions,\\
how language lurks on the right side,\\
while singing emerges from the left,\\
unheard now. Death hushes all.\\
No impulses jump, no neurons\\
electrify messages or melodies.

Just me on this floral couch\\
with my brain inside my skull.\\
I can not reach past bone, can not\\
unwrap ropes of my gray prison.

My brain misfires,\\
my therapist says. Think\\
of fireworks, an explosion\\
of sorts. You want\\
a safe place. Lock\\
your brain, chain your door.\\
Nothing is enough, though.\\
The world will intrude.

I see, I answer.\\
Outside my window,\\
I can tell you how\\
broken bits of leaf\\
stick to rusted screen.

Ants trek along\\
my windowsill, a line\\
of broken bodies\\
wrestle with bread\\
crumbs, food for their queen\\
who lives in sand.

Past this frame, seasons change.\\
In autumn, wind rips leaves\\
from red maples in my side lot\\
and drops them helter-skelter\\
around my neighborhood. Covered\\
in winter by snow. If only\\
you could see how squirrels skid\\
across frozen surfaces, you would not\\
ask why I fear cars on icy roads.

When crocus surface in spring, I\\
recall the locust tree with no buds\\
and the willow tree whose roots\\
snuck under my basement floor,\\
cracking concrete with stringy brown\\
roots claiming my house. I imagine\\
how roots can climb stairs, strangle\\
me as I sleep in my own bed, unaware.

Let me tell you about\\
last summer, when I walked\\
these familiar streets\\
needing some eggs for baking.\\
Yes, it was test.\\
I took my cane so I felt safe.\\
But the neighborhood had changed.

The grocery store was gone.\\
Houses had boards nailed\\
across windows. I did not know\\
those hooligans who sashayed out\\
of the old Baker house,\\
chains dangling from belts,\\
rings in their ears,\\
chests bare except for tattoos,\\
horrid ones of red-eyed dragons\\
breathing fire. Those boys\\
smoked and smelled of drink\\
when they got up close. I froze\\
in all that heat. They had knives,\\
all silver flash at noon blinded me.\\
They grabbed my purse, laughed, shouted\\
awful words. I keeled over and saw\\
those fireworks you mentioned.

You do not understand.\\
I have traveled in my days\\
seen English gardens, Ohio cornfields,\\
oceans full of salt and sharks. Even rode\\
an elephant once, in India, I think.\\
Every blade of grass, breath of air, pink\\
petal, courthouse brick, city street, every\\
one swims in my heart. We play together\\
in waves behind my steel door.

You speak of brains and how they work.\\
I speak of other body parts, even skin.\\
There is no way I can let you in.

\clearpage
\section{Anatomy of an Argument}

I.  The Head Violates Rules Logically

All day I fumed about what you had not done,\\
my love. You did not\\
cut the lawn, did not turn off the television when I asked.\\
You were not at work when I called you there.\\
My anger climbed like the red blaze roses on our trellis.\\
Each petal shriveled to brown at the edges, brittle as my heart.

What's that you said? You will do it now if I remain calm? I am not\\
calm. Do not use logic on me. Haven't you caused enough damage?\\
I threw my antique Fiesta plate\\
at you, the yellow one bright as the banana you dared eat as if
nothing\\
were wrong. That plate cracked down the middle and shattered. I knew\\
where you were when I phoned. You were not able to answer\\
anyone except the other her who has no plates and don't you dare\\
walk away. I am in a magenta rage.

II.  The Heart Screams for Blood

Ah, you are not here, my love.\\
You deserted me in autumn\\
after the oak leaves flipped from green to yellow-\\
orange as if they were on fire, about to explode\\
with my hatred of you. I loathe the day\\
you drove away with your books\\
packed in boxes. You grabbed the \emph{Da Vinci Code}\\
I gave you as a gift before I ripped out\\
all the pages you had not finished reading.

You write to me as if love can be\\
reignited, as if passion does not\\
fizzle like spent firecrackers. We end this chapter with a bang\\
in a tail of smoke, in our free fall to hard ground\\
where spectators wait for disaster to fall in their laps,\\
something big like heart blood and voodoo hate,\\
a spell of death I call forth with black, black words\\
chalked on all our memories, an X that marks our past love.

III.  The Soul Searches for Some Way Out

My love, you are no more\\
than the lies you told me about her\\
about you, about the way things were with us\\
before you left. You lied or should I say\\
you lost your path. Your fantasies moved too fast\\
and you had no bread crumbs to drop. She was your mirage,\\
a curved shape in the distance standing on white sand\\
beside blue water while palm fronds swayed\\
under a round yellow sun. Your mouth still twists\\
with dishonesty and my soul trails ribbons of mourning.

I acknowledge the truth, how I bite through\\
the thick skin of the bitter pomegranate that I eat.\\
Red juice dribbles from my lips that once tasted of you.\\
That's the truth the best I know it.


\clearpage
\section{An Invitation to Answer}

Just a question. Do you mind?\\
Not the big one like do you love me,\\
yes, no, maybe\\
today and not tomorrow\\
or, perhaps, even yesterday\\
when intimacy starts and ends

in niggling questions\\
that tug on your shirt buttons\\
prance through your bloodstream\\
squat in corners of your eye\\
while the tip of your tongue\\
probes your eyetooth

betrays doubts\\
you cradle all day long\\
afraid to drop one splat\\
on the floor, a ripe tomato.\\
Black-seed answers\\
skitter everywhere.

Just one question\\
crisp as a peapod?\\
You slit that green curve\\
and three round pellets\\
wobble out:

Have you tasted\\
those three green globes\\
plump with your betrayal?

\clearpage
\section{Anonymous}

We start and end in blackness\\
like a period that ends the sentence\\
of life or death -- and all the in-between.

That man there is one of us. He swims\\
in darkness. For nine months he is\\
a blind fish in the Mammoth Cave.

No name for a day. Just baby\\
at Hutzel Hospital. One scream\\
among many open mouths. No teeth.

The glass streets of Detroit\\
shave his cheeks, break his nose,\\
light the night with gang fights.

At the Rouge Steel Plant, he sweats\\
pieces of himself. Salty brine\\
puddles in his left eye, an ocean of sharks.

Dirt collects in his hair. Each strand\\
a tributary, a stream of muddy water\\
eroding all he is, bit by bit.

His hair recedes. His eyelids droop\\
beneath water. He floats under black cross,\\
like us. He returns to \emph{x}, the unknown.

\clearpage
\section{The Art Teacher\\ and Her Granny Smith Apple}

Yesterday, the art teacher asked her third grade students to bring an
apple to class. Today, one by one, the children arrange their apples on
her desk in perfect disarray: Side-by-side, stacked in toppling apple
towers, right-side-up or -down, brown stems poking out every which way.
Twenty Red Delicious apples, no bruises anywhere. Just glossy red,
taught skins and the smell of cider poems in the air.

Inhale apple poems through your eyes, the art teacher says. Let the
taste dawdle on your tongues, rush through your veins and pulse in your
fingertips. Now, close your eyes. Touch apples behind your eyelids. Move
your paintbrushes against paper. Gently. Ever so gently, smudge the
edges.

In the beginning, the art teacher tells the children, there was one
apple that was plucked from the tree. Now, we have our pick from the
grocer's shelf. So many apples. It's hard to choose. Twenty apples here:
all tempting Red Delicious.

With a flourish, the art teacher pulls her white linen handkerchief
edged in red lace from the sleeve of her dress. She gathers up the third
graders' nose -- all their buzzing words, all the screeching of their
crepe soles against oak floors, all the swishing and whirlpools they
create when they dip their paint-soaked brushes into Mason jars filled
one-third full with water run from the faucet. She muffles that noise in
her handkerchief and draws it, again and again, against the palm of her
cupped hand.

Silence now, just as it was in the beginning.

Twenty sets of eyes full of questions watch the art teacher's hands
shake out the handkerchief, spread it across her desk. In the middle of
that white, she places one Red Delicious apple. Like your paint poems,
she sighs, this apple absorbs the light. She lifts that apple to her
lips and takes a bite.

One by one, twenty children choose an apple and eat. Red Delicious drips
between their fingers an onto their paintings as they chew to the cores
and catch seeds between their baby teeth.

In the palm of her hand, the art teacher holds something covered with
her handkerchief, whiter than clouds. She whisks off the covering.
Nestling there, a Granny Smith apple, green as the leaves on any apple
tree. She cuts twenty small slices and drops them into twenty small
mouths that pucker at the tart taste.

Remember this, the art teacher says. Whenever you paint a Red Delicious,
use soft-bristled brushes or your fingertips. For Granny Smith apples,
apply layer upon layer of green and yellow oil paints and shape your
life poems: too opaque at the end for light to penetrate and textured
enough to make it worth the taste everafter.

\clearpage
\section{Archeological Dig}

In my father's garage, I uncover artifacts of his\\
past lives -- my third-grade math tests boxed with\\
his wife's valentines, Merrill Lynch statements\\
filed back to the 1950s, his mother's note when he\\
escaped the Ohio farm and hitchhiked to Detroit,\\
an empty box of Miracle Grow.

What I do not find speaks loudest.

\clearpage
\section{Ask for the Moon}

Why settle for moondust\\
when you want the whole rock?

Lasso that silver love disc.\\
Loop the rope around your wrist.

Let the silver balloon bounce behind,\\
reflect luscious lips open tonight.

Lovers desire moonlight,\\
the fury of fire once removed.

\clearpage
\section{Aunt Amaryllis in the Dead of Winter}

Before Detroit can shake its winter skirt\\
free of tire tracks, Aunt Amaryllis\\
jitterbugs under January's quarter moon\\
while juggling her loose change

usually kept safe\\
in her back pocket. Her bulbous rump\\
girded in brown cotton and crinolines\\
for over sixty years -- all that rich-veined flesh dormant\\
through spring, summer and fall\\
sprouts a fleck of green\\
inside the boxed night

where two husbands left her\\
alone and barren,\\
wondering what to be\\
or not be, just her

tiny voice squeaks\\
as she pushes cash register keys\\
at Farmer Jack's for forty years\\
and bags bananas, soup cans,\\
skimmed milk for mothers cutting back\\
calories every day in denial

these mothers crave new bodies\\
not wanting an ounce more\\
while she hangs pounds of fat\\
on her bones -- over two hundred\\
bones converge\\
into her

into this fleshy, green stalk\\
of a woman, into Aunt Amaryllis\\
who unravels her layers of brown\\
and blooms four crimson heads:

One blushes the blood of birth.\\
One gulps the scarlet moon rising.\\
One ignites her curls of fire\\
while the other one winks

one red eye for Aunt Amaryllis\\
who struts her green-booted self\\
into this night of reckoning.

\clearpage
\section{Aunt Glory's Estate Sale}

Just spread out your life\\
why don't you.\\
Leftover dibs and dabs\\
of forty years\\
laced into Grandma's colonial\\
brick: two stories

or so I've heard. One\\
full of knots you tied\\
when your lover left\\
unnamed. One\\
a gothic tale of midnight\\
and morning glories forever\\
closed, tight-mouthed\\
prim as you

who want more than a dollar\\
for that bone china\\
cup of your life.

\clearpage
\section{Bag Full of Fall}

\begin{quote}
\emph{Fill this bag with pretty, colorful\\
leaves. Bring it in Wednesday.\\
\mbox{}\qquad\qquad\qquad\qquad -- Miss Amy.}
\end{quote}

The day the leaves dropped\\
higgledy-piggledy at the park, my grandson\\
squeezed two maples and one birch\\
in his bare hand, crumpled them to bits\\
he threw in the air like confetti,\\
pigments of marigold orange, red,\\
dusty brown he breathed in when\\
he heaped leaves over his head,\\
over his green-leaf body.\\
He licked dry oaks from his lips.

My grandson crammed the bag\\
with crinkled leaves, acorns,\\
two beach stones, a Canada goose\\
feather and two pretzel rods,\\
salted away for one more fine fall day.

\clearpage
\section{Balancing on the Barstool\\ at Tommy's Tavern}

She sat tilted to one side, her fingers wrapped\\
around the beer mug, touching the glass center.\\
Ice melted at the edges,\\
puddled in the palm of her hand\\
as if the glass itself had returned to sand.\\
Her fingers traced the path, down\\
onto the napkin, soaked and shredded\\
by the cherry red nails of her other hand.\\
She refused to drink beer, warm now,\\
and called for another, colder one.

\clearpage
\section{The Balloon Man}

His eyes bounce\\
around the crowd,\\
float like two balloons,\\
blue as the sky,\\
light as hope.

One more huff of air.\\
The balloon man flicks\\
his wrist. He locks a leg\\
in place, another leg,\\
two more. A neck, a face\\
and one slender body.

A green poodle\\
prances into the arms\\
of a pink-ruffled girl\\
Two long puffs\\
plump the giraffe's yellow neck.\\
The balloon man breathes\\
blue tulips into bloom.\\
They dangle from his fingertips.

Faster.\\
The balloon man blows,\\
pulls balloons tight at their\\
mouths,\\
lets out squeaks of air,\\
twists paws and muscled legs,\\
curves backs and turtle shells.

Hands reach out for colors:\\
gold, green, purple, vermillion.\\
Bluebirds and parrots feather his hands.\\
Deer leap from his lips.\\
Elephants trumpet silver surprise.\\
Red ladybugs speckle the air.\\
Daisies spring from children's toes.
\clearpage
The balloon man burps each balloon,\\
leaves space for breath and unicorns.

\clearpage
\section{Baptized in Suite 225}

Every day, our 22 female voices\\
hoarse with conviction\\
galloped through phone wires\\
sold insurance: home, health

life full of catastrophic\\
exclusions, commissions we counted\\
with gusto, shrill vowels rose\\
a torrent of words wrapping

our sound around fluorescent tubes\\
too hot for fantasies while we were\\
crammed behind this door, stuffed\\
in tweed swivel chairs where

we were worn thin as upholstery\\
rotted by dry air, recycled each one\\
knew this job would last long enough\\
unless we wanted more like the woman

fired yesterday for ushering in a creek\\
its sac of water breaking just in time\\
to birth 21 other women whose lusty screams\\
melted ice, fed buds on leafless trees

ready now to sprout foliage, shade us\\
when we dip bare toes in currents\\
as we follow wetness wherever it may lead\\
us without flood insurance, without question.

\clearpage
\section{Becoming Unraveled}

Loose strings marked my trail\\
as a child so I could find\\
my way home. Blue threads of me,\\
dangled down from the apple tree\\
on the corner of our street\\
where I climbed to sky,\\
plucked Red Delicious globes,\\
shined them on frayed jeans\\
before this big bite I savored,\\
letting juices dribble down,\\
scent my clothes as if I were\\
an apple burnished in the sun.

As my body curved into woman,\\
I gathered up my threads,\\
rolled each piece into one big ball.\\
Out of this tangled skein,\\
I wove myself into a Detroit scene,\\
my bones a loom. I wrapped\\
the waft around the warp,\\
and created the tapestry of me.

My boss did not think\\
I could weld bolts on those chassis\\
at the GM Assembly Plant.\\
Faster than men, my bolts held\\
until I got too hot from all that fire\\
spurting from my welding gun. Even though\\
the money was good, I burned out.

My weft was too loose\\
when I welded bolts and screamed\\
at voices in my head. Those voices\\
were like liquid steel. I bubbled up\\
and no one could melt me.\\
Such heat was unbearable.

I was becoming unraveled.\\
All that held me in place\\
were loose threads, a loom,\\
strings of my childhood.

\clearpage
\section{Bedtime Fables}

You speak to me of wolf stories,\\
pull me close against your waist\\
like some benign tumor growing there,\\
hard, solid tissue, a fist against your skin.

The wolf sucks up straw houses, you say,\\
spits out grain baked into biscuits\\
I carry hidden in my basket.

Your finger spears and slices words,\\
serves them up on paper plates,\\
a picnic in the woods with wolves\\
who prowl in daylight and howl\\
all night. You, father,\\
warn me of the cutting time.

When father works\\
the night shift at the factory,\\
you, mother, tell tales\\
of mythical beasts who rise\\
above my bead like bread.

Your hands,\\
kneading knuckles,\\
shape unicorns with horns\\
of ivory bone\\
and hoofs of ebony.\\
You carve spaces I can mount.\\
My head bends low into the wind.

We ride, we two, you behind me.\\
Our legs straddle both sides,\\
giving birth to the morning dove\\
who circles the ceiling\\
and nests under my pillow,\\
fluffing feathers\\
soft as down on my arms.

Your voices melt, merge\\
at the bottom of the page, thick\\
as raspberry brandy, fermented\\
fruits piling, peeling my brain.\\
I toast with crystal flutes, head back.

I am tipsy with wolves,\\
fur brown as beer, full of malt\\
and hops. Lips sheathe teeth, wet\\
with blood, mere camouflage.\\
Out of silk pouches, I shake\\
unicorns. Free of bridles, of reins,\\
we canter past churches, cemeteries.\\
The mourning dove's message\\
covers my heart like a sheet.

\clearpage
\section{Betrayal}

Pin up Bertrum that twerp\\
that snail whose slimy track\\
wound around your neck,\\
his lies a velvet choker.

Your turn. Now breathe out\\
profanities, a whole string\\
and spin the fiber, tough\\
enough. Leave him hanging

in the name of love and all things\\
clean as your bleached cotton\\
towels strung on the line, rows of hands\\
waving down truth, the bits and pieces

once scattered, now unite and whip\\
the breezes into a gust, a whirlwind\\
of dust and pollen grains, a funnel\\
full of yellow powder. You embrace

the fertility of the goldenrod. You\\
bury your face in towels of truth,\\
fold each oblong scrap of white\\
and wring out Bertrum wet with lies.

\clearpage
\section{Between Me and Thee}

My camera clicks\\
quick again. Again,\\
in rhythm like birdsong,\\
I frame, aim, shoot.\\
These mountains blur.\\
White clouds shroud dark\\
peaks sharp enough to pierce\\
white sky with gray.

Snow streams carve crevices,\\
a parable of black and white\\
where any man can sink. I\\
capture still shots,\\
freeze the lake, the blue\\
wet smear feeding green\\
pine. A dense forest\\
speaks of elk, deer, men lost.

I need a tripod, a time\\
exposure, a steady hand\\
in wilderness. If I survive\\
this journey, I may catch your image.\\
Just between me and thee, I confess\\
I am an amateur. I cannot find\\
the right setting for destruction,\\
for stripped logs floating to market.

\clearpage
\section{Beyond the Horizon}

I am the woman\\
who touched cumulous clouds,\\
shook them in my teeth as if\\
they were laundered sheets,\\
pillowcases where you and I\\
sunk our heads in down\\
and slept our dreamless sleep.

I am the woman\\
who found a toehold,\\
shimmied belly to stone,\\
from one red rock ledge to the next\\
where I sucked water from stone.\\
Alone I climbed, your shadow buckled\\
in my backpack. At peak, I set you free.

I am the woman\\
who strolled lake bottom,\\
sifting sand between my toes,\\
each hard bit held against skin\\
and released. In this postcard I send,\\
you can see me, my blue ruffled skirt, my head.\\
My arms spread open to what lies beyond you.

\clearpage
\section{The Big Brass}

I have never seen\\
the big brass, the faceless one,\\
all angular arms. His pinchers\\
grip our round: a few countries\\
or maybe even blue oceans\\
wet enough to drop. No safety net.

Centuries of wars, famine, plagues\\
tarnish our ball, pile weight on one side,\\
then the other. We all scurry,\\
here to there. We slide past\\
his brass forearm, past his wrist bone,\\
right into his palm. We dangle there.

The big brass can not\\
be bribed with gold. He will protest\\
declarations, explanations, proclamations,\\
tallies of wrong that may be right. No,\\
he commands balance, up, down, across\\
and around. Resist and go to jail.

The strategy of brass is white button.\\
A fingertip temptation.

\clearpage
\section{Big Pink}

All girlie stuff whispers pink\\
as a flamingo's neck turned\\
right. Go ahead and stare us down.\\
Play the checker game. King\\
us with red glory. We prefer\\
warrior queens, the scarlet hue.

We can rope you in, you men\\
who do not gaze past big pink.\\
See, we are\\
all sweet flowers, a garden of round\\
suns with petals curved. No danger\\
there. Just wrap your shoulders\\
with our quilt, warm yourselves to hot.

Arms bare, we transport our arsenal.\\
No help needed. Together, we women\\
untangle strands of life. Our hair\\
is curled and brushed. Our shining locks\\
radiate, beckon you. Your man face\\
winks. You lick your own pink lips. Beware\\
the hair spray. That fine mist holds in place\\
our shining glory. Our blood-red nails claw.

\clearpage
\section{Bits of my Dead Self\ldots{}}

\ldots burrow into grey-haired angora\\
bunny fluffs under my king-\\
sized bed where the pink-nosed\\
congregation mates all\\
wiggle-eared and warm-blooded.

\ldots waltz in triple time\\
with pollen flecks as partners\\
in the step, slide, step, the whirl\\
and dip. With quick silver stitches\\
I hem the sunbeam's yellow skirt.

\ldots swirl off my skin in one day\\
as I shed the outer layer\\
dust of who I was.\\
Now my iridescent flesh\\
revels in the possibilities.

\clearpage
\section{A Black and White Photograph\\ of the Cosmos}

Four snowflakes ride brown wing feathers\\
to earth, marking their space in the cosmos.\\
A white satin sheet covers the ground.\\
All is slick white ice. A crunch before the slip.

White ground rises up to meet the muddy sky.\\
The sky turns black at midnight, crinkly,\\
blank as a big sheet of construction paper, the kind\\
children have drawn stars on since the beginning of time.


\clearpage
\section{Blonde Showgirls}

When I narrow my field of vision,\\
squint a bit, I see only daffodils,\\
blonde heads lined up\\
in the roadside ditch by my house\\
where drivers toss empty beer cans\\
and McDonald's bags full of trash\\
that seem to disappear under paper\\
petals spread to conceal what is\\
out of place, what will not grow.\\
I will not pick these daffodils,\\
will not disrupt the show.

I preserve yellow petals\\
as they drip from the sun.\\
They melt the winter snow\\
storm raging in my head.\\
With my eyes shut, I can't see\\
daffodils, can't see anything\\
except the darkness they come from.

My gaze penetrates the surface\\
soil, slices deep, all the way down\\
to the roots that anchor the bulbs\\
in place. They sprout brown\\
underground. Stems burrow upward\\
until they burst the crust with buds\\
like long-legged women in a chorus line\\
invisible\\
until the curtains part.


\clearpage
\section{Blue Chenille Tales}

I nudge my ratty-tatty bathrobe\\
from the white plastic hook\\
behind the bathroom door.\\
Worn blue chenille slides\\
across my bony shoulders,\\
thins down my back\\
and falls just below knees\\
bent repeatedly against wood floors\\
when my grandsons raced tiny metal\\
cars around corners, chased whatever\\
lingered around the curve. I, too,\\
roared my souped-up Mustang,\\
behind theirs, my knees slower.\\
Back then my robe ripped\\
on an exposed nail.

I finger the blue torn fabric,\\
reel as if caught in a whirlwind.\\
A spiral of memories spin me,\\
wrap me in lightning flashes\\
of stories found in the folds of my robe:\\
Blackberry jam stains the right sleeve,\\
dragged across the sweetness of surprise\\
when my husband flashed two tickets\\
to Hawaii for our 40\textsuperscript{th} anniversary.\\
My daughter's tears rained\\
salt on my shoulder\\
the day she left for school.\\
Coffee stains sloshed and dribbled,\\
marked all those mornings\\
Mary Jane slipped across my grass\\
into my kitchen. I admit\\
I lost a button, stitched a seam,\\
removed spots, washed that robe\\
until almost all the blue is gone.
\clearpage
I wrap my history around me,\\
glory in its blue chenille tales.


\clearpage
\section{Blur the Edges}

That splotch of black\\
speckled with orange spots\\
pretends the dog flames\\
the fire. Watercolors run\\
while the scruffy mutt gallops\\
across fields and its silky mane floats\\
in the sky where an upside-down tree\\
roots in the cloud. Leaves\\
round as plump clocks ready to drop\\
a newborn sparrow,\\
a tiny blotch of brown. Tick-\\
tock before noon. The child's fingers,\\
wet with color,\\
curl around wood,\\
tilt the brush\\
so edges smear,\\
lap the shore\\
and kiss the lips\\
of this child\\
who paints\\
his own swirly world.


\clearpage
\section{The Body Cast}

Yesterday,\\
You broke all my bones\\
and patted wet plaster into place\\
promising with kisses, hugs, and sweet words\\
that everything would be better.

Today,\\
you cut off the casts\\
and exposed grotesque flesh and limbs\\
crippled from confinement.\\
There I stood---a parody of your dream woman.


\clearpage
\section{Box of Disquiet}

If you were here\\
you would see the bruises,\\
purple blossoms on my knees.

For you, I knelt all morning,\\
tugged weeds from April soil\\
so they would not seed again.

Nightshade roots dried white\\
in heat like your lies\\
wilted into the warm room where we slept.

I still remember the season of snow\\
when you left, how your tail lights blazed\\
red as two tulips against white.

Yellow daffodils opened at noon today\\
while I sipped chamomile tea, hoping\\
yellow petals would bring you back.

Even though I rid our garden of winter debris,\\
my heart is a box of disquiet,\\
buzzing with honey bees. Like you they sting.


\clearpage
\section{The Boy in Blue Boots}

the boy in blue boots\\
kicks stones into mud puddles---

water drops lead home


\clearpage
\section{Breakthrough}

No one called him Richard ever,\\
never called him anything,\\
could not even remember him,\\
but, then, he could not\\
say himself who he was\\
or what he wanted,\\
only that he did not want\\
to be who and what he was.

For an hour each week\\
he was mine, a case study\\
I would write\\
after I heard him out.

``Richard,'' I asked,\\
``what happened today?''

``Not much,'' he said.\\
``I punched out my boss\\
when he docked me one hour,\\
as if he had the right\\
to clock my life.\\
Who does he think he is?''

``And who are you?''\\
I had the answer\\
written in the chart,\\
the diagnosis\\
of this man named Richard\\
who never had a father,\\
who was passed from place\\
to place and acted out\\
or acted in. He was lost\\
and no one cared but him\\
and me, of course.

Last week, Richard dreamed.\\
``In my nightmare, I was buried\\
alive in wet cement\\
that hardened, gave me shape.\\
I punched my concrete shell\\
until it cracked.\\
Then I could breathe again.''

I, too, have had that dream\\
many times. I need Richard\\
to tell me how\\
he inhales and exhales\\
when the world is too much\\
with him, with all of us.


\clearpage
\section{Bronzed Baby Shoes}

Above the open loops\\
of laced up, bronzed\\
baby shoes, her mother\\
could conjure the whole.

The curl of birth, the wobble\\
steps, the shush of leather,\\
the slide ahead before the fall,\\
the awe of it all, the sudden hush.

Two bare legs mosquito bit,\\
a waist, the reach of arms,\\
peaches picked and dripped\\
from her baby's lips,\\
hair static with electricity.

Purple rhinestone heels\\
tapped against Elvis songs\\
as her baby's pelvis tilted\\
into purple satin pinch.

Even when her baby\\
left for hotter states,\\
her bronzed shoes poked,\\
jabbed her mother's\\
memory of soft\\
shaped by\\
years of hard shells.


\clearpage
\section{The Brown Spotted Dog\\
  on the Dining Room Table\\
  Last Thanksgiving}

dipped his snout into the blue-rimmed bowl\\
of mashed potatoes, then snatched a turkey leg\\
and gnawed and growled as he hoarded the juicy\\
bits between his front claws. No one from the family

or even guests dared say anything unkind. No one\\
fussed or shooed the dog off the table. No one mentioned it\\
in their bountiful words of thanks. No one even saw it\\
except maybe in their minds when my sister

giggled at some memory of having\\
a white purebred poodle that rolled in horse manure\\
in Mama's garden when she was into girly pink\\
tank tops at fifteen and disowned the wayward poodle

while no one else recalled such a dog. Maybe\\
her brother or uncles were the dog lovers.\\
They once went to the pound, they agreed,\\
and brought back an ugly mutt with floppy ears.

Both turned out allergic to dander (or so they thought\\
today) but were grateful, indeed, the rescued\\
such a misfit even for one starless summer night\\
as they laughed now at what boys did back then.

My papa contradicted the boys (now men) while he cut\\
turkey breast into small pieces for my palsied aunt\\
who remembered owning a goldfish she won\\
at the county fair but never a dog, never ever a dirty dog.

I witnessed all the passing of butter and broccoli casserole,\\
all the retelling of family stories weaving us together\\
at Thanksgiving, another year when no one dared acknowledge\\
that brown spotted dog on our table gobbling up the best parts.


\clearpage
\section{Camellia}

waxed every limb:\\
two arms, two legs\\
hairless as four river stones.

Even her platinum mustache\\
which no one could see but her\\
and twin eyebrows\\
she whisked away\\
with chemicals\\
that puddled,\\
sunk through open pores\\
and burned her roots.

All this fire dammed\\
behind pink eyes\\
stoked her clay-red Georgia heart\\
baked in this clay-red Georgia heat

and scorched those souls\\
who dared\\
sit on funeral folding chairs\\
buttocks pressed against polished oak\\
when Camellia preached

re-creation of ourselves\\
in our own sweet scent\\
in our own sweet time.


\clearpage
\section{The Cape Verde Hurricane}

\emph{September 21, 1938}

Radios blared,\\
promised rain,\\
the hurricane was lost.\\
At equinox, Apollo sailed into the eye,\\
his twin, Diana, at his side. They waltzed\\
the seas from Cape Verde, hiding in the bays,\\
the same dance floor as 1815\\
their feet remembered the way\\
through New London and Stonington,\\
even the great South Bay.

Apollo breathed deep, spit out the salt\\
that stained church windows in Vermont.\\
Diana wrapped women in clouds\\
and pulled them to her breast.\\
Python rose from the washed-out roads\\
and lurked in lofts\\
until Apollo's arrow split the serpent's skin\\
and people prayed, said their rosaries.\\
At Napatree Point, cottages crumbled\\
and bath houses washed to the sea.\\
Parrots played uncaged.\\
People clung to shingles\\
sat atop the trees\\
and fox trotted with the gods.


\clearpage
\section{Case Closed}

Turn out the light, mother hollers. Windows slam, doors lock on the
first floor, shut tight against hitch hikers, rabid bats who fly by\\
radar. At 12, the midnight age, I open shades she pulls, call forth
witches and the Seven Sisters from winter skies.

Heads bent low, we read mysteries. Under sheets bleached white with
dreams, we spot clues sprinkled like cherries over mother's\\
buttercream frosting. We dig for chocolate cake buried, layer upon
layer. It crumbles in our mouths. We guess it is the short bearded
stranger or even the vicar, never the girl in red polka dots. Motive is
all, we say. Where is her intent? We like her, recognize innocence. On
the last page, her sentence is life.

Mysteries crop up everywhere at school. Like what dress to wear\\
or not wear for luck. I solve algebraic equations, where \emph{x} equals
\emph{y} squared and unknowns become known. Answers are right, teachers
say, or wrong, depending on the answer key. I hope for multiple choice,
a chance. When I check Patrick's papers, I give him a couple of points
for trying. Trying to kiss me last night, with Mother flicking the
porchlight off and on like some lightening bug. Still, his fingers touch
bare breast and rest, a period on the page.

So much blood flows every chapter. Hard-boiled PIs, names like Mike
Hammer and Sam Spade, go it alone. Tough guys, quick with their fists,
they pound flesh and pavement. I fall in love with these characters.
They've seen it all. And, they have done it all. At least that's what I
tell Thomas when he wants to know why I quit my job and how we expect to
pay our bills, much less make it and get ahead. Ordinary questions, I
guess, from someone who prefers reading biographies of dead men, models
to live by.

Locked room mysteries confuse me since what can't be is. The key usually
is in the victim's pocket. When Thomas hires a security firm to install
electronic sensors, I can never remember the code. Mother says I'm lucky
he's such a good provider. He puts deadbolt locks on her front door and
back, showing her how to work them fast in case of fire. At my house, I
set off alarms, punching in wrong numbers. Finally I'm fed up with all
the noise. No one understands why I want a divorce. Of course, he says,
you'll change your mind.

Miss Marple clicks her knitting needles, knowing all the time. Liars and
murderers are no more than microbes in a glass of water set on the
kitchen sink. I drink gallons of well water, live dangerously.


\clearpage
\section{Casino Confession}

Dark does not descend\\
in this house of neon lights\\
where I play all the angles,\\
grab my chance at life.

My eyes do not close\\
against the probability\\
that I may toss snake eyes\\
when the red serpent bites.

Venom pumps in my veins,\\
the promise of a lost paradise\\
I must regain. I shake two die\\
and watch them as they fall.

I confess I am no innocent.\\
I have slept with Lady Luck\\
more than once on white sheets.\\
When I woke, she was gone.

I search for that woman\\
who abandoned me when I held\\
all the hearts. Flush with losing,\\
I will lure her back again.


\clearpage
\section{The Catcher}

Hunched over,\\
his text cradled to his chest,\\
the student scuttles sideways into the room,\\
a ghost crab.

I sit on the desk,\\
dangling my feet in\\
seas of books I read\\
as a teen, a fish hooked\\
on Holden Caulfield.\\
I, too, waited\\
for the worm to wiggle,\\
spotted plastic flies,\\
called them phony.

``Did you,'' the ghost crab asks,\\
``run away?''

``In my head,'' I answer.

He buries his eyes\\
in pages.\\
Words at high tide\\
wash his cheeks.\\
He catches\\
specks of sand.


\clearpage
\section{The Cats}

Poe's raven led them.\\
Of that I'm sure.\\
Those two cats outside staring in\\
through black and glass\\
electric eyes green before the flash\\
heating these arthritic hands\\
until they melt\\
red against students' words\\
that come to life\\
like stick people\\
and loom large\\
above my head.

Those dripping, purring words\\
like claws against my skin\\
begging, pleading for a way in.\\
Angora strands stroke my throat.\\
Those gentle sounds are soft meows\\
warm against my breast.

Those two cats the raven brought\\
crept into my house.\\
Their lightning eyes\\
illuminate\\
the words\\
my students said.


\clearpage
\section{Caught Driving at Night Without Glasses}

Just like that\\
death zips up beside us,\\
flashes candy stripes of red/blue light\\
strokes the night\\
and streaks our eyes. As if

Our radar detector malfunctions,\\
doesn't register the signals sent\\
by the cop behind the wheel. The one\\
who cruises beside us or waits ahead\\
under the overpass, wrapped in its shadow.\\
That officer wants off traffic patrol,\\
tired of entrapments and tickets he signs\\
with his name and ours. Our infractions

Are merely minor violations. Nothing even\\
criminal about our refusal. No, sir.\\
We preserve our rights. We can\\
not breathe into the breathalyzer,\\
not strap ourselves into seats that hold\\
us down. Snap and unsnap. As if\\
it were decreed somewhere

that we must wear our glasses, must\\
magnify the 12-wheeler zooming toward us,\\
blot out its headlight halos. As if\\
we need to update our driver's license,\\
appear in court on the date written,\\
lock all doors against carjackers,\\
press the accelerator only halfway down. No,\\
we will not be caught in his jurisdiction. We will look

out windows and see the woods. We will\\
count bristlecone pines and red maples.\\
Watch dinosaurs as they lumber\\
into the clearing, winter wheat caught\\
between their toes. Must be the heat

that drives them out. Or the alligators\\
with hooded eyes, back again. Our woods\\
are greener than their marshland. Forest\\
green. Too green for alligators. Maybe\\
hunter green. We pray for rain

and dinosaurs. Their breath,\\
moist with maidenhair ferns,\\
mist our window pane.


\clearpage
\section{Cherry Poem}

Through the black-stick winter,\\
I snapped words underfoot as I waited,\\
watched for the burst of white buds\\
the cloud of cherry blossoms\\
pale as my poems, no more\\
than fragile petals I picked\\
and flung, snow in spring wind.

When the cherries came,\\
I rushed to pick them\\
before they dropped,\\
before the cardinals claimed\\
these red tart poems as their own\\
globes of wild life\\
plucked from my word trees.

Each cherry was not a cherry.\\
In my mind, beneath the skin,\\
the pits grew orchards of poems,\\
spurts of pages clustered at the ends\\
of branches, a dangle of words,\\
a flash of green visions fruited cherry red\\
with glossed stanzas as far as the eye could see.


\clearpage
\section{Chicken Bones}

Go where fear lurks, back\\
behind the garage, hunkered down,\\
no more than a dog pound mutt\\
digging up chicken bones, too fragile,\\
too sharp for any dog, for any human who might\\
gnaw through that skinny thigh bone,\\
yearn for the crunch before the final break.

Wear these blanched bones in pierced ears, dangle\\
them in sunlight. Let every friend ooh and ahh\\
as they caress such slippery truth hanging loose\\
enough to pinch between two slim fingers.\\
Such a gem bone tantalizes, tricks thieves\\
into thinking they can heist something\\
beyond the pointed toe of fear.

Rub bone edges smooth so they don't snag\\
the wool pockets of anyone's red plaid mackinaw.\\
Don't tangle the dog's hair after the groomer brushes\\
each strand clean of burrs, shapes butterscotch curls\\
along the ribs, marks every vulnerable spot with gold swirls.\\
This dog guards chickens as they eat. Their beaks\\
peck grain and puffs of bone dust from the dry, dry dirt.


\clearpage
\section{A Child's Game}

Little sister, you are it.\\
Your turn to shout out sunlight\\
while crows flap wings.\\
With plump fingers splayed,\\
you peek at ragged bark,\\
count numbers out of order.\\
Ready or not

you warn me\\
as if I am not\\
older, as if\\
I have not\\
played this game. I\\
who ran out of bounds\\
when father left us living

alone with his blood, his whiskey\\
wife. Our mother hid\\
in a bottle, a secret\\
message to us, if only\\
we knew where to look,\\
how to wake her, how to\\
break brown bottles

labeled with a crown we\\
would not wear until later\\
when we rousted crows, tagged friends\\
hidden on dance floors, in red\\
vinyl booths or at Formica tables\\
in plain sight waiting\\
to be found by me, by you

by men who knew\\
how to count to one hundred.


\clearpage
\section{Chroma: Gloria Colors Her World}

Gloria spins the globe\\
on her office desk,\\
faster and faster\\
until blue oceans gobble\\
green land. Whole countries blur,\\
longitudes and latitudes merge.\\
No equator circles earth.\\
Such a plastic world tilts, gains\\
momentum. The whirl of speed\\
sets it free. Nothing holds\\
its place. One sphere\\
jumps its orbit.

Gloria, the gray dove, plans her retirement, her wing-\\
flapping revolution. She opens\\
white Venetian blinds, saturates her\\
office with morning light.\\
Her faded flesh turns pink.\\
Her fingers move rat-a-tat-tat\\
across the keyboard. Her nails\\
frosted robin-egg blue, not natural.

Gloria flings one fist\\
full of rubber bands\\
across gray carpet.\\
She empties boxes\\
of brown circles\\
she once stretched\\
over piles of reports,\\
binding black words on white\\
paper, filed in beige drawers\\
alphabetically trapped in darkness.

Gloria whips off her sapphire\\
blue jacket she bought for today,\\
spreads it over gray carpet.\\
She scoops rubber bands\\
onto silk lining, fills two arms,\\
three pockets, one back with elasticity.\\
Each band soaks up one color: emeralds\\
she dangles from her ears, garnets\\
on each finger. A flurry of color\\
lurks in light when mundane rubber\\
bands snap Gloria back to life.

Gloria weaves a band of red, a crown\\
for her feathery gray hair. She tucks\\
her blue world under arm, struts\\
office aisles, tosses color everywhere.\\
She flies out into open space, her birdsong\\
lingers: Glo-ri-a, Glo-ri-a, Glo-ri-a.


\clearpage
\section{Collage}

Rising from the streets,\\
the yellow fog cradles\\
bits of voices:

curse words, shrieks,\\
prayers, absolutions.\\
Confusion

balloons in her brain\\
in her third floor flat\\
where she rocks,\\
clutches her cat,\\
strings street words\\
into nonsense lullabies.

She ignores dust,\\
settling on her windowsills

while he struts the streets\\
costumed in yellow, purple,\\
green. All the time, he

fingers the switchblade\\
warm through his pocket\\
against his thigh,\\
remembers edges\\
sharp as butcher knives.

Forgets. He hums off key\\
a mating call\\
to whores who turn their heads\\
toward storefront mirrors\\
and vacant buildings

while the newborn\\
Salvation Army Band\\
dressed in heavy coats\\
blast their trumpets\\
ring their bells\\
count their change\\
tinkling in the brass bucket.

``The Savior will come,''\\
or so they say\\
to the kid magician\\
conjuring cards\\
for the salvation crowd,\\
collecting money\\
meant for the bucket.


\clearpage
\section{Collector of Petty Grievances}

My grandfather hand-fed each slight\\
as if it were a gerbil, a round\\
ball of fur in his pocket he could\\
stroke, feed a fistful of sunflower\\
seeds until the creature expanded,\\
exploded. With each telling\\
he added a phrase, another reason\\
why the world took against him:\\
a man who did his best,\\
a man who never gambled\\
except on luck.

Always, he fretted\\
the mill might close\\
before he filed his grievances.\\
His supervisor chewed him out\\
for steel shavings on the floor\\
when he punched in late\\
after his lunch of spoiled tuna\\
which he never liked, even when\\
it was on white bread that\\
would go down better with\\
a beer he couldn't have. His\\
superintendent called the union rep\\
when his work slowed and he\\
threw a punch at no one\\
in particular, just whoever\\
got in his way which was everyone\\
back then when he was too young\\
and had no luck and now he had\\
one grandson he didn't really like\\
and a pocket full of gerbils\\
he poked and coddled.


\clearpage
\section{Come Here a Little Closer}

In my father's closet, I find\\
bits of you in old Polaroids\\
faded from black to brown,\\
stained by the many hands\\
that held these photographs\\
exposed in sunlight,\\
stored in shoeboxes\\
rippled from damp\\
days before I reclaim my life.

You are now a shadow shot\\
without features, a familiar\\
shape yet ghostly as if seen\\
through gauze. I would draw\\
you back, color your eyes\\
black as watermelon seeds\\
or were they a lighter shade\\
of brown, more like\\
peach pits and apple stems?

I smell your rose fragrance\\
from when you leaned\\
your head to mine,\\
when you read Golden Books,\\
told me I could do it all,\\
tugged my braids, teased me into bed\\
laughing, carrying the scent\\
of peanut butter cookies warm\\
from the oven every day, our snack.

I remember the purple sundress\\
you wore on my fifth birthday\\
and the chocolate cake with my name\\
on top. I blew out all the candles,\\
wished for a dress like yours\\
and your long straight blonde hair\\
or did it just touch your chin\\
back then? Had it turned to brown?\\
I cannot sketch your mouth, your nose.

You, my mother, were gone for good\\
before I started school. I felt your hand\\
in mine as I walked through that door,\\
vowed you would always roam my memory.\\
I painted my mouth with your red lipstick,\\
creamed by body hoping to bring you back.\\
Even photos blur, have wavy edges.\\
The acid bath of time burned your image\\
from my mind. Please, come a little closer.


\clearpage
\section{Continuing View}

Pretending is a delicate wing\\
layered with ebony feathers.\\
You are crow in air. You see\\
below, above, beside yourself\\
all that is and can be. You

become that sage green ribbon,\\
weave a fluid fence. Your tapestry\\
sprouts trees that reach toward yellow,\\
curve toward shades of life. Lavender\\
hues, all purple and red. A puddle\\
of children. Call all\\
crows. Glossy feathers perch in trees,\\
speckle air, draw close\\
as children stoop to ground.

Begin the lesson of feathers.\\
These children open eyes.\\
Their pupils dilate. They continue\\
their view through veined leaves\\
up close, transparent and forever real

until they blink. Dark lashes\\
skim their cheeks silent as a wing.


\clearpage
\section{Cross Current}

The undertow yanks me\\
just below the surface. Sucks me\\
back to birth, that long canal\\
of dark. I wiggle to light,\\
crawl to beach. Rest where\\
I am defined. A woman wreathed\\
in seaweed, damp with promise.

Like you, I have a soft body, hard\\
mollusk shell on my back. I can\\
creep on shore, retract my head, leave\\
a trail of where I've been. I have once\\
risen from the salty brine, the ocean floor\\
safe, now with deadbolt locks on every door.\\
I hang suspended, a shell of myself.

At midnight, I will come where you are,\\
baptize myself in ocean. I will drop fathoms\\
to where you swirl in quicksilver cross currents.\\
We will bathe ourselves, braid Neptune's beard.\\
Let us embrace a wayward current, etch its moves\\
on our armor. We will jitterbug a storm\\
full of white lightning and thunder words.


\clearpage
\section{Crouching by My Desk\\ on the Fifteenth Floor}

Ready to spring, sneak out\\
mitered doors, hinged to swing\\
outward. I freeze in the glare,\\
fluorescent headlights hold me\\
there, a doe dodging the hunter's bow.\\
At dusk, phones ring, machines answer\\
in bullets, ripping flesh.\\
My voice spouts numbers.


\clearpage
\section{Cupcake Cathedral}

The bishop of chaos reigns.\\
His scepter is a peppermint stick,\\
his puffed hat a marshmallow.\\
He anoints his congregation\\
with soda fizz miracles,\\
clears stains. He will not

hear confessions of gluttony\\
or counsel hyper spirits fueled\\
by sugar: white, refined. Sweet\\
heaven on earth, two cupfuls\\
per cupcake. We praise confectioners:\\
buttercream frosting a sin

we commit. We redeem ourselves,\\
bear witness in this sacred place.\\
Helter-skelter, we rush the altar\\
where the bishop stands, a potholder\\
and cupcake tray in each hand. We supplicants\\
lick crimped paper clean. Each crumb consumed

sparks visions. We speak in tongues,\\
a babble of birth and maroon spires rising.


\clearpage
\section{The Curious Affair\\ at Sampson's Used Bookstore}

I bought a bagful of murders\\
half price, pages dense with clues.\\
The blonde behind me\\
sucked lemon drops. She nudged\\
\emph{Death on the Nile} along the counter,\\
the same book\\
I traded last month.\\
My splash of black coffee\\
still scarred the cover.

My worn recliner squeaked\\
as I pushed back with \emph{Dead Man's\\
Folly} in my lap. A later Christie\\
mystery with Hercule Poirot.\\
I exercised my little gray cells.\\
No name inside the cover.\\
A faint sniff of lemons.\\
A blonde hair at the start\\
of chapter two. Darker\\
at the roots. This woman lingered\\
today, behind me\\
at Sampson's Used Bookstore.

Pale pink lipstick smudged\\
the upper corner of one page\\
as if it were licked open by this blonde\\
greedy for words on her fingertip.\\
I, too, chewed these inked phrases,\\
pressed my prints against hers.


\clearpage
\section{Daylight Savings Time}

When the light goes on\\
and on, refuses the moon\\
its rightful place alone,\\
I inhale scorched stars falling\\
unnamed, lost

and found by the arson squad\\
sifting through ash and serial killers\\
who month after month in Detroit\\
crave the ritual death, and April hour\\
snatched in spring, returned in fall,\\
too late. No identity. No challenge\\
for those who dare tinker with time

impose what is false. Those bureaucrats\\
sprint for office, fast\\
off the mark and breathless\\
as their minds turn back the hands\\
and alarm clocks lie about waking

and sleeping. I rise in fog,\\
in rain and darkness, maybe.\\
My dreams are a dull pulse, no\\
more like a forgetting, more like\\
a lost hour when a woman\\
who resembles me

meets a man who takes her life\\
in that abandoned hour.


\clearpage
\section{Death of the Bird Child}

As I slide one sunny-side-up egg onto her plate, I smile\\
an upside-down frown, at least\\
from her view looking up at me looking down. Her five-year-old\\
arms sprout red down.\\
She is becoming a robin, a sweet warble of a girl, my girl who\\
wants to fly high, higher\\
than a tree, higher than the sky itself. She collects feathers.

All along the beach, my bird child dips and rises. Her pockets spill\\
yellow, green, purple\\
feathers, especially black. The glossy black of a loon's wing feather.\\
Grey fluff from the\\
chest of a ring-billed gull. A splash of orange from no known\\
species. A long tailed\\
duck dives for a round goby, a flash of black and brown fish scales.

Within weeks, feathers wiggle with bugs and slugs. Wet,\\
moldy feathers drip from\\
mounds of bodies. Dead from eating gobies, dead from botulism.\\
The common loons,\\
long-tailed ducks, gulls, white-winged scooters, red-necked grebes,\\
red-breasted\\
mergansers, double-crested cormorants have all fallen down from the
sky.\\
The rot along the\\
shores of lake Michigan, decay\\
at my daughter's feet. That summer my girl threw away\\
her feathers.


\clearpage
\section{Deathwatch}

Prognosis poor. Scalpels slice,\\
cut statistics any way.\\
Just hone the blades, razor-sharp,\\
metal scraping stone. Faster,\\
grind edges precise\\
as doctor words. We can

cauterize old wounds. Let's burrow\\
into ourselves like earthworms,\\
into the darkest loam\\
never turned, fertile still.

Dig deeper. Past sand, clay,\\
shale stained red. Veins\\
of green, of gold paint promises\\
in stone. Colors pale under ground, under\\
words crisp as the earth's crust. No

more skulking behind skin. Shed the stuff\\
that holds us in place, in time.\\
merge with me, a single cell.

Let's circle the world. Leave the tropic of Cancer,\\
cold. Scuttle south:\\
a zodiac crab, tail tucked, a major constellation\\
brighter than the fourth sign. Nurses can chart

malignancies and biopsies. All positive\\
negativity spread on a slide,\\
magnified twenty times. We will

channel through granite, open passages\\
straight to the iron core.\\
Liquid thick as lava, hot,\\
erupts: Come ride the waves

white as shark's teeth,\\
black as raven feathers.


\clearpage
\section{Delivery}

Days passed hang\\
heavy on my calendar.\\
Pregnant women\\
outwait due dates.\\
Edges curl.\\
Red magic-marker months\\
newborn pink, fresh\\
from mother's womb.\\
Stretch marks, stitches\\
across bulging bellies\\
shrink now.\\
Years are so much\\
afterbirth.


\clearpage
\section{Dementia}

All direction is lost. You move in concentric circles, ever\\
closer to the center, your earth orbit charted, hung at the end\\
of your hospital bed.

You linger, some ancient memory calling. You glide through\\
earth-shadow, await the total lunar eclipse.

I cling to your copper glow, fear reflected sunlight, the new\\
moon rising.


\clearpage
\section{The Demise of Eloise Insane Asylum}

\emph{In Memoriam, 1832-1981}

Windows never closed tight enough\\
at Eloise where night slipped under sashes,\\
carried the odor of skunks cornered by a dog\\
dumped in this wilderness where wood rotted\\
and brick crumbled inward like my grandmother\\
who once lived there before she died an insane death\\
full of empty spaces she swallowed with her final breath.

When I was 10, I visited my grandmother five stories up\\
inside the walls, below a clouded sky, above the place\\
where fire engines waited for the burning time. The powerhouse\\
spat smoke that rode my shoulders past the screaming man\\
who reached out to me but Mama yanked me back,\\
back from all humanity as if I might catch what they had,\\
as if I might drag home the musty smell of death.

On the highest floor, in the furthest room, we found\\
my grandmother huddled in the darkest corner\\
where she stared straight ahead and spoke to no one\\
except us who knew she wanted to go home, wanted\\
nothing more than her dress of morning glories she wore\\
on summer Sundays when she hit all the high notes. Her\\
voice once raised in celebration now slides down the scale.

Bit by bit, Eloise lost pieces of itself.\\
The wrecking ball swung death at the end\\
of its steel cable while late night lovers carved\\
their transient state of craziness on walls and floors\\
about to fall as land reclaimed its ground and sprouted\\
green leaves with ivy twined like voices harsh with hope,\\
ragged with longing for one last chance to save themselves.


\clearpage
\section{Detroit}

Rising from the streets,\\
the yellow fog cradles\\
bits of voices ---\\
curse words, shrieks,\\
prayers, absolutions.\\
Confusion\\
balloons in her brain\\
in her third floor flat\\
where she rocks, clutching her cat\\
against her shriveled breasts,\\
stringing the street words\\
into nonsense lullabies,\\
ignoring the stillborn dust\\
settling on her windowsills

while he struts the streets\\
costumed in yellow, purple,\\
green, all the time\\
fingering the switchblade\\
warm through his pocket\\
against his thigh\\
remembering fear\\
forgetting\\
whistling into the night\\
a love call\\
to whores who turn their heads\\
toward vacant buildings\\
as he slinks by

while the newborn\\
Salvation Army Band\\
dressed in heavy, navy coats\\
against the darkness\\
blast their trumpets\\
ring their bells\\
count their change\\
tinkling in the brass bucket.\\
``The Savior will come!''\\
or so they say to\\
the kid magician conjuring cards\\
for the salvation crowd,\\
collecting money\\
meant for the bucket.


\clearpage
\section{Diary of a Night Watchman}

In my nightmares, it is noon.\\
the sun sags,\\
round as a pregnant woman,\\
flushed with sin, silent,\\
waiting to drop\\
behind clouds or ride\\
elevators to the top\\
floor where I can't track

her footprints, crushed in carpet,\\
which lead to unlocked doors. She\\
smiles through mail slots, vanishes\\
at five o'clock, full of life.\\
She ferrets out riverbeds,\\
drinks carp and seaweed,\\
her belly a fish bowl.

On my shift, I turn out lights.\\
I beat my head on concrete walls,\\
read scripture aloud. My own\\
voice rises to ceilings,\\
splinters giant redwoods\\
rooted for centuries. Behind bark,\\
concentric circles mark off ages.\\
These rings, like her, forget the axe,\\
the prophecy of bones.


\clearpage
\section{The Diet}

I'm going on a diet tomorrow.\\
No more frosted, spongy words,\\
sauced phrases, or\\
mashed truisms that soften\\
my sharpness. I'll snack,\\
instead, on crunchy, cuss words\\
and lean, bloody images\\
until my bones poke through\\
my skin and people see\\
the stuff I'm made of.


\clearpage
\section{Dissonance}

I answered\\
nothing, not doorbells\\
or phones ever silent\\
since I lost myself

when cash registers jangled\\
me deaf at Piggly Wiggly\\
where I cashiered for 20 years\\
too long, grew used to dark

green money clutched by hands\\
callused, knotted with worry\\
at rising costs of bread and milk\\
without honey which I gave for free

when no one asked but the manager\\
who fired me for giving too much\\
when all I yearned for was a way\\
home, where ivy flourished

teemed with life, sprung from dirt\\
unnourished, tough as its own green\\
leaves shiny with hope, with climbing\\
brick by brick upward and over\\
my window, lush as eyelashes

each leaf winked in wind\\
trembled in storms I knew\\
its roots would hold\\
no matter what

when ivy green rustled a jazz tune\\
I answered with rain\\
sliding from tone to tone\\
improvising leaf to leaf


\clearpage
\section{Distances}

You, in the black leather chair.\\
I, on the floor mounting photographs\\
while TV people drag\\
their leaden legs through scripts.\\
Your video voice measures distances\\
across frozen ponds, between titles and credits.\\
My winter eyes check your marks\\
against the glassy surface.

Our dubbed voice-overs:\\
marvelous, mechanized genies\\
rising, falling, cursing the day we joined\\
like Siamese twins\\
in our own circus\\
tearing at our flesh\\
even as we touch.


\clearpage
\section{Dispenser of Good Will\\ and Spotless Garments}

Mr. Thomas Granite piles his stained clothes\\
on my stainless-steel counter at Glendale Cleaners.\\
Juicy plum lipstick greases one shoulder\\
of his wrinkled, white shirt\\
entwined with the blues and yellows.

A blotch of red wine taints another\\
shirt that reeks of cigar. Blue ballpoint\\
slashes diagonally across the right side\\
of his striped shirt as if he had\\
awarded himself a blue ribbon.

Mrs. Nora Menton, a pale newlywed,\\
plumps the heap of pearls, satin,\\
and tulle. She will pay to have her dress\\
preserved as if her wedding can be\\
folded into a box and sealed.

She may not even notice\\
black shoe polish on the soiled hem,\\
the ripped train she may have forgotten\\
to detach or hold when the music blasted\\
her into the curved arms of other men.

Mrs. Arthur McKinnon, no first name\\
of her own with her husband still alive\\
although his hands wobble when he eats,\\
so she told me once, and he drops\\
bits of Sunday's pot roast on his wool pants.

As she smooths each garment\\
into shape on the counter,\\
she tells how ever blotch happened\\
when her husband's attention wandered\\
and how difficult it is to get him back.

I examine their dirty clothes,\\
pin on tags, pre-treat spatters, clean what\\
they wear, press garments flat before\\
I give them back just like new.


\clearpage
\section{Diving Lessons}

My sister dives off the high board at ten,\\
head first on a downward trend. Her body arcs,\\
an apostrophe in air. She hovers there\\
above the pool, her body slick with passion.

Under blue spruce, she sheds her skin for men\\
who wash over her, zip her into wetsuits\\
dark as pubic hair. She walks on ocean bottom,\\
marks her path with cowry shells, a ghostly galleon.

At fifty, my sister hits a drought year\\
hard as a bellyflop. Her arteries clog with sand,\\
with men who dig dry ditches and stand guard\\
with tubes and latex gloves, their second hands

plunge into unlit caves. Her eyes sink inward,\\
two divining rods pulling her toward salt springs\\
where she bobs with pirate lovers who baptize her\\
in the name of the red moon at night.


\clearpage
\section{Divorce}

She sold the washing machine\\
the dryer the cups the sheets\\
even the bed---sliced mattress\\
and all except the letters\\
no one bought or took for\\
free: even the garbage man\\
returned them tied\\
in butcher's string.


\clearpage
\section{Down a Long Hall}

Father makes sure everything is in good repair at our house,

the back screen door barely squeaks as I slip out.

Into the night, my black hair dark as a black bear pelt,\\
my brown eyes fertile as the trail we tread, my arms barely able\\
to grasp the daffodils with their trumpets blaring.

My blond doppelgnger will not go away. She appears\\
in the right-hand bottom corner of every screen. Her outline is\\
stronger, darker, more threatening with each intermission.

I pull the shades, turn her off the big TV and computer screens and\\
cover the mirrors. The blond scampers down a long hall,\\
disappears into the corner.

As my father turns to face me, I see he has become Zeus,\\
gold flecks in his arched eyebrows and severely cut beard.\\
His lance of lightening lifted, ready for the blond.


\clearpage
\section{Dream Forest}

I dream myself awake\\
in this misted forest,\\
caressed by fog's fingertips.\\
I rub face against bark,\\
wrinkle to wrinkle. Every crease\\
of flesh and wood converge. We\\
speak in oak and ash, the way\\
lovers merge. I nest\\
in your lowest branch,\\
fluff leaves not yet rusted\\
orange-brown or dried to dust.

Here I rest, cradled\\
in your gnarled arms.\\
Droplets of mist\\
collect in cupped leaves,\\
bowls of nectar we sip\\
together. We embrace\\
in this quiet space.\\
I listen to\\
the rustle of you\\
who know autumn days\\
break early and wane.

Explain, if you will,\\
why your branches flutter\\
in wind while I sway,\\
lulled by whistling breezes.\\
A few moist green leaves cling,\\
not ready for the fall.\\
They brush my cheek, whisper\\
of seeds, the growing time\\
come spring when you bud again.

Our roots grow deep\\
in solid ground. Our bodies\\
sweat sweet melodies,\\
as each note rises, drops\\
with songs of us.\\
We reach for moon,\\
a slice of light to hold\\
as tunes blanket us\\
this warm night of fog.

Owls hoot their eerie sound:\\
who, who, who drapes mist\\
like a white shawl\\
warming bare shoulders?\\
Who will wear my wing\\
feather full of flight\\
in her onyx hair?\\
Who walks among trees\\
in dense, wet air?

I do, I answer.\\
I dream each tree,\\
each forest, each night as I sleep.\\
When I wake to day, night sounds\\
echo in my heart, rattle my soul\\
until I breathe in night truth,\\
exhale a cloud of fog,\\
a royal cloak for trees.


\clearpage
\section{Driving Lesson}

I cupped speed under my toes,\\
held it down hard on the accelerator,\\
curved my whole body into the snake turn,\\
heard my father beside me whisper\\
slow down, slower, as the wheels\\
jumped the curb. The man on the ground\\
mowing his grass saw me,\\
dropped the handle and ran.\\
He pumped his legs, jumped on his porch,\\
opened his mouth and screamed loud,\\
called me names, bad names.\\
He spit those names. I whipped past.

My father stretched beside me,\\
lapped his leg over mine, jammed the brake.\\
My chest bit into the steering wheel.\\
My heart flew out the window,\\
a red bird feathered in flight,\\
a flame of fire across the noon sky.


\clearpage
\section{The Drop into Sleep}

My balance teeters\\
as I walk the tightrope\\
of twist your knee and fall.

Heel to toe I totter on the invisible\\
hemp stretched across the yawn\\
and stammer of one foot up, one foot down.

The slush of slippage\\
pools in my silver slippers,\\
drips from one worn sole.

My legs tremble my mind\\
into memories of toes uncurling,\\
an arch falling, the letting go.


\clearpage
\section{The Elephant Man}

The elephant man chases me.\\
His whirling trunk is a lasso\\
dropping neatly over my head,\\
tightening about my neck.\\
He, ever so slowly,\\
lifts me to him.\\
My legs, like palm fronds\\
in the African night,\\
slit the darkness\\
until my heart bursts;\\
the blood flood blinds him,\\
soaks the leathery noose\\
shrinking around my neck,\\
crushing my scream.

Rearing back,\\
he unwinds his trunk,\\
hurls me to the ground.\\
Dead.\\
Until\\
tomorrow night\\
when the elephant man chases me.


\clearpage
\section{Epistemology}

Not knowing what\\
to call those two birds\\
by name, by species,\\
indexed in field guides,\\
lost alphabetically.

In the same way,\\
your name melts\\
on my tongue, tart\\
as cinnamon candy. Red\\
dye stains\\
tongue and teeth.\\
Our talks, our touch\\
too tentative, too full

of nothing but\\
sugar, spun at high speed,\\
puffed with air, not enough\\
to grab onto. Like planes,\\
our marriage recycles oxygen,\\
dry as old arguments.

Just that once,\\
I spot them\\
at the feeder, wings tucked,\\
heads cocked like guns

before metal hits flint.\\
Detroit on New Year's Eve,\\
near dusk. Their saffron crests\\
burst brown night,\\
two lone crocuses above\\
snow. The whole world

upside-down, in flight.\\
I see two\\
birds, feel feathers\\
brush my eyelids, listen\\
for wings fluttering

against my lips, the sound\\
of breathing in and\\
out. Not even words,\\
hard as sunflower seeds,\\
can soften the spaces\\
between us.


\clearpage
\section{Even Wonder}

Woman transforms herself,\\
chops off her hair,\\
pinches strands into spikes.\\
She blends in\\
crowds, cleavage mounded,\\
rounded and held\\
up with Madonna's bustier\\
featured in late night\\
shows she can't escape.\\
Without her tiara, her bullet-\\
deflecting bracelets, she lost her\\
jangle. She is one of us\\
who battles on technicolor pages.

Now she paces Main Street, enters Walgreens,\\
circular in hand, her camouflage.\\
On her own, she goes\\
straight to comics, aisle 2\\
at the end of the rack. She thumbs pages,\\
where Superman pales in fluorescent light\\
and Spiderman spins his web. She will\\
break gossamer threads, free herself.

At sixty, we both court\\
peace, a quiet space\\
where votive candles\\
scented with orange\\
burn black night\\
into day. We will rise\\
refreshed, lick dew from ferns,\\
make our way across Stoney Creek.


\clearpage
\section{Everyone I Know Is Dying}

Even me and my green-feathered parakeet\\
whose clipped wings ground him\\
to this birdseed life.

In Detroit Aunt Gloria sleeps\\
with a steak knife under her pillow\\
just in case she wakes to breaking glass.

A baby boy forces his way headfirst\\
into life behind Miller's Bar\\
where his mother tends the men who gather there.

All muscle, my ex-husband benchpresses 300 pounds\\
at the fitness club in Birmingham where young things,\\
exercising their rights, watch him sweat.

Friends lose parts of themselves:\\
a breast, a bladder, teeth, a bone.\\
Even hearts forget, bypassing what has been lost.

Each day now I cram my pockets full of peacock feathers\\
and watch for iridescent green wings\\
flapping in this hardscrabble year.


\clearpage
\section{Face to Face}

You mapped your face in mirrors\\
for fifty years. Traced\\
every open pore, closed\\
mouth bracketed by worry\\
lines. Two furrows

cut deep between your brows\\
cradled your babies, clutched them\\
close. You swaddled\\
them in your own imperfections,\\
reversed reflections.\\
Skin to skin, they knew

you inside out, out\\
side in

as I did, my friend. Do you\\
want to chart again\\
your whole geography? Change\\
our ways of seeing who\\
you are, where you've been? All

this medical sorcery\\
the cost of skin\\
pulled tight, holding\\
in what should be let out? Do you

choose a pattern\\
in your size? Does it fit\\
with seams, stitches, tucks\\
like a child's outgrown smock?\\
Will it

make your skin transparent,\\
luminescent, hold your death\\
day in abeyance: a distant\\
fluorescent green glow?


\clearpage
\section{Fill 'Er Up}

I left behind my blue sweater\\
and unmade bed when I went away.\\
These familiars spoke of me and this place\\
I was born into without having a say.\\
Here I grew and learned from you, my father,\\
how to:\\
grow crops in the foothills\\
irrigate this rocky land below mountaintops\\
bale hay for the winter months.

The chill of snow shivered my body.\\
My vision blurred. I could not\\
see where sky reached out\\
and touched the land.

I left a note,\\
told you how mountains blocked me in\\
when I needed open space. I hovered\\
on empty.

I fueled myself and my red pickup\\
with regular at Lou's. No high octane needed.\\
Just a full tank of gas could carry my mutt\\
and me past these dusty roads\\
I navigated on the darkest nights\\
without stars.

My tires rolled.\\
I crossed borders,\\
arrived in other states unknown.\\
People asked my destination,\\
that point on the map I headed toward.\\
Sometimes my mutt chose the route\\
as simple as a paw on a squiggly line\\
and we followed.

We saw what we saw,\\
went where we went, my mutt and me.\\
We lost ourselves in open space,\\
on crowded streets, under Red Rock Lake.

If I had been born\\
here\\
or there\\
would the snow still drop to ground?\\
Could I have been content with rain?

Hard to tell if we were coming\\
or going\\
when the calendar read spring\\
From what we've seen of city faces\\
and roadmap places, we were\\
headed back to what we left.

I would sew the button on my blue sweater,\\
tuck the blankets under and fold them tight


\clearpage
\section{Feet at the Funeral}

Follow the prayers.\\
Cross and uncross. Re-cross\\
four hundred legs covered\\
with tweed wool trousers or taupe\\
silk stockings tapered to toes\\
rigid with reverence. Some legs

encased in synthetic fabric\\
waver and wrap one long limb\\
around the other. This ancient dance\\
choreographed: apart, then together.\\
At the chord, tap the carpeted floor\\
twice and move forward, caught up\\
in long, tangled embraces. Their

dance an offering, a sacrifice\\
to distant foot gods. Their best guess\\
at life. They rise in unison

a morning glory vine in bloom. Heart-shaped\\
bodies fill space along the pews. So many\\
leaves. Each face pink, lavender,\\
white and funnel-shaped. Upturned\\
eyes above their open mouths. Ready

for the preacher whose thirty mouths hang\\
slack like wet laundry. His head droops\\
on his neck, dangles above his feet\\
in prayer. Shoes resoled and polished.\\
A black shine beneath his robe. Now\\
behind the pulpit. His hidden feet

are flat against waxed oak floors. Rubber\\
soles slide an stick, a squeak\\
of music, his own refrain\\
drains away.

Spider veins wind around\\
old women's ankles,\\
streak their legs red. Break\\
even as they unlace their shoes,\\
knead flesh, swollen like some\\
exotic flowers. Maybe magnolias\\
or white orchids turned brown\\
around the edges. Their corns,\\
calluses, bunions throb to the tune\\
of the preacher's prayers. Our Father,

who forecast plagues and floods\\
and death to all things on earth,\\
does not wear shoes, does not\\
have feet. Only His son\\
washed disciples' feet

clean of dirt, caked on thick\\
from daily travels up\\
and down the hours. They slipped\\
on pebbles that cut through

flesh, through leather sandals\\
buckled tight\\
against the day of one betrayal, one infidel.

Today, friends, family, neighbors\\
and complete strangers converge.\\
They drag their feet across\\
the sepulchral floor. They\\
search for a foothold, settle\\
for the static charge of nylon\\
carpet fibers created by\\
four hundred feet sliding.

Through the valley of death,\\
bent heads and withered vines.\\
Small, tortured cries and whines.\\
Amen. Amen. Amen.


\clearpage
\section{Final Approach}

Two great egrets whiten the sky\\
with wing feathers, an omen\\
I cannot read in darkness.

I am not a superstitious man\\
who thinks\\
death would hitchhike south.

Like me, these two birds have simply lost\\
direction. They will mate on this island\\
refuge where life writes itself in sand.

Palm fronds illuminate\\
the runway as these two messengers\\
drop their long black legs and land.

These birds follow me into the center\\
of mystery where some say\\
dinosaurs once wore egret wings.

I do not need proof of ancient wood\\
or fossil bones. I have charted the sun,\\
held one blue egret egg in my hand.

Under moonlight, I read my fortune twice.\\
I swallow one white feather,\\
chew a huge chunk of blue sky.


\clearpage
\section{Fireflies of the Manic Moon}

I have reversed the straight line\\
of time, curved it back on itself,\\
traced the trail leading to my child

days when time evaporated\\
at dusk, when fireflies fed my lunacy\\
carrying each crazy dream above

over the manic moon back to\\
where I lived inside my skin,\\
inside the grey-shingled cottage

outside where the soft swish of shadows\\
were moving toward the center, somewhere\\
near the muscle of my heart

which remembered all\\
my body houses, all\\
my body craves:

lust luminescent as fire\\
flies who flash their promise\\
of procreation, their legacy

of life. I have come back to where\\
no map can record the terrain,\\
where daylight gives way to night.

Under the ribcage of this cottage,\\
my memories dangle from curved bone.\\
Hope flaps its wings, feeds its young.


\clearpage
\section{Five Women in Their Baths}

1\\
The white feather of a dove. Soap\\
so soft against her skin, fragile\\
as an eggshell membrane. A splash,\\
a water drop from above. Another\\
trickle. Another hand wipes her skin.\\
Not a bath.\\
Not a bath.

2\\
Her soap dreams whimper,\\
rise in mist, pop in humid air.\\
Droplets dribble bit by bit\\
back onto her bony shoulders.\\
Old now,\\
bones bent with the wet and heavy load.

3\\
Lavender petals lick her skin purple.\\
In her bath, she soaps her wet body.\\
Water breath whispers\\
promises into her cupped ear.

4\\
Water still runs. Soap slips\\
from her hand. From the other side\\
the tap, tap on the door,\\
the voices of her kids, the phone,\\
a heavy thing hits the floor,\\
the dog barks.\\
Part of her life goes on\\
without her. She is not\\
ever alone.

5\\
At 10, she molds her breasts with strawberry bubbles.\\
When such roundness floats to the ceiling,\\
wipes away with a washcloth,\\
drops into bathwater,\\
dissolves into her hand,\\
disappears,\\
she pulls the bathtub plug.


\clearpage
\section{Fluttering Heart}

My heart skips beats\\
while yours pumps too fast.\\
We survive our hearts,\\
ignore days we forget love.\\
Our blood thins to pink\\
after thirty years of marriage

Tonight I reach back\\
past all those years.\\
I pull myself forward\\
and am still what I was then:\\
a prom queen in peach organza,\\
shoulders bear. I show you skin,\\
raise one shoulder,\\
beckon you for one dance,\\
then leave.\\
That blue-eyed crooner\\
caresses piano keys,\\
lures me back to you.\\
We nestle,\\
shuffle across the floor\\
like two lovers hanging on.

My love, the black vinyl is scratched.\\
The needle cuts deep,\\
skips some grooves.\\
That man's voice\\
meanders through heartaches\\
we have embraced all these years.


\clearpage
\section{From My Watch Tower in the National Forest}

I spiral upward, round\\
and round such angled steps.\\
I grow dizzy, live on less\\
oxygen up high where few go,\\
where wax faces burn to bone.

My bare feet blister\\
against sun-heated iron\\
steps, rusted from rain,\\
from too few feet rapping,\\
tapping their metallic tattoo.

One orange-vested hunter below\\
searches for signs: bent twigs,\\
a white-tail flash, pheasants,\\
even a rabbit. Something, anything\\
out of season, worth the risk.

One family pounds in stakes\\
strong enough to moor canvas to ground.\\
They lose themselves on unmarked trails,\\
argue over the way back, wondering if\\
their green tent houses brown bears.

I scan for smoke, for fire\\
during this drought month when I run\\
from air-conditioned life, yearn for heat.\\
I am far above ground, an aerialist who swings\\
from earth to air. I can just let go and burn.


\clearpage
\section{Frozen Fantasies}

\emph{or, The Ice-Cream Lady's Modern Romance}

The ice cream lady cruises slowly\\
ringing the ice cold bell\\
reading a \emph{Modern Romance}\\
closing her eyes against children\\
running alongside screaming

wishing she were a heroine\\
living between glossy pages where\\
men are constant\\
women beautiful\\
and children all grown up.

But the white truck claims her\\
wrapping her in steeled flesh\\
drawing her into frozen depths\\
whispering with frosted breath\\
hinting at the age to come

while she dreams of raspberry swirl orgies\\
where chocolate melts into vanilla cones\\
flows over butterscotch ripples\\
and smothers mint surprises\\
in the heat of the day.

And the children feed on her\\
licking off her rouge\\
combing her darkest curls\\
kissing her earlobes\\
drowning out the bell

until she accelerates,\\
escaping\\
teeth white as picket fences\\
and hands full of gravel roads.


\clearpage
\section{Garden Tour}

Come view our landscaped life:\\
a labyrinth of green hedges, lush,\\
leaves wide as lily pads, cut\\
back against their own growth

daily. Prune with gardener shears\\
the mazes, the complications.

Over there, earthworms\\
aerate soil, rise in rain, plump.\\
Worms dry out on our concrete path. No more\\
than leather laces, fit for fastening\\
boots, grounding souls (yours and mine)\\
against electrical storms. Sudden,

your words hang like dust motes\\
suspended. Air currents crackle\\
around us full of light. Rise and fall

tall as red cedars, deep\\
as willow roots spread\\
underground. Crack reservoirs full\\
with water, pounds of pressure. Push

against walls. Flood lowlands, valleys\\
dry too long. The rush of it all.

Cup hands. Drink\\
from white water rapids\\
with me. Abandon rafts.\\
Touch boulders smooth\\
as skin, wet with moss. See

overhanging branches bend, red\\
with berries, very tart. Taste

curls our tongues: two pink buds\\
flower, unfurl petals

fragrant as spruce trees,\\
wide as sails spread to the wind.


\clearpage
\section{Gemstones}

Anger is her jewelry box,\\
wormwood worn thin, a veneer\\
of voices fill the grooves.

All day she mines gemstones (a word,\\
a look, a thing undone) and hoards them,\\
pins them to her breast, hangs them\\
from her ears, her neck, her wrists.

She strings amethysts. Purple as plums\\
her husband's sighs peel skin, chew\\
pulp while juices puddle on his lips,\\
a slick patina. Their marriage holds\\
the pit, the center shrunk to size.

In this season of her sanity, she\\
wanders, wonders if\\
she lets go, drops down abandon\\
shafts, she'll find\\
herself or at least carbon\\
traces of her in dreams, crystallized,\\
she'll wear in her hair.

Under the hunter's moon,\\
she pirouettes, a circle of light.\\
She dips into her jewelry box,\\
drapes her body in opals.\\
Her ruby rings click\\
like castanets, summon forth\\
old arguments, her dancing partners.

Her legs incandescent, flare\\
into flames, a conflagration\\
of recriminations. She swallows\\
fire, face flush with new desire.

She twirls faster, calls forth winds\\
that whip white coals, strip her\\
naked. Ashes smolder,\\
harden into black onyx\\
under her feet, solid ground.


\clearpage
\section{Genealogical Descent}

Three living ancestors escaped\\
their right-angled boxes\\
and splattered my lineage\\
with third and fourth marriages\\
to fifth cousins of uncles\\
I never called\\
uncle, never

called my father\\
anything but father\\
the same name\\
as the priest on the other side\\
confessed to me\\
and I forgave him\\
his absolute faith\\
without relatives

without blood family like mine\\
who all flaunted gold-plated crosses\\
and glittered in stained light\\
as my Irish Aunt (on my father's side)\\
went straight to Troutsville\\
Virginia from her northern island home

and in that southern state\\
she lost her virginity\\
and her rosary of beads broke\\
under her feet, rolling her right\\
past the ticket taker, past\\
the driver of the Greyhound bus\\
who winked and drove her straight\\
to us in Detroit

where the Ford factory\\
promised her work\\
tightening screws on dashboards\\
as she assembled her life\\
under twin black smokestacks\\
rom which clouds flamed\\
and burped grey ash\\
igniting her stories\\
of paychecks she frittered away\\
on men who swallowed the fire\\
in her red, red hair

and smoked her brand,\\
unfiltered Lucky Strikes,\\
she lit and puffed\\
out in perfect circles\\
I impaled on my ring finger\\
as she and my mother dressed\\
for high mass every Sunday\\
at the Sacred Heart Catholic Church,\\
not missing a white-glove chance\\
of coaxing nylon\\
over their slender legs,\\
checking that the dark seam\\
ran straight up the back

where four covered buttons\\
dangled from elastic hands\\
front and rear.\\
Snap, snap!\\
Snap, snap!\\
End of the line.


\clearpage
\section{The Girl Who Wanted to Be a Heroine}

In the northeast corner of our basement,\\
I huddled under the rough wood\\
of my dad's workbench where he fixed\\
whatever broke. Only I filled space\\
below as wind scrabbled,\\
clawed locked windows, wanting\\
everything inside. Even my small body.

The screech of Detroit's emergency siren\\
split the dark, rainy June afternoon. Even\\
the transistor radio in my hand shone grey\\
as it spouted static-filled tornado warnings\\
and watches when I didn't know what\\
I should look for or what I should do\\
if I spotted a tornado twirling\\
down our street slicing houses in half,\\
picking up, then dropping\\
whatever stood in its way.

The rattle-tattle wind rumbled\\
down the steps, and intruder full of threat.\\
maybe it would swoop down on my mother\\
pulling her grocery cart behind her\\
as she hurried home from Kroger with carrots,\\
hamburger, milk, and oranges\\
we might never eat. Or what about my father\\
at his jewelry store, glass counters\\
jammed with diamonds, emeralds, and precious\\
stones that would never be found if glass\\
cracked or if my father was already gone\\
on his way home where he would find me\\
safe under his workbench, radio sputtering.


\clearpage
\section{Giverny in April}

\emph{for Claude Monet}

blossoms with slate roofs,\\
pink and green houses----and life\\
plump as rhododendrons. Fuchsia

petals cluster into constellations,\\
a whole galaxy you can clutch. You\\
clean fine-bristled brushes.

You embrace lily pads stroked\\
by weeping willows kneeling at pond's edge\\
before they shed leaves. You crave truth

the reality of transient light, breezes\\
and pigment. You dig in dirt, paint\\
sunflowers-\/-\/-reflect life from the ground up.

Your paintings hang now\\
in museums, noisy with tourists.\\
You can be found

on postcards, in guidebooks\\
jammed in canvas bags. People want\\
a piece of you to carry home.


\clearpage
\section{Giving Directions}

``I was a student once,\\
still am, in fact,\\
a student of students,\\
ready to learn\\
from you about me\\
ready to wear\\
your faces if I must\\
when I read your poems\\
late at night, alone and tired\\
of lies and words you think\\
I want to hear about\\
`nice' friends and fun.\\
Tell me instead\\
about Uncle Ed who\\
drinks too much and dances\\
late into the night while\\
Aunt Martha sleeps, hiding\\
from words\\
better left in the book\\
on the bedside table.

``We'll awake\\
together and clutch\\
at words\ldots''

``How many,\\
words, I mean, for\\
tonight's poem?''


\clearpage
\section{Glaucous}

Gaze long enough at me\\
you see\\
yourself. You\\
lead me to blue\\
where pieces of sky\\
plummet into Stoney Creek.

Your arms are bare. Blue rivulets\\
flow under your skin, warm currents.\\
You dip my finger into blue\\
and I touch liquid green.\\
We reach for white, those strands\\
clouds leave behind\\
when they rise moist.

Look there. That heart-rock\\
plump with age-old mystery.\\
All those river rocks in bed\\
lulled by dragonfly buzz.\\
We stroke lichen sun drops\\
that line of green-yellow fur\\
on blue rock road.

Come let us soak\\
in wandering waterways,\\
our toes gripping river rock\\
as all that wet murmurs\\
the ancient blue-green song.\\
Who we are.\\
Who we will be.


\clearpage
\section{Grandfather Never Gambled}

My grandfather misplaced himself\\
one cold Ohio morning when he cracked\\
one hundred years. With double zeros\\
like unbroken egg shells, he cheated\\
those small town gamblers,\\
his farmer friends who never believed\\
so many years could accumulate while\\
they hoed and harvested crops,\\
patched their own roofs and gathered\\
eggs for breakfast, met at the barbershop\\
and never said much beyond\\
what they did that day when grandfather\\
never showed at the collection celebration,\\
even though he must have known\\
they all had the wages they owed\\
stuck in their plaid shirt pockets.

When these men found my grandfather\\
alive and upside-down, his ankle caught\\
between branches of the green apple tree,\\
I figured he was whistling\\
as he watched the purple waves of sun\\
drop behind the ridge, waited for\\
the start of his next century, knowing\\
his friends would come on foot\\
whooping and hollering, calling him\\
home from the field, home from the tree.


\clearpage
\section{The Great Backyard Bird Count}

The green webbing breaks\\
and hangs down one side\\
of my aluminum lawn chair\\
and I worry\\
I might break\\
in that same unexpected way,\\
maybe even today,\\
this February day of the Great\\
Backyard Bird Count,\\
this frozen day everyone will recall\\
as snow blots out the sky.

The town closes\\
and everyone stays home,\\
even me who huddles in my back yard\\
in my dilapidated lawn chair\\
just waiting to jot a check\\
mark in the box next to the robin,\\
the hope of spring clutched between its beak,\\
its promise of damp soil and fertility.

I am prepared with my Field Guide\\
of Birds, binoculars, thawed birdbath,\\
a thermos of coffee, two sandwiches,\\
some cookies, blankets, my check sheet,\\
and one pile of small rocks to chase squirrels\\
from my own backyard. I yearn for birds.

Mourning doves coo but I cannot\\
spot them. The starlings, crows, and grackles:\\
all those black birds blend into dark branches.\\
I cannot see even one feather.\\
I cannot see my boots\\
covered with snow below me.\\
I am afraid the birds have eaten all the birdseed\\
when I wasn't looking,\\
have flown off on their journey\\
and left me here alone.


\clearpage
\section{The Guided Tour}

{\bf I.  At Pioneer Square, Seattle}

A ticket gets us in\\
and we descend, step by wooden step,\\
into the subterranean city\\
below Doc Maynard's Pub. Above in the Square

tourists perch on our heads,\\
flap their arms and shake loose dirt\\
that dribbles onto our bare shoulders,\\
logs the guide's upturned mouth\\
with fire dust and a city buried\\
in its own ashes all the way up\\
to first floor windows\\
through which we view

more dirt. Enough dirt\\
to smother us in open catacombs\\
beneath our crepe-soled shoes.

{\bf II.  Tourists Once Removed Automatically}

We load our auto-focus cameras\\
with high-speed film\\
in one second flat and snap up, down,\\
around, and close up if we dare. We do it\\
from far away using telephoto lens\\
unscarred and clean of lint

for the clearest shot once removed\\
and framed in the viewfinder.

Our flash backfills unnatural light\\
of electric glass as we sight\\
our guide and shoot.

{\bf III. Photographing the Blind Woman's Music}

No one resurrects\\
or restores or ignores the blind woman\\
whose grey hair frizzes in this humidity\\
into a halo of split ends\\
above her sleeveless ivory blouse\\
embroidered with purple violets and emerald leaves,\\
each tipped with tiny mirrors reflecting back

our faces as she tucks one lemon-yellow cosmos\\
between her breasts, its stem wilting\\
against her sweaty flesh. Then she empties coins\\
from a cup at her feet into her palm. Silence now\\
while she rubs each silver and copper circle\\
between her fingertips as if they speak to her\\
of those who flip them in her plastic cup\\
as they pass by

when she plays the dulcimer hanging round her neck.\\
She does not see us lick our raspberry yogurt cones,\\
tongues chilled by the heat\\
of her dulcimer's metal strings.

That woman strikes with two hammers\\
one in each hand and a mouth full of hymns.

Words round as God's name puff her cheeks\\
into crimson cushions that curve toward bone\\
when she plucks one string and ends.

Click, the camera catches her then,\\
the last shot at the end of the roll.

We know the finish, that whirring noise,\\
that internal winding cellulose\\
from one black spool to the other.

{\bf IV.  In the Darkroom at Home}

Under the developer,\\
the blind woman evolves again\\
into herself and brings back our sight

so we remember the melody of her coins\\
clinking in her skirt pocket\\
when she lifts her arms, just before\\
she hits the strings in the right spot\\
and shapes echoes from another place\\
that settle deep in our bellies\\
and we hear our own eyes blink in light.

{\bf V.  Squares of Time in Photo Albums Everywhere}

We carry away\\
slices of places and people\\
in Kodak color\\
and paste them in photo albums\\
or shuffle them like decks of cards\\
as we show friends, family, anyone\\
who will look at where we've been.

In other photos\\
a blind woman who may be her stands\\
behind us in Times Square playing the flute\\
ten years ago. Even then,\\
our cameras dangle from our necks\\
like black pearl amulets.

We spot the blind woman\\
staring out at us\\
from antique albums\\
on dusty bottom shelves\\
at flea markets\\
in other towns\\
as we bargain for her image:

her eyes flat as the world map\\
and full of buried cities\\
that are silent as the pulse\\
beneath our skin.




\clearpage
\section{Heart Murmur}

Mama crashed into her 90\textsuperscript{th} year\\
full speed and braked\\
for no one except me\\
who knew

her heart thumped erratic beats\\
and murmured, a whirring sound\\
as tapes rewound decades\\
past these bedsheets\\
trimmed in violet and pink eyelet

past her third husband she left behind\\
in Venice on an air-conditioned tour bus\\
eyes closed as Mama's second husband\\
who drank Bourbon straight, a southern man

from Lexington who bet on thoroughbreds\\
called Mama Too and lost himself\\
as Mama dressed in her white satin\\
gown with beaded bodice zipped

against her grief for one husband\\
who came first and slipped underground\\
beneath a bluegrass sheet one August day\\
when his heart leaked blood

drops red as cardinal feathers\\
shed along the path, markers of the way\\
for me when I was born in blood

and Mama saw him\\
in me and rocked us\\
in her arms, back to life again

and again she sang the cardinal's song\\
all along knowing what women know

how warmer than normal, this night would fall\\
through the open window, into her room

while we two women\\
listened for father's feet\\
and heard red\\
red wings tipped with black.


\clearpage
\section{Her Father Spied What Was Not There}

Her leather diary:\\
skin like a ripe red delicious\\
apple, freshly picked from the tree\\
of lies and wishes and what might be.

Her father:\\
twisted a key in the tarnished lock.\\
He might as well have stuck\\
an ice pick in her back.

Secrets hidden:\\
blank pages crammed\\
with blue lines\\
and empty space unbroken.

Only her heart:\\
cradled all she remembered\\
of life until now, the horrible\\
longing for something else.

The rest she wore\\
ironed on her sleeve.


\clearpage
\section{Her Night Mares}

Her swollen tongue\\
fills her mouth\\
with mushy screams.\\
Smothered words\\
in thickening saliva;\\
a marsh\\
of muddy waters,\\
decayed leaves,\\
rotted bodies.\\
No one to hear her\\
as she drowns

in white-cold papercuts,\\
her scissored dreams\\
press against\\
the schoolhouse windows\\
cover unwashed streaks\\
as her child fingers\\
scrape the winter glass.

Her frosted breath\\
warm against the panes,\\
melts her frozen mind

until behind her eyes,\\
she finds\\
a poem\\
she can mount bareback,\\
stroke with breathless whispers,\\
spur into a gallop\\
across the desert sand\\
where blanched bones\\
nestle at night,\\
where she laughs,\\
races toward mountains\\
until the mare's sweat is hers\\
and she slips\\
into lined white emptiness.


\clearpage
\section{The Hypnotic Mrs. Little}

The flap of skin under Mrs. Little's chin\\
waggled as she warned our third grade class\\
about bombs and flash burns during\\
our monthly air-raid drills while we stood behind\\
cement pillars in the school basement with others\\
whose legs trembled and eyes scrunched as if\\
they might cry or sink to the ground from the weight\\
of the whole school on top of them.

Back in class,\\
when I messed up the multiplication tables\\
and erased numbers so hard the paper ripped,\\
Mrs. Little rapped my head\\
hard with her sharp fingernail, polished\\
blood red. Just my whisper at her back\\
whipped her around. She aimed\\
the eraser at my head. White chalk\\
fluffed the air, settled in my hair.

Mrs. Little drilled us in disasters:\\
how fast Newton's apple fell from the tree\\
of knowledge, how the sky might freeze\\
and break into silver hatchets\\
that could split a head in half.\\
A crow could plunge through air,\\
snip off a bit of our ears. She could\\
hold us chickens back a year.

I pinned the sky in place with my eyes,\\
opened my arms wide to whatever might fall.


\clearpage
\section{A History Lesson}

Mark Faneuil Hall: tell its story on bronze.\\
Let tourists stare and snap their instamatics.\\
Inscribe the Declaration, etch it on the wind\\
whipping around the corner of the Old State House.\\
Never mind Prudential Tower; its foundation isn't set.\\
Guide them on the Freedom Trail past the Park Street Church\\
Erect wooden signs to hold the spots,\\
scrawl on paper if you must.\\
Sell them guidebooks, maps, and souvnirs.

Then take their faces, wet with sweat,\\
and turn them toward your own.\\
Field stones build the base,\\
wall off space for chin and cheek,\\
sink in eyes at high tide.\\
Uprooted sugar maples ridge the chin,\\
pour blood-warmed syrup over ice.\\
Lips like cod swim the face,\\
close before the hook.\\
Mourning doves coo in your throat,\\
remembering the first winter.\\
The covered bridge between your eyes\\
uses king posts like an arch.\\
You breathe the blacksmith's fires,\\
melt iron for your skin.\\
Two northern red-bellied snakes for eyebrows\\
underline forehead furrows plowed\\
by Puritans, chanting scriptures\\
after the horses were fed.\\
And in your eyes are alpine gardens,\\
clinging to the rock.


\clearpage
\section{Hit By a Hearse}

Just because I didn't look\\
both ways, right\\
and left, from air to ground,\\
I didn't see me or anything else\\
reflected in the chrome grille\\
burnished that winter morning in Detroit\\
behind the closed double garage doors\\
of J.W. Morely's Funeral Home

perched for forty years\\
on the corner of my street\\
like some black crow with marble eyes\\
who chants: \emph{caw, caw, caw, caw,\\
}and waits for something small\\
to move across its sight line.

Just ice patches, I know\\
but a girl like me can slip\\
if my boots strike the wrong angle\\
and puncture the thin layer\\
of trust that freezes\\
and holds my full weight, all the pounds\\
I put on since marriage, since the children\\
moved out, since

my husband retired early. He watches\\
the TV weather woman predict sun\\
somewhere in Florida\\
and tips his 90-proof vodka bottle\\
up toward heaven in absolution,

his hair once brown as the paper bag\\
that will hold my apples, brown\\
as fresh-turned dirt. Now gray\\
and slicked back, its strands thick\\
with lies are wound round his head:\\
his heavy crown of thorns.

I cross the street\\
alone on my way\\
to Kroger's for tart, green\\
Granny Smith apples\\
plucked from the trees in Washington\\
and transported in crates\\
to this city without orchards\\
as if a taste of truth, a trickle\\
of juice down my chin from the first bite\\
that breaks the skin,\\
can make me see the black,\\
sleek bullet of the hearse.


\clearpage
\section{Home Again}

I poke in corners of the world\\
and return naked to my clean,\\
uncluttered space.

I stand still\\
as the lily pond outside my window,\\
my thoughts in bud, about to burst.

My fingertips curl toward a palm, a cup\\
of air. I breathe in natural light,\\
exhale shadows, my flesh-tone poem.


\clearpage
\section{Home from the Sea}

I unpinned a story of the North Sea\\
from the braided bun my Middlesbrough\\
grandmother wore. The silver strands\\
unraveled, flashed waves of whitecaps,\\
a foam of bubbles, a school of gray-green cod,\\
a glitter of fish scales, a wealth of shine\\
polished in the port city where ships docked.

On the day she almost died\\
icy water lapped her ankles, she said,\\
as the fork-tailed terns spread white-cloud\\
wings above her. The current dragged her\\
toward Denmark, maybe Germany, some place\\
far from the English beach, the sand, the shore.\\
Deep swirls of wet raked her back,\\
filled her mouth. Splashes of red beaks\\
pecked herring from the green-blue\\
wet beneath her fingertips. She lost\\
direction. Brown-speckled boulders broke water\\
and spoke the secrets tangled in kelp beds.\\
The seals' sharp barks herded her home.


\clearpage
\section{The Hook}

punctuates my life\\
with its head bowed\\
and spine stretched\\
toward a single dot

a speck at the end\\
when the hook\\
scoops white space\\
embraces emptiness

while the question wiggles in my voice\\
like a nightcrawler\\
impaled on barbed wire\\
and dangled from a line\\
whipped out\\
by the booted man\\
who waits on sand

and does not ask himself\\
or me if he can\\
scale and gut a trout\\
and eat what\\
once\\
swam free

as his love for me\\
without question.


\clearpage
\section{Hospice}

\emph{For the woman with Alzheimer's}

She shrinks daily, a sliver\\
of herself peeled to the core\\
of apple seeds, moist\\
with juices. Tipsy and cider-drunk,\\
she steals out windows, out doors, out\\
any open crack, a contortionist\\
bending low under the limbo stick.

Her body a drum, skin stretched tight,\\
reverberates with memories, too loud.\\
Hammers pound, play her bones\\
like a xylophone, up and down the scale:\\
the mating dance. Back and forth,\\
her legs straddle the stick, this side

then that. Silence, sudden as white\\
elephants. Smooth as ivory tusks.


\clearpage
\section{House of Good Fortune}

The pink-haired teen drags a pile of troubles\\
in a gunny sack. Everywhere she goes,\\
people stuff their problems into her burlap bag\\
as if she can ease their weary days.\\
She shovels the whole mess into a compost pit\\
where it mixes with the bright heads of marigolds,\\
muddles with rot and potato peelings\\
and curls around strangers' problems.\\
Everything breaks down eventually\\
and oozes over the boundaries.

The teen dyes her hair brown, the color\\
of fertilizer. In one coat pocket\\
she strokes her worry stone,\\
sighs as the shush of flesh rubs\\
over the rose quartz and settles her mind.\\
Behind each ear, she tucks a four-leaf clover,\\
protection against the evil eye\\
that could shrivel her to nothing.

An amulet dangles from her neck, given to her\\
by a cousin who stitched powerful magic into each seam.\\
The belly of this deerskin pouch cuddles one creek\\
pebble, one tail feather from the common sparrow,\\
and one dried rosebud plucked during the heat of the day.\\
She wears her silver charm bracelet so she can fondle\\
the book, the piano, the wishbone, the tiny scissors\\
for cutting threads, the Scottie dog that almost barked\\
and wagged its tail and the heart engraved\\
with her name and birth date given to her by her\\
mother who pinned all her hopes on this baby girl,\\
now a teen who believes in luck, forgiveness\\
and second chances. She gives up worry\\
when she enters the house of good fortune.


\clearpage
\section{How to Find Happiness}

Reach back behind crunchy granola bars. Grab\\
the wisp of smoke. A gasp of hope curls in sleep.\\
Shake it awake. Make it talk sounds of cotton-candy pink.

Listen as a tiny gold-haired girl giggles\\
up and down the scale. Improvise\\
the jazz of peppermint, the razzmatazz of chocolate drops.

Slip-slide twelve steps forward. Reach out for Neapolitan ice\\
cream. No, go for rocky mountain road, the crunch,\\
the pecans cracked between vanilla teeth. The pleasure of it all.

Forget the flour and milk. Add very berry ice cream\\
and red licorice, of course. Savor the tantalizing tickle\\
when happiness hits the roof of your mouth.


\clearpage
\section{I Do Not Write Poetry}

it writes me\\
into the blue-black center\\
of my birth back then\\
when I slid head first\\
into sterile white with no words

for my life pushed into that mid-afternoon\\
glare of Detroit time clocked in and out\\
at the Ford Body and Assembly Plant\\
and ticked off by the White Castle\\
belly-buster burgers slammed one after the other\\
onto the greasy grill and patted flat by the slender cook\\
who knew her blank-verse days ended Sundays\\
in the Temple Baptist Church on Woodward,\\
the main drag for the '43 Ford V8 DeLuxe coupes\\
revving and running lights too red\\
after the world war I read about in poems\\
without rhyme

and later, words\\
slapped me flat as a White Castle\\
when poetry sizzled blue in my mouth\\
dribbled onto pages of my life\\
and wrote me into a simile\\
as if I could puzzle out\\
my birth and death rites\\
and scrawl poems in between.


\clearpage
\section{I Got the Dirty-Devil Blues}

I got the no-place-for-me blues,\\
the lay-down-and-worry blues, way down\\
since the Chevy plant dumped me on the street.\\
I got no laces for my shoes.\\
I got bunions on my feet.\\
My unemployment done run out.\\
I'm down so far I can't get up.

My welding gun has lost its flame\\
and everything needs soldered.\\
No way to melt the metal.\\
No way to join the seams.\\
Things just seem to fall apart.\\
The cold wind blows thru cracks.\\
This old house has broken glass.

No one's calling me for jobs.\\
Everyone forgets my name.\\
The phone don't ring.\\
I don't ever pick it up.\\
The front-door buzzer broke yesterday.\\
Everyone stays away, far away.\\
No one wants to catch these down-and-out blues.

Maybe tomorrow I'll get dressed in my best,\\
make the rounds of businesses,\\
put in my applications for work, any work.\\
I'll try McDonald's with golden arches.\\
Minimum wage means it can't go no lower.\\
The only way is up or out---and I ain't out yet.

For now, I just got the dirty-devil blues.\\
That old devil sucker-punched me good.\\
That old devil knocked the wind right out of me.\\
That old devil whispered hurting things in my ear.\\
That old devil sucked my soul.\\
For now I'm just laying back, eyes almost closed\\
waiting for that awful sound of the final knock-out bell.


\clearpage
\section{I Have Become}

a chair, slipcovered\\
in neutral beige, bound

by the decorator's taste.\\
His fingers shape my face,

tack the fabric tight.\\
He levels legs and leans

against my back, hoping for\\
an empty space, a place undone

like stepping off a cliff.


\clearpage
\section{Impromptu Essays}

You write to me on desktops.\\
Jagged passions etched in wood\\
with steel paper clips and dry pens:\\
Lopsided hearts pierced with arrows.\\
Mary \& John in '79 and class of\ldots{}\\
Advice to one before\\
from one who sits now\\
knees gum-glued to metal,\\
his motionless daydreams\\
(heavy against black oak)\\
explode in magic marker\\
expletives,\\
the permanent kind\\
as I pass out the mimeo sheets---\\
no peanut butter smudges\\
or erasures---\\
to cover the splintered hardness\\
of your fears.

The assignment\\
created in four nights\\
in my draped den\\
calls to your mushroomed souls\\
to write their spongy truths\\
in measured lines. How\\
can I tell you\\
my red pen's out of ink,\\
my text is coffee-soaked,\\
the grammar rules lost\\
in red maple roots?

You sweat spring rain,\\
build wadded paper glaciers\\
between us,\\
write pencil essays\\
in dictionary words.\\
Hand them in.

Now, carve your woodcuts\\
until bells ring\\
and I'll read\\
your impromptu essays\\
after class\\
when I sit\\
where you sat\\
and touch\\
your word slivers.


\clearpage
\section{In a Double Bed}

She dreams\\
oceans beneath her bed.\\
Eels slither upward\\
electrify the sheets;\\
the octopus extends its arms\\
wraps her\\
and drags her down\\
to salty depths;\\
she breathes\\
within the embryonic sac.

He dreams\\
mountains rising\\
peaks tipped white;\\
snow melts between his lips\\
cools his belly's hunger\\
floods the dams below\\
until people look up\\
see him plant the flag\\
and breathe\\
the frozen air.


\clearpage
\section{In Anticipation of a Poem}

My inspiration left me yesterday\\
for a long, cold sleep under ground.\\
It curled its tail around itself\\
and ceased its chatter. I was alone

above ground, afraid death would wander\\
into my house like a black fog\\
carried on bat wings. I chewed silence\\
on my back teeth, swallowed all this

nothingness. Without words,\\
I, too, hibernated inside my burrow.\\
Blank walls boxed me in. Only\\
my cache of nuts promised sustenance

as I cracked hard shells,\\
dug out the sweet meat of truth.\\
I waited for the melting time\\
when soil loosened its hold on living things.

The rhythm of the seasons\\
shook me from my lethargy. Blood\\
rippled in my veins where words bobbed\\
above river stones. I created poems of air

that rode the wind currents as if mere images\\
nourished this brown chirping creature\\
whose paws curled on my shoulder\\
as I grabbed a handful of bubbles rising.


\clearpage
\section{In Her Own Image}

Smaller than she remembers\\
her childhood house,\\
no more\\
than a green bungalow set up\\
on a Monopoly board\\
collecting rent.

Hands roll dice,\\
come up four: father,\\
mother, sister, her.\\
Then three, unlucky\\
number after mother dies\\
an ugly death.

Father\\
in less than a breath\\
displaces empty space.\\
He takes his photographs,\\
portraits black and white,\\
hand painted\\
by his assistant,\\
his new wife\\
and he vanishes.

Snapshot of him:\\
face full of blood, flush\\
with a winning hand.\\
Let her, the youngest\\
girl, shuffle hearts\\
into diamonds, snatch clubs\\
from behind his ear.\\
Master of magic\\
he saws women in half.\\
His whiskey words,\\
caustic as an acid bath,\\
coax images\\
into photographs.\\
Roll flat and frame.

She's three, all braids\\
and bows. Her knees,\\
bare beneath pleated skirts,\\
bend like pipestem cleaners\\
as she poses for him.\\
Chin tilts up, eyes down,\\
mouth a wedding wreath.\\
Shutters click, capture light\\
and dark. Shadows of her soul\\
reverse on negatives.

Unblinking, his third eye\\
zooms in, yanks her\\
into the black box,\\
turns her upside down,\\
arms dangling like linguini\\
hanging off her plate.

Her teen years\\
glisten with developing\\
solution, wet and slick as sweat.\\
Sour as citrus fruit, he\\
smells of screwdrivers\\
and peppermint. In his darkroom,\\
he airbrushes blemishes,\\
shapes her mouth\\
into her mother's smile.

During her forties, men\\
meander through her\\
living room like tourists\\
at art museums, looking\\
for paintings of places\\
they've been\\
and left behind or nudes\\
whose hips curve\\
into thighs, muscles\\
tense and tight.

Like her father,\\
these men reek of drink\\
and blackouts, punch huge holes\\
in her book of days\\
as she forgets to remember\\
the whiskey chronicles,\\
that wet, cold current of lies.


\clearpage
\section{In Our Blackberry Days}

I roll those August farm days in my mouth\\
like blackberries, thick, sweet on my tongue,\\
round as that summer memory of us.

Wild at ten, mosquito-bit and bruised,\\
all legs running, arms reaching for more\\
than my cousin or I bargained for that week.

She taught me how to dribble\\
chicken feed through my spread fingers, call\\
home the jersey cows and lock the gate.

In the sun-soaked days, our skin burned,\\
blistered in Ohio heat. Back in the woods,\\
we found a wild bramble bush, heavy with fruit.

With broken hoe handles, we whacked that thicket\\
until a black rat snake slid past our ankles\\
and even the bees fled on whirring wings.

My cousin jabbed her finger at the most juicy blackberries\\
hidden under sharp-toothed leaves, huddled in the tangle\\
of arched stems, the place where thorns hooked skin.

With curved wrists, we gently tapped the branches\\
until blackberries drifted down into our open mouths\\
and juice drenched our lips, our tongues, dribbled down our chins.

We rescued blackberries from a purple finch, throat\\
puffed with song and a dark-eyed Junco in mid-snatch.

A waterfall of berries spilled into our silver buckets.

At night, my aunt stirred big pots of blackberries\\
while we prepared wide-mouthed jars\\
greedy for the dark nectar that stained our fingertips.

In my parents' car, I clutched three jars of purple preserves\\
cooling in my hands, hardening into something precious\\
I would spread on biscuits in a white bungalow in Detroit.

In the August of my retirement, blackberries\\
pop up in the pink of my wild roses. I know\\
the purple finch carried my cousin's seed here in its beak.


\clearpage
\section{In Search of a Healing Place}

For her health, Mother rowed a boat\\
across Pleasant Lake, then caught\\
a train headed toward Black Springs,\\
Michigan, all by herself,\\
searching for a night-\\
colored spring with the tar\\
smell of a rotted house.

Huddled at the very tip\\
of the rabbit's ear, Black\\
Springs was not on the route.\\
Mother hiked the final mile\\
full of pines and endangered\\
trillium white in moonlight.\\
Wilderness grew in her mind.

Mother wandered on red\\
maple leaves farther along the trail\\
and slipped right through their covering\\
into black spring water some people say\\
doesn't exist. If it does,\\
they say, no one survives the extremes:\\
the bubbling fire or the frozen floor.


\clearpage
\section{Interchange}

You come back to what you leave behind.\\
Like me.\\
Like litter on the roadside: Kentucky\\
Fried Chicken cartons. Bones spill\\
from red cardboard, bleach white in summer sun\\
or cover themselves with bluegrass

and grow toward you. You\\
roll down windows, breathe in\\
skeletons rising up, new life\\
alongside Interstate 52. You and me

dust to dust. The concrete median\\
crumbles moments into hours. Plaster\\
particles and sharp pieces of stone\\
puncture clouds above the moon roof\\
of your '99 Windstar. Metallic seeds

float there. Settle into cumulous\\
formations flat on the bottom.\\
Fuse with fish bones. Fragile\\
as your flight you chart with Magic\\
Markers on the map from AAA. You accelerate

past safe limits, past lavender\\
twilight on your right. You look\\
left. Vacancy signs flash red ahead.

Iridescent neon tubes twist words\\
we said and unsaid. Our silence\\
cracks an open space between

us, unconnected cartilage. Bones\\
bring you back.

You press your breastbone\\
against mine and both soak\\
in the salty brine of birth,\\
our fused vision of bones.


\clearpage
\section{In Woodlawn Cemetery After Dark}

Who's to say what darkness is\\
or when it falls to ground, \emph{kerplunk}?

Not I who usually visit during lunch\\
when I'm hungry for life and company.

I memorize names of those who will be\\
by me when my number is called for residency.

I research their names, read their obituaries,\\
introduce myself by telling my life story.

Kenneth, I scold, you should not have driven\\
yourself around that curve smack into a tree.

But I understand how you fell asleep worrying\\
about the economy and your boss who screeched.

You are a man I can talk to honestly and trust\\
to take no liberties with an old woman like me.

Helen, we will lay head to head. I read news\\
accounts of how at eight a stranger strangled you.

You were just beginning then and I am fading from\\
this world to yours. We will talk of girlish things.

So many names, so many years, so many causes\\
have buried you here. I have my plot, paid in full.

Sun lulls me on this fine fall day when leaves drop\\
geese fly north and I fall asleep above my place.

I have read closing hours on the gate, amazed\\
that we who breathe have a curfew, a deadline to meet.

I wake to dark, a slice of moon and stars for light.\\
These violated hours lock me behind fences with you.

Come out, I plead, and chat with me, your company.\\
Mist shrouds headstones, drapes trees and bouquets.

Do not hide from me for I cannot play that childhood\\
game. I fear tripping on such hilly ground.

Is that thick haze the shape of you? Or that shrub\\
huddled over there? Your secrets will soon be mine.

Must I wait? I can recite your names, a litany\\
of life I offer you. I am trapped in a forbidden place.


\clearpage
\section{It's My World}

I have eaten off the blue plate,\\
swallowed each grain of rice,\\
crossed my chopsticks,\\
signify that I am done,\\
my stomach full\\
of undigested food,\\
of questions that roll and growl.

I have removed my slippers,\\
a sign of respect, of openness\\
as I stand on the edge\\
of the spinning red vortex\\
where fire can consume me\\
before I even know\\
who I am or who I could be.

I have allowed my cat curling space\\
on the rounded surface of my head\\
as if he were my hat dangling black tail\\
one side, the other side a white paw of claws\\
holding me in place, safe against red\\
tongues that rake my back, purr words\\
too slurred for me to understand.

I have written notes to myself,\\
my book of changes with pages smudged,\\
lined up like squares of doubt above me\\
ready to flutter down, smother me\\
if I don't reach past the green-blue\\
perimeter where turtle and snail carry their houses,\\
where blue butterfly and ladybug color my sky.

Below me sits the city where skyscrapers sway\\
in wind and bend their concrete sides toward earth\\
where I yearn for yellow day and dark night,\\
for their unity as one cycles to the other like me\\
who stands poised for the dive onto my gold couch below,\\
a velvet landing into my world where I am me.


\clearpage
\section{The Italian Restaurant on the Corner}

Some neighbors say\\
the owner torched his restaurant\\
one bitter December night when\\
his till did not balance\\
and the weatherman predicted\\
a long, hard cold spell.

That night fell below zero as I recall\\
and the stars hung in the sky like icicles.\\
I heard the sirens, saw the red flash\\
of fire engines from my bedroom window.\\
Flames licked the moon with many tongues\\
while water failed to drown the fire's thirst.

When the moon fell from the sky,\\
even firefighters dodged the sparks\\
as the restaurant roof caved in and cans\\
of tomato sauce exploded. Tin shrapnel\\
struck the owner who stood in the ruins,\\
arms raised over his head, cursing.

For days the restaurant kept shedding\\
pieces of itself. I found a fork, a shingle,\\
a charred board and a restroom sign\\
in the rubble coated with ice,\\
a place off-limits where neighbors walk\\
their dogs searching for evidence.

These neighbors desire a crime to solve\\
while I embrace the mystery of ice.


\clearpage
\section{I Wish for Bradley}

\emph{With Much Love from Grandma Carol\\
February 3, 2005}

I wish you white-feathered wings\\
that slice the air and carry you\\
safely past the first star\\
you see tonight,\\
over the giant redwood\\
where you will land\\
and listen to the rustle\\
of green-needled truth.

I wish you a pinch of pepper\\
for zest and a hearty appetite\\
for adventure as you explore\\
all life's nooks and crannies\\
and value what you find there.

I wish you forget-me-nots\\
that color your garden\\
with the blue from the sky\\
and the green sprouts of growth\\
as friends gather round\\
and walk with you through life.

I wish you eyes that shine\\
like the sun, illuminating all\\
you see with hope of what can be\\
and what you can do\\
to make the world a better place\\
for everyone.

I wish you books and music\\
you can share, walks in the park\\
off the worn dirt path\\
where ideas scamper about\\
and feed your curious mind.

I wish you health and happiness,\\
a love of life to see you through\\
and most of all, I wish you dreams\\
and a heart full of love.



\clearpage
\section{The Japanese painter}

inks their dusky souls\\
on rice paper:\\
fine lines thicken\\
lengthen into bamboo\\
lips sucking\\
persimmons;\\
juices drip\\
into body hollows\\
where twin plums\\
eclipse the moon\\
and the painter switches\\
to soft-bristled brushes.


\clearpage
\section{Kneeling in Front of\\ the Rain-Battered Mums}

He leaves the mums alone,\\
not tied against stakes\\
with garden twine\\
or covered with burlap\\
when frost crystallizes night.

He hoes his dreams, chops chickweed\\
rooted in the ground,\\
eats seeds like sparrows\\
nested in the downspout. Gone now,\\
chased by November rain.\\
He stoops, alone, crouching\\
wishing he pinched the early buds,\\
folded in green jackets.\\
These mums, too tall, too leggy\\
to not fall, droop into dirt.

He props them up,\\
weaves stems through cyclone fences,\\
cuts dead heads, white petals now rust.\\
His knees stained soil-dark and his hands,\\
moist, move among the broken stalks,\\
and his voice, hoarse with age, chats\\
a promise, a prayer as he kneels there.


\clearpage
\section{The Language of Waves}

The moon-gold sun balances against my hairline, rolls to my nose\\
wet with sweat,\\
then back again\\
as I steer my bomb of words through these rough waters, under\\
this neon sun.\\
Such unbearable yellow heat\\
that explodes green, blue, white waves of words, churns a froth\\
of sentences,\\
long strands to meaning I read from the banners towed by fighter\\
planes\\
over Whisper lake\\
where bullet words pierce my bare belly until I beg for mercy\\
and waves run red

with blood that will not clot, will not settle into war stories that\\
turn

to stones,\\
sink to the bottom\\
of the lake where they wear smooth, become no more than sedi-\\
ment,\\
particles left from the ice and the fire, from dinosaur trails\\
and fish scales,\\
bits of historical ruins,\\
tales told in hieroglyphics, liquid as the waves that wash out,\\
wash in,\\
wash over me with ancient alphabets I scribble in blue water\\
and decipher\\
the drift of night currents.


\clearpage
\section{Last Day of Class}

On the last day,\\
we come together\\
to record our history.\\
I start it out, a story of us.\\
Alice tells of \emph{The Scarlet Letter}\\
and Ed who read CliffsNotes instead.\\
Mary remembers the day\\
the sub laughed at Paula's puns\\
and the whole class got so loud\\
the principal came in.\\
Ted recalls the test\\
that everyone flunked\\
when I used the wrong answer key.\\
Ellen screeches chalk and pops her gum.\\
Barbara's poem is in \emph{The Clarion}.\\
``The Most Dangerous Game'' is\\
cheating on tests and walking in late.\\
Elmer got locked out and Mike got As.\\
They didn't forget\\
the apple they brought all waxed and shined,\\
the fire drill when no one came back,\\
stomachs that growled at \emph{The Iliad},\\
notes they passed when things got dull,\\
things I said when they pushed me too far.\\
We got it all down and played it back.\\
Voices distorted or so they thought\\
but I knew every one.


\clearpage
\section{The Last Night of the Year}

Even my dog let loose howls\\
my neighbors pretended they didn't hear\\
tonight. Too many one-note party horns blasted the air.\\
Too many feet slunk up the stairs and around the corner\\
and tap-tapped into the bedroom on the second floor\\
where men snatched winter coats\\
worn by those women who slipped out the back door\\
and later snuck back to tip whiskey down their throats\\
thrilled at the risks they dared before the last\\
tick of midnight faced them forward.

Maybe twenty neighbors crammed together\\
under blue crepe paper twists in the basement\\
where the whole neighborhood could fit\\
if they didn't settle for too long\\
under aluminum stars dangled from silver threads.\\
Arms reached out helter-skelter, fingertips slid along\\
cheek bones, lips pressed lips and feet slip-slided\\
on the tile floor. Wiggly bodies dripped sweat.\\
People yelled for their spouses, aroused\\
when the end moved nearer and a last kiss\\
became the first, hovering\\
between lost and found.

One man scurried home\\
alone before the final, fatal tick.\\
The changing of the years,\\
the tick tock of clocks\\
echoed in his head.\\
He stroked his gun\\
three times, tilted the barrel toward the sky.\\
When he squeezed the trigger,\\
the bullet split the air above his head.\\
He did not see\\
where it lodged or if\\
blood flowed in the new year.


\clearpage
\section{Legacy}

Bones blanch white in desert sun.\\
Skulls testify, mourn flesh and rivers\\
of blood that stain rock sunset red.

The sky is on fire, a desolate orange.\\
Only one Saguaro Cactus remains, arms raised\\
in mock surrender, roots ready to detonate.

A bomb blasts open earth's black mouth\\
spewing boulders from the underground.\\
Green tanks roll on caterpillar treads.

Still, the gods wager against all odds,\\
shake two dice and throw. No sacrifice\\
appeases blood thirst if they throw seven.

Such gods play high stake games\\
through centuries. They rake in rubble,\\
watch which way twin towers crumble.

A soldier guards the death cart\\
while the brown moth guides spirits\\
underground and the serpent strikes.


\clearpage
\section{Lemon People}

Maria resides in a box, in a place\\
she calls home with four walls\\
and no address. She plasters\\
photos everywhere, buildings she wants\\
for her very own on a street she lost somewhere.

Frederick crosses the line, steps off the curb\\
in front of a yellow bus going too fast.\\
The cops hold back crowds who want a peek\\
at a man they never noticed on the street\\
where he snatched purses, money to buy dark birds.

Minerva sweats at the Rouge Plant\\
where steel bubbles, melts, pours into molds\\
that shape her very soul as she clocks in\\
and out. She counts the wages of her work\\
and yearns for the cardinal's song all along.

We are all lemon people dropped from thorny trees.\\
We clutter our space, forget the fields we farmed,\\
ignore the ocean roar---and suck our own tart juice.


\clearpage
\section{A Lesson Written in Ice}

Trust nothing you learned that summer\\
fishing on Pleasant Lake. All you needed\\
was a boat, a baited line and time\\
enough. You caught your limit then.

It was all too easy. The bluegills spawned\\
in shallow water. Females laid their eggs\\
and fled. Males stayed behind as guards\\
and you, my father, set the hook.

I watched you from the dock. You knew\\
I could not swim with bluegills, would not\\
invade their nests. I was your son\\
who never ate the fish you fried.

In winter, bluegills lured you to the center\\
where you claimed your space and carved your hole.\\
You dropped your line into darkness, hoping\\
for fish you could not see---and a son who was not there.

During these frozen months, you wandered\\
on the surface. You heard the ice floor crack,\\
tracked the jagged line of our relationship\\
all the way back to the first bluegill I never caught.


\clearpage
\section{The Lexicon of Streets}

When I was seven, I believed\\
in street names, their promised\\
mystery. I pedaled my bike along Margarita,\\
let my tongue roll around the exotic sound---Margarita.\\
I just knew a Spanish woman clicked\\
castanets somewhere on a porch\\
with a metal swing,\\
big red and purple and yellow carnations\\
painted on the back. Margarita would\\
tuck one real flower behind each ear,\\
clutch one long-stemmed yellow rose between\\
her white, white teeth. She would drape her scent\\
along the swing, put down her castanets,\\
beckon me with her long pink nails, pink\\
enough to braid my tangled hair\\
atop my head and thread\\
two carnations at the crown.\\
She would call me Margarita, too,\\
and it would be true.

By eight, I knew Margarita was nowhere\\
I could find. My tires rolled along\\
Rosemont, a street without roses,\\
across to Shaftsbury where Sue's dad\\
drank Bud from cans he crushed\\
in one calloused fist, his knuckles white.\\
He swore at the union and the company\\
day in and day out until all his grass\\
turned brown and the neighbors agreed\\
he got the shaft. That summer I gave up\\
street names. I believed only\\
in myself and the apple tree I climbed\\
at the corner where I caught everyone\\
doing something.
\clearpage
If I were patient as the pink-tinged flowers\\
time would turn into apples\\
sweet enough to eat.


\clearpage
\section{Lilacs Bloom in the Vacant Lot}

Other eighth-grade boys bring tulips\\
and daffodils that last\\
a week on Ms. Williams' desk\\
without releasing one sweet smell.\\
If I shut my eyes,\\
I do not even know\\
they are there.\\
Their leaves wilt to brown,\\
and the petals drop.

In the vacant lot, next to\\
the party store, I pick\\
lilacs. Purple, pink, lavender,\\
white stars cluster,\\
arch in bunch after bunch,\\
so perfect\\
they look fake. But fake\\
reeks of plastic. Even I\\
can smell that kind of lie.\\
All the way to school,\\
my lilacs bloom\\
perfume in the palm of my hand.

One by one, I arrange my lilacs\\
in Ms. Williams' cut glass vase.\\
The woody stems touch bottom\\
as buds on each bunch pop open.\\
My teacher takes huge gulps\\
of lilac air as if she cannot\\
catch her breath.

Ms. Williams picks\\
a sprig from the vase,\\
tucks it behind her ear\\
and gasps at the truth of it all.


\clearpage
\section{Living Still in Tornado Alley}

\emph{Moore, Oklahoma\\
May 20, 2013}

Monday, mid-afternoon, my world\\
went silent. Robins quit singing.\\
Grey-black clouds hung in violet sky.\\
Even Mimi, my Pomeranian,\\
abandoned her yipping, nudged\\
my feet, her pleas to flee this place.

My fist froze over the small\\
hole I dug in my spring garden,\\
the soil already turned,\\
lush and ready for lettuce seeds\\
clutched in my motionless hand.\\
Plump raindrops splattered, then hail

hit the earth, dented my Ford\\
Focus, bounced in their death dance\\
before the melt on hot ground,\\
before the fierce funnel cloud\\
spiraled down. I grabbed Mimi, my phone, my\\
radio, and huddled in my porcelain tub.

When I peeked over the edge, couch\\
cushions crouched on the tilted floor. I flipped them\\
on top of me just before the tornado's roar,\\
a prolonged growl, a deafening rumble before\\
the weight of the world tumbled down, settled\\
on top of me, trapped until found.

In our confused space, I could be anyplace,\\
even over the rainbow with Mimi\\
who licked me into consciousness again\\
and again before I called the good guys who rescued me,\\
the woman who would plant seeds.


\clearpage
\section{Loaves of Bread}

The little girl who ate words\\
chewed them, savored them\\
on her pink tongue as she curled\\
against her father's side, her head\\
balanced on his arm, her eyes on whatever\\
book he held palms up, tilted toward her.

Each letter on each page\\
jumped into her mouth, tasted\\
new, sharp as cheddar cheese.\\
The whole alphabet jumbled,\\
melted together in a casserole\\
of spaghetti noodles and red,\\
round tomatoes with an acid\\
sting, a sit-up-and-take-notice\\
zing when Heidi, an orphan,\\
met Peter, the goat herder, and shook\\
hands with the hermit in the hut\\
who was her grandfather,\\
his voice cold as the snow-\\
crusted Alps Heidi had climbed.

The little girl snuggled closer\\
as she swallowed Heidi's story,\\
nourished by each chapter,\\
warm as dough baked into loaves,\\
the crusty bread dipped into goat milk,\\
her daily sustenance.


\clearpage
\section{Lullaby}

Rising from the streets,\\
the yellow fog cradles\\
bits of voices:\\
curse words, shrieks,\\
prayers, absolutions.\\
Confusion\\
balloons in her brain\\
in her third floor flat\\
where she rocks, clutching her cat\\
against her shriveled breasts,\\
stringing the street words\\
into nonsense lullabies,\\
ignoring the stillborn dust\\
settling on her window sills.


\clearpage
\section{Madame Tina}

resides in the House of Spiritual Renewal\\
just down Miller Road, a block off Main,\\
in a converted hardware store\\
once owned by Maynard Maynard\\
first and last names\\
the same, no middle,\\
not even on the last will and testament\\
where he left his land and bank account\\
to Madame Tina, who read his fate\\
in the squiggly lifeline dead-ended\\
half way across his palm.

Maynard's final seven years\\
were lived full blast, circling\\
to the top seat of the Ferris Wheel\\
where he cuddled Madame Tina's plump hand\\
against his calloused one, fingers laced\\
together against the chill, skin to skin\\
rocking the whole world away.

Now by day, Madame Tina cradles other hands.\\
Palms up, she pronounces their destiny\\
carved in grooves. Deep ruts speak to her\\
of marriages, births and deaths, no more\\
than jerky starts and stops\\
when the carnival barker fills empty seats\\
on the Ferris Wheel and allows one more cycle\\
before shutting down power at midnight.

Outside the House of Spiritual Renewal,\\
a green neon hand, fingers spread,\\
trolls with filaments of light\\
across the freeway, slender threads\\
of hope Madam Tina weaves into red banners\\
when the carnival comes to town.


\clearpage
\section{Mama's Windows}

Never a smudge\\
on windows Mama washed daily\\
and left cracked all seasons\\
so air skittered across our foreheads.

Like being outside, she said,\\
and right up close\\
in the front row pew\\
with nothing in between\\
but our lives; such fragile souls

needed nourishment and a clear pane of glass\\
that reflected us back to ourselves. Together\\
we guzzled water straight\\
from the garden hose\\
when days broke one hundred degrees\\
in mid-afternoon and bluegrass lawns\\
stiffened into yellow brushes\\
that bristled against our bare feet\\
and water dribbled from our lips\\
wet with words of foam\\
and ocean waves.


\clearpage
\section{The Man Who Loved Paper Wasps}

angled the aluminum ladder\\
into the shadows of his brick colonial\\
as the sun bounced\\
behind the earth's curve. Red\\
welts streaked the wasp-stung horizon.

Dusk called the moon\\
into the sky as he stepped\\
onto the first rung and sized\\
up the paper wasp nest tucked\\
under the eaves, up at the very peak.

The gray cells of the nest\\
resembled moon craters, holes\\
he could fall into if he got too close.\\
Hand over hand he climbed,\\
each rung a foothold in the June air.

All it took, he knew, was a shift\\
of his body. In one unbalanced\\
second, he could topple. A lethal drop,\\
without the wings of a wasp to hold him up.\\
He dared not look down.

He loved the wasp nest\\
above him, the beauty of construction. Admired\\
the workers who spun each paper cell\\
from dead wood and their own saliva.\\
The rested tonight, their wings folded.

At his own door, the wasps stung\\
him more than once. On silent wings sucking\\
the nectar from his sweet peas,\\
nestled against the rusted cyclone fence.

The paper wasps would not hear the poison hissing\\
in the air, the sudden spray of death.


\clearpage
\section{The Man Who Took Baby Steps}

backward arrived nowhere\\
except places he had already been.\\
He practiced twirling as if\\
with the curved beak of his wingtips\\
he might drill the soil\\
deep enough for oil,\\
for a gush of rich black-gold,\\
for a slippery skin\\
of wealth, an ooze of dark\\
promise poured in cans,\\
pumped in silver tankers,\\
zoomed across mountains,\\
cruised through apple valleys,\\
direct to the Speedway station\\
across Second by McDonald's\\
where the man who took baby steps\\
stood on his tip toes\\
by General MacArthur's statue\\
in the town square\\
in the bed of pink petunias\\
with one hand over his heart,\\
his other waving to the girl\\
who left him years ago,\\
waving toward his lips,\\
waving for one black kiss.


\clearpage
\section{Mare's Tales}

I scratch my initials\\
in earth's crust,\\
lose my fingernails\\
in dust at the bottom\\
of this sand hill\\
I will climb\\
and raise my arms\\
in victory like the pine trees\\
that flourish in rock.\\
Roots dip deep,\\
spread underground toward lakes\\
while the circle of spruce\\
holds up the sky\\
with green-needled limbs,\\
keeps it from crashing down on me.

I will stick to the path\\
etched by others' feet,\\
ascend rock by rock through sand\\
just as the wild horses do\\
on this island where tourists tell\\
tales of three mares viewed\\
at the peak, silhouetted in moonlight\\
when mist sweeps in and some say\\
they are ghost mares waiting\\
for a run with the north wind\\
like me who years to gallop\\
with them if the stories are true,\\
if I am strong enough to hang on.

I count each step as I ascend,\\
suck water from this humid air\\
until I reach the top\\
where I rest in trees,\\
waken to pine and know I am\\
not done with my journey\\
when bark and branches beckon me\\
upward to wear their crowns\\
and hand over hand, I move higher,\\
stretch in air\\
where clouds of mares' tails\\
drop low, tickle my fingertips\\
with white, wispy strands\\
long enough and strong enough\\
to wrap around my fist\\
as I kick free of wood,\\
dangle in air,\\
wait for the rain.


\clearpage
\section{Marquette Under Snow}

I have come to the edge of winter\\
where Lake Superior\\
creaks in the cold snow blow\\
jingle jangles a zigzag crack,\\
its message of melt.

I stand in blur\\
as a soft-bristled breeze sweeps\\
across trees, dusts pine\\
needles with blue-white flakes,\\
a flash of light, the swish of tulle.

On my forefinger, I hold a fleck\\
of snow as if it were a black-capped\\
chickadee, a white-feathered wing,\\
a fluff of down, a puff of snow\\
poised in flight, illuminated by sunlight.

I swallow liquid snow,\\
pooled in my cupped hand\\
as warm water rises, vaporizes,\\
hitchhikes in the belly of cumulous clouds,\\
drifts around the world without me.


\clearpage
\section{Marriage}

Soup boils on the burner.\\
Noodle numbers count\\
fall limp\\
onto ransom notes:\\
Carrot pasted to potato\\
slices sanity\\
rises red or orange\\
to slit night\\
to rock moons\\
to thrust words\\
into crevices\\
too dark for moans\\
and vegetable words.

Vitamins\\
throb at her temples\\
as she scorches\\
cotton collars\\
and pays\\
in unmarked bowls.


\clearpage
\section{The Mass Media}

Sitting\\
there. A tall girl (or maybe\\
small) with pointed nose and\\
breasts pushing against her T-shirt\\
shouting (whispering) in bold, black letters\\
(faded slightly from the wash)\\
DO NOT TOUCH.

Waiting\\
for me, you, or anyone\\
to turn her on (or maybe off)\\
to some news flash of a foreign crisis\\
(or even a domestic quarrel) that needs\\
her intervention. She knows how to sit-in,\\
stand-out, join-the-crowd while adjusting volume.

Wondering\\
if she were alone (forever),\\
would her tubes blow,\\
spewing parts\\
through the twenty-one inch screen,\\
slashing passers-by (who stop to stare).


\clearpage
\section{Mata Hari Goes Undercover\\ in Our Neighborhood}

Crouched behind Dad's yellow Buick,\\
my brother and I spied on Mrs. Valentino,\\
guessing she could be an enemy\\
settled in a safe house next to us,\\
a double agent who wanted too much\\
from men who rattled her screen door before\\
they strolled right in with all their secrets.

Everyone in the neighborhood gossiped\\
on hot Kentucky nights when her bedroom\\
lit up, then darkened. Only my brother and I\\
suspected she could steal our secrets,\\
kill us and disappear in another small town\\
where she might be safe from boys and men.\\
Peeking in her open window, listening\\
as her high heels clickety clacked on wood floors\\
answered by the slush, slush slide of a man's crepe\\
soles and then, the sudden silence, heavy in the humid air.

At the dump that summer, my brother and I\\
searched for lost diamond rings and buried treasure,\\
maybe even all the cash from that Brinks robbery\\
just one town over. My brother drew the map\\
of most likely spots and we dug.\\
We found a bike we fixed with a new chain,\\
a picture frame and bones too small to be human.

As long as we could stay away from home,\\
as long as we wrote the stories, we could\\
escape our father's fists and mother's tears.


\clearpage
\section{Memorial}

Let me suckle at your breast\\
of forgetfulness where milk gushes\\
the geyser of life. I do not know

headless statues, craters or bronze\\
plaques under falling bridges.\\
No proclamations, declarations

preparations are enough. I can never\\
recite the chronology of devastation.\\
Too much, too long. Such causes crumble

brick by brick, stone by stone,\\
word by bloody word. The portrait\\
of now and back then. Way back

when death dropped a black cloth\\
over our heads and held the shutter\\
open. You know the acid bath

the acrid fumes I have not sniffed.\\
No memorials to death. No leafless limbs.\\
Just flesh and muscle. Streams of mothermilk.


\clearpage
\section{Midnight Shift\\ at the GM Body and Assembly Plant}

No one ever says\\
Hank is losing the touch\\
or shirking the work, double now\\
with shifts cut and machine cycles revved\\
past capability, past endurance, past\\
perimeters where metal and flesh fuse.\\
No more

does his welding melt surface\\
to surface or hold like lovers,\\
limbs locked, navel to navel.\\
Scarred at the very center,\\
he cuts the cord\\
again and again.

Belly full of fire\\
burns blue\\
as Hank's anger. The hottest flame

after thirty years, his layoff.\\
As if he can\\
just switch off currents,\\
forget the welding gun,\\
the way it fits his hand,\\
transforms his fingertips. Now

five torches blaze. Touch the robot's arm,\\
caress the steel joints. Patch\\
broken parts with solder. Strong enough\\
to hold against repetitive moves, against

vibrations that reverberate\\
off factory walls, off scorched bones.\\
Then burn to ash.


\clearpage
\section{Migration of the Monarch Butterflies}

You know the high price\\
of the fog-filled hours\\
as we wait for sun, wait for our body temperatures\\
to rise. We prepare

for flight. Our bodies\\
feed on fat, on thoughts\\
we stockpile---our fuel\\
beneath our flesh.

We are bound\\
by our instincts, joined\\
in our migration. Our destination:\\
that small valley\\
high in the Sierra Madre\\
of middle Mexico

where the air is too thin\\
for entomologists. They cannot\\
track wingprints or see\\
beyond violet.

I winter in these mountains\\
with the Monarchs: millions of them.\\
They perch in my hair\\
dangle from my earlobes\\
and brush against my neck. They settle\\
on my shoulders. Their wings

paint the when and where\\
of our return flight. We will\\
mate in midafternoon, then\\
lay our eggs on milkweeds.

Our descendants will rise up,\\
rake the clouds with their wings.


\clearpage
\section{Missed Calls}

I didn't hear you knock\\
on my oak door. You know\\
the type of solid rap\\
rap, rap, two or three times with fists\\
curled hard, pulled behind your back;\\
then pound, pound until\\
the frame splinters\\
into sharp, slender wood,\\
fragments that pierce your finger,\\
cut dep as a memory\\
you forgot to remember\\
that day weeks ago\\
you set your wedding ring\\
on the top of our dresser,\\
the shiniest circle\\
among your loose change and\\
you slammed that oak door tight,\\
rattled everything left\\
alone on the other side.

I didn't hear my phone ring\\
even though I kept the volume\\
turned up to the highest notch\\
but I answer it anyway\\
just in case someone needs\\
to talk to me, just me.\\
You, my sister, are already there,\\
about to call me about nothing\\
important. Still your words\\
rush over the long distance between us.\\
You talk, talk, talk and talk some more\\
about tomatoes in your garden\\
red enough you will pluck them tonight\\
red enough you might eat nothing else\\
for days. I hear you\\
over the rattle of\\
blue-flowered plates and cups\\
as you load the dishwasher.\\
you won't start until you're done\\
with your chat with me or until the kids\\
run in from school full of news\\
they don't tell me, their only aunt.

I receive my mail every day\\
from grocery stores where\\
bananas are twenty cents\\
a pound, where flyers urge me\\
to buy ground round for the grill,\\
corn on the cob for a discount,\\
the more I buy the cheaper I get it\\
all, even the plasma TV bigger than\\
what hangs on my wall fed by cable\\
I pay for every month for programs\\
I don't watch, not even the news\\
with its bloody headlines that tell\\
of a guy who shoots another guy\\
who calls him some name he doesn't really\\
hear but all the same, he doesn't like\\
anyone calling him anything.\\
I drop all the ads into the yellow bin\\
for recycling, for making something new,\\
maybe something with my name on it.


\clearpage
\section{Monarch on my Windowpane}

\emph{For Nancy, who counts butterflies}

I do not see you land on glass,\\
do not look up\\
until you raise your wings\\
above your back and lower them:\\
three orange patches on your forewing\\
bordered in black. You signal me\\
with silent message flags, with eyespots\\
where they aren't supposed to be.

I cannot see how you roost there, how\\
you cling as if held by silk threads\\
from your caterpillar days. Invisible\\
strands hold you motionless

against my windowpane. You stay still\\
as my memories. My dreams hang moist,\\
expand under this June sun\\
and unfold into butterflies.

These butterflies\\
carry the sky on their backs, move it\\
from field to field, from day to night\\
without letting it fall.


\clearpage
\section{The Moon Behind the Pines}

Just walk nice and straight.\\
Put one egg in your shoe and one shoe\\
in front of another as if you were taking\\
gigantic steps. No, giant steps.\\
No, humongous, slipping, sliding, flopping steps.\\
Straight lines, angles, skating through\\
velvet ropes. Simon says, ``Find\\
the moon.''

I see the moon. Big. Round. Orange.\\
Sweet edges. All is sugary. You point\\
a finger north.

No you don't. The sky is empty. No moon.\\
No stars. Nothing but darkness. No\\
beginnings. No endings. Just emptiness.\\
Look behind you. It's just the reflection\\
of our hanging light in the window,\\
as if someone took a big bite\\
out of the moon.


\clearpage
\section{Moonstone on My Finger}

You stare at my nudity,\\
at my hands spread flat before you.\\
My nails are polished clear.\\
These transparent half moons arch\\
into pink and grow toward night,\\
reach backward.

You've been there;\\
you are there still.\\
Your hands tugged cow's teats,\\
firm, full of milk,\\
emptied into pails at your feet.\\
The moonstone, the one he gave you,\\
mounted on a silver band,\\
slipped over your knuckle,\\
locked there on your left hand\\
held then against the mare's flank,\\
both sweating in heat.

Inside your frame, hedged in rosewood,\\
there is no room for breath. Like a Monarch\\
butterfly in a mason jar,\\
lid twisted tight, your\\
face caught behind curved glass\\
hovers above my brass bed.

I lie beneath you, translucent\\
as the moonstone, worn down,\\
barren for fifty years.\\
My belly is flat as yours at nineteen.\\
My fingers flexed and restless\\
feel your ring. The metal touch\\
tarnishes my skin, turns it dark\\
as my inheritance from you, my kin.


\clearpage
\section{The Mourning Dove's Missive}

We live in blue onion skin\\
envelopes used for air mail\\
letters to ourselves\\
we sign with wing tips and quills

our feathery scrawl, our message\\
scratched in blue sky, flutters\\
this April morning as you teeter\\
on knotted pine I split yesterday

for beams in my cabin, all wood\\
brown as you, fertile with sadness\\
birthing our blue song:\\
``cooah, coo, coo, coo,'' unrelenting

as we mourn the death of night\\
full of ghosts, a circle of strangers\\
and friends heads bowed, reading what\\
life and loss swirl around us

land in our hands where we hold on\\
another day transparent as our letter\\
fades, glue dries, flaps open wide.\\
I ride your white flash of light.


\clearpage
\section{Mourning Papers}

Uncle Ed reads the obituaries every day\\
at breakfast. He will not eat a bite of toast\\
until he knows who has slipped past him\\
in the navy night, who has ridden\\
the blue heron of death. It stands watch,\\
on one leg, ready for his fall from life.

My uncle has seen the blue heron twice:\\
once at noon when his tractor tipped him off\\
and he lost consciousness and once at night\\
when his heart stopped beating. He swears\\
it is the same blue heron that returns each year\\
and scoops up minnows from the shallows\\
in Pleasant Lake. That heron lifts up,\\
wings spread wide. Its legs are two fishing lines\\
dangling in the blue sky. My uncle watches,\\
shades his eyes from the harsh June sun.


\clearpage
\section{Musical Staff}

Sour note, Mother shouts. Her\\
voice floats from the next room,\\
catches in her unmade bed.\\
Rumpled words, whiny\\
as my violin, a string turned\\
tight as the cat's gut caught\\
under my thumb.

My fingers\\
slide between bridge and neck.\\
I press\\
hard to coax sound,\\
cradle the bow,\\
sink across b flats,\\
c minor notes.

No major\\
movements mar my\\
mother's breath, hot\\
as pepper on my tongue.

The cadence set,\\
we practice every note,\\
every night. Under light\\
too dim to read\\
the score of lullabies,\\
the book of songs\\
is propped on the music\\
stand.

I stoop; I drop\\
my arm, my bow.\\
I bite my violin,\\
chew wood, swallow\\
silent symphonies.


\clearpage
\section{Musical Warfare}

The bugler blows a wake-up call,\\
dark and mellow as one ripe plum\\
about to fall. That sound\\
rolls out the tube. Note by note,\\
reveille drips into a cupped ear,\\
slides down the bone channel,\\
straight to the heart. The soldier\\
wakes to the ancient hunting blare,\\
ready for the battle of brass winds.

The French horn player claims first seat\\
in the symphony. Metal tubes\\
spiral into circles. One deep breath\\
travels through brass,\\
transforms itself into notes.\\
Listeners lean forward, hear\\
that muted woodland song.\\
The fawn drinks from a stream,\\
as water flows and wind blows\\
storm clouds overhead, a warning\\
that war is about to begin.

Troops polish brass and silver\\
until metal glints in sun, reflects\\
musical scores. Lips vibrate\\
into mouthpieces. Tubes resonate\\
with tones. Pitches climb and fall\\
in harmony. The battle\\
flags unfurl from musical staffs\\
like mountain roads wind to the peak,\\
the final stop before descent\\
into green, hushed valleys\\
while the red sun sets.

The war of brass winds\\
reaches back,\\
sounds the hunting call\\
through carved bones\\
and the horns of elephants\\
who won't forget\\
the bugler always\\
plays taps.


\clearpage
\section{My Drawings Have No Relation to Reality}

A simple summer sun becomes\\
a round head with arms cut off\\
at the wrist. Nothing under the sun,\\
not one person, not two trees, not\\
three stick people who are my family.\\
Just blue scribbles like waves\\
or wrong-colored clouds crayoned\\
below yellow on manila paper\\
taped upside-down on the fridge.\\
Not even the apples in the green bowl\\
notice anything out of the ordinary.

Trapped on a glass slide, protozoans\\
whip their tails, wiggle and fidget,\\
my lab partner claims. But I do not\\
see them through the microscope.\\
These one-celled creatures split\\
in half, create their own families.\\
I draw only blackness before adding\\
what should be there.

Doodles decorate every note. In every\\
blank space I sketch hopeful hearts\\
and purple violets. Every letter\\
I write home contains nothing\\
real about my life, only drawings\\
of what will never be.


\clearpage
\section{My Doors}

Fish swim in\\
and out my five open doors\\
carrying messages, wet mysteries.

That catfish at my threshold\\
suctions mud, purifies an entryway\\
for walleyes with hypnotic stares.\\
Large-mouth bass speak in code\\
from a door above. A single minnow\\
tows a line and a Sunday fisherman\\
whose lips brush mine. The rush\\
of lust dangling, bobbing\\
in and out my doors as if I were\\
the Detroit River, a channel to somewhere.\\
The northern pike, round goby, yellow perch,\\
black crappie, sucker, even the common carp\\
bait me with a silver flash of truth.\\
They spawn, feed, move on\\
upstream, downstream. Some stay near shore.

Fish wiggle my maze of veins,\\
open my sixth door. I see\\
no doors lock. I am river.


\clearpage
\section{My Father Grew Dahlias}

``Nothing,'' I told my father,\\
``I want nothing.'' For sure,

he wanted his job back before Christmas,\\
wished his hands still tightened the vise,

operated the shaper, cut metal once more\\
within tolerances. Layoffs bloomed

that hot summer in Detroit. Tempers split the humid air\\
when work trickled in as if dripped from a leaky hose.

My father planted dahlias, one hundred tuberous roots\\
and staked them against the whoosh of the west wind.

Soil lined our fingernails as we yanked weeds and father\\
fed me Dahlia names: Blackberry Ice and Cherry Drop,

Candy Cane and Apricot Sun. He answered newspaper ads,\\
hoped for a call back and deadheaded dahlias, whistling.

At first frost, the dahlias turned brown. My father predicted\\
snow on Christmas. He tracked jobs, kept in touch and bought

nothing new. White flakes glistened like wet December\\
dahlias. Petals of snow crystals piled up.

Out front and three houses down that Christmas day,\\
my father had shoveled the walk. When I straddled my new

red bike and pedaled away, he shaded his eyes\\
and waited until the tires rolled me back to him.


\clearpage
\section{My Grandson's Preschool Assignment}

\begin{quote}
\emph{Typed note: Fill this bag with pretty colorful\\
leaves. Bring it to class Wednesday.\\
Your Teacher, Miss Amy.}
\end{quote}

The day the leaves dropped,\\
higgledy-piggledy at the park, my grandson\\
squeezed two maples and one birch\\
in his bare hand, crumpled them to bits\\
he threw in the air like confetti,\\
pigments of marigold orange, red,\\
dusty brown he breathed in when\\
he heaped leaves over his head,\\
over his green-leaf body.\\
He licked dry oaks from his lips.

My grandson crammed the bag\\
with crinkled leaves, acorns, two beach stones, a Canada goose\\
feather and two pretzel rods,\\
salted away for one more fine fall day.


\clearpage
\section{My Machine Project}

In my mind's closest cave,\\
I scratch dimensions on walls,\\
calculate the speed of time,\\
a linear line that stretches\\
from past to present\\
where I reside and mark my days\\
as clocks tick. Alarms sound\\
my wake-up call. I add numbers,\\
subtract for loss, multiply minutes\\
into round-circle days\\
without end.

I sketch loops of my infinity machine\\
with no openings for forgetfulness.\\
My mind sees the possibilities.\\
In my dark cave, my charcoal traces\\
geometric shapes and how they fit\\
together with straight lines,\\
the shortest distance between\\
what was, what is\\
and what could be\\
if there were no limits,\\
no boundaries.

Others have been here before me\\
in my cave, in my mind, in my muscles\\
which remember another time\\
when people crawled,\\
then walked upright on green land.\\
In my machine, I will take you\\
back to the beginning\\
where we started,\\
where sunlight illuminates\\
shadows we carry forward\\
and backward through time.


\clearpage
\section{My Sweet Onion}

Ah, sweet Vidalia,\\
my wild onion child,\\
I watch your bare feet\\
root in this Georgia field\\
as your golden-brown legs\\
draw nourishment from red clay\\
days under sun, a harsh spotlight

as you twirl from row\\
to row full of bulbs and seeds\\
bursting through soil, cultivated\\
by you whose grasshopper legs\\
leap lettuce leaves nibbled\\
by deer who eat from your hand\\
when you whisper in their ears

your plans, the rhythm of your heart\\
pumping visions of pirouettes\\
on stage where you strengthen\\
toes, muscles tough enough\\
as you whirl along the side\\
of your rented shack, performing\\
for family who hum, play harmonica

while you, sweet Vidalia, ride\\
each note until it holds you,\\
a season of layers circled under\\
your translucent skin no one\\
peers beneath but you who know\\
one day your body will be ripe,\\
ready for northern cities, the dance

where you wear purple ribbons\\
curling in your hair as you\\
shed dusty skin behind footlights\\
and stretch toward one sweet note.


\clearpage
\section{My Two Children Leave
  \\ their Childhood Home}

If you must leave this house,\\
do it quickly before dust settles\\
in my hair, before I grow old.

Take seeds from forget-me-nots\\
by the drive. Plant them deep\\
as you think of me raking our garden.

Don't forget that chunk of quartz,\\
that piece of driftwood you carted\\
from the shore of Lake Michigan.

Remember our redbud tree blooms\\
in full sun when purple leaves explode\\
to pink. Still, some call it the Judas Tree.

Life is not all the peaches you eat.\\
I ask you to watch your backs,\\
cross at corners after looking both ways.

I grow toward you, you bend away\\
as it should be when you leave.\\
This house was built for me.


\clearpage
\section{The Nature of Vipers}

I was born on the dark side\\
in the year of the snake.

When I slept coiled,\\
my knees grazed my chin.

As I grew, I sunned myself,\\
stretched in the arms of bad men.

I kissed the lips of cottonmouths,\\
sipped their venom and whiskey words.

The gossips said I shed skin after skin,\\
and tacked them on my bedroom walls.

Tales about me plumped with lies,\\
spread across back roads where I rested.

In the heat of August days, I savored\\
the true dangers of the raised rattle.


\clearpage
\section{New World Spices}

A small dark woman who had no words\\
in English, only a pan on a gas stove\\
where she whipped exotic spices into a frenzy

of ginger, cardamom, saffron, cinnamon\\
that pinched my nose, tugged at the back of my throat\\
burned my tongue with memory

of smells she dished out to neighbors\\
along with the recipes her daughter translated\\
into a new language we all spoke.


\clearpage
\section{Night Passage}

Stars lead me to the edge\\
of Lake Superior where I find myself\\
alone. Sand scrapes my bare feet.\\
I step on stones, stumble on bottles\\
full of emptiness, left behind\\
on this beach where I curl my toes,\\
scoop up mounds of sand. Each grain\\
carves its message on my skin,\\
the lesson of horizon\\
where earth meets sky.

I baptize me in cold water, shiver\\
in warm air. Draped in black night,\\
I spot a Great Lakes freighter slice\\
through the calm at a steady pace.\\
The moon illuminates one side\\
as if that vessel were a shadow\\
of me steering toward a destination.\\
In the Pilot House, someone else\\
moves this night, awake in starlight.

When I look up from the horizon, I see\\
all those stars, too many to name. The sky\\
blazes with star gas. Celestial candles\\
flame white and yellow,\\
reflect the freighter's water self\\
as if it floated\\
up from the bottom of the lake.\\
Even the smokestack duplicates itself\\
in sky and water. Smoke drifts to clouds\\
above the horizon, marks my night passage.


\clearpage
\section{Nighttime Meditation:\\ A Woman and Her Tabby Cat}

This is a quiet poem,\\
full of hush and purrs.\\
My tabby cat claims the kitchen\\
corner in the evening after supper.\\
Her nose nuzzles her ginger paws.\\
Black, cream and cinnamon hairs\\
quiver as she surveys her territory.\\
Her green eyes glare at a blue\\
cup out of place on the counter,\\
pierce me as I scoop\\
tender ocean whitefish from a can\\
into her blue, hand-painted china bowl.

Tonight, I am slow-moving,\\
even contemplative. My steps\\
to the chipped porcelain sink\\
shrink in size as I think of all\\
the news I refuse.\\
At least tonight\\
I will relish domestic acts\\
of nourishment---the slosh\\
of chamomile tea against\\
my cup, the silence of dust\\
as it settles on the TV screen\\
and the ripple of bath water\\
warm against my skin, soaking\\
off dirt and stains I will not let\\
color such a tranquil scene.



\clearpage
\section{The Night Watchman and His Flashlight}

After the sun doffs its top hat. After it boogie-woogies across the
slippery oak floor. After its patent leather shoes tip-tap out the grand
finale\ldots then, and only then\ldots do we applaud daily at the Gem
Theater when spotlights dim.

Two stagehands (one stage left and one stage right) pull the drapery
cords. Hand over hand. Almost as if they can shimmy up those ropes,
thick as their wrists. They strain against gravity until the tattered
purple curtains, frayed at the bottom, drag the strings of our days
across the unlit stage. Velvet seam laps velvet seam at stage center.

Smaller than the sun. Powered by four D batteries (negative and positive
charges at opposite ends like a fable and a truth, linked by electric
metaphor)---the night watchman's flashlight. It is his only weapon
against darkness on his appointed rounds when he tests outside security
locks on exits. And yes, he checks the locks on the iron gate secure
across the entrance.

The night watchman times his sunbeam flash to the second on his
luminous, green, glow-in-the-dark, digital wristwatch strapped against
his moonlit skin. Meticulously, he prints numerals and documents each
time and date in log book columns lined with red. He, the bringer of
light to night, believes re-chargeable batteries in his light of flash
will never die.

A crash, a clang, a crinkle. The watchman flicks his flashlight on/off.
On again. His light lines write their poem (or is it a story?) across
the Gem's brick faade. Inside no concern of mine, he mutters. Outside
safe, he nods.

A skittering sound. Inside. A creature drawing close. Off again, his
flashlight dangles from his fingertips. Crackle, slurp, crackle, slurp,
bam, blast. The orange-blue fire within the Gem flares up and crunches
beveled windows between its jaw. Then, ever so politely, this flame
wipes its own burned lips with smoke.

On, on, on. He clicks the red button. Shakes the chrome cylinder. Four
batteries hiss, release their charge.

The night watchman can not record the exact moment, can not log the date
of darkness when the Gem turned inside-out.

Phantoms float, stick\\
in his throat,\\
damp, a bitter taste.\\
Over the years, stalagmites\\
grow up from the soles\\
of his feet,\\
pierce his heart\\
where fish swim\\
blind from lack of light.

He breathes hard, fast,\\
hyperventilating, giddy\\
with oxygen. He strikes\\
matches. Sulphur burns blue.\\
Sparks drop, catch again\\
on cardboard cartons,\\
rise to mate with diesel fumes.

He keeps watch. Alert\\
behind bulletproof glass.\\
He waits; hands fold in prayer.\\
Floors explode. Geysers of fuel\\
erupt into flames. Burn clean\\
the night. His visions of light.\\
The legacy of loss.


\clearpage
\section{Northville Convergence:\\
  An Architectural Landscape}

{\bf 1820s: Settling In the Natural Way}

A free place, a space\\
of imagined shadows chased\\
with light, with wind\\
wiping air clean, whipping

through evergreens,\\
branches brown,\\
blue-green needles of spruce\\
and oak, roots\\
reaching deep into rolling hills.\\
Cattails and weeping willows\\
anchored in marshland.\\
Nature's architect sketched

possibilities. Logs, trees, earth,\\
water, air: the stuff\\
of life. Women, men and children\\
came, claimed land,\\
cleared their property. The Tibbits,

McFarlands, Starkweathers, Bartows,\\
Phillips, Bentons,\\
Simons, Hungerofrds: more than\\
merely names. Pioneers who built

foundations, laid dupporting beams, shored up\\
roofs, erected homes of heavy timbers. Frames\\
for their existence, their comings and\\
goings in and out. The sun, the phases\\
of the moon their calendars. Days colored

by seasons: cyclical, rhythmic\\
as their heartbeats, that pulse of blood,\\
a natural way of settling in.
\clearpage
{\bf 1830-1900: Truth Tellers}

Churches rose from farm soil, homes\\
of worship. Presbyterian first, then Methodist,\\
Baptist, Catholic, Lutheran: arched doors,\\
frames, bricks, stained glass. Geometric\\
shapes of faith. Worshippers kneeled\\
in aisleways, against the wooden pews,\\
fingers touching trees and visions\\
of what could be, would be\\
as they tended\\
crops of corn, orchards\\
of apples. All truth tellers\\
transforming wilderness. Rows\\
and rows of wheat, hay,\\
rye hoed and harvested. Stored\\
in barns whose plank sides weathered\\
over winters, swelled with water. Warped\\
doors shaped themselves, skimmed snow,\\
cleared paths for women gathering eggs, warm, white/brown ovals.

Smooth/rough buildings brick or frame, one\\
story or two. Joined in rural\\
celebration: a community at work.\\
Houses, a school, the doctor's office\\
in his home on Main, a general store.\\
Roads rutted from wagons and horses'\\
hooves. Linked paths, packed tight\\
from loads of produce,\\
logs, and livestock, even\\
people moving up and down steps, outside/

inside, entering/leaving one place for another.\\
Mills processed logs, split\\
into more structures. Victorian\\
architecture: intricate carvings, parapets\\
of stone, steep-pitched roofs, porches shaded\\
by sloping eaves. Places where folks\\
could talk and live. New building

plans with mystical rooms. Attics\\
nestled under high mansard roofs,\\
the double slopes, the angles.\\
Basements, cemented\\
in ground, strong as the White\\
Mulberry on High Street. There\\
for the duration. This hardy

stock of people, not afraid. They\\
bred horses, thoroughbreds\\
with lineages flowing back through\\
history, across the ocean. Embraced

change and controversy. Workmanship,\\
after a fashion. They carved moldings\\
scrolled above their heads, up\\
and down entry ways, leaded\\
glass, windows shuttered, painted\\
wood and stained walnut, mahogany\\
and Black Maple. Dark

as their stallions standing guard\\
against the night. Unbridled, unmounted\\
moving muscle, poised for flight.

Form and function. From\\
the Yarnall Gold Cure\\
Institute on East Main, promising\\
and promoting miracles, to Northville's\\
spring-fed well. They came

and drank. Silver Spring\\
Mineral Waters bottled and sold,\\
cold. Electric\\
railway tracks down Main,\\
past a grocery store, clothing\\
store and all essentials. A blacksmith\\
shop where fires burned hot, melted\\
metal shoes. Lined up, eclectic

buildings, blends of this\\
and that. Purity\\
of Queen Anne arches,\\
built to last. Studies\\
in grace, simplicity.

A century closed, opened\\
doors. Truth tracked through\\
architecture; the story\\
sung by warblers perched\\
on gables, the highest peaks.

{\bf Early to Mid 1900s: A Gathering Together}

Fire's flames licked up houses, Park House\\
Hotel, Globe Furniture Factory, other\\
places as if they were\\
ice cream cones melting\\
on a hot day. The train picked up

dropped off people who came and\\
went at the railroad depot, into/\\
out of Northville. The old crow's nest\\
at Main \& Center circled\\
by crowds. They heard round\\
notes of trumpets, drum beats, the whole\\
band above them. Sounds wafted\\
down, settled like cloaks\\
around folks' shoulders. People

came from nearby towns, stayed\\
for fun. Pinned on blue ribbons.\\
Cows, quilts, jams, cakes judged\\
at the old Wayne County fairgrounds,\\
in season. Just down the way,\\
a tournament ski jump. All wood\\
and snow. Layers of white winter, soft,\\
slick and smooth enough. Ski\\
tracks, ending at the end\\
as skiers soared, their poles\\
tucked like wings\\
into air, before the jump

and landing. Whole families balanced\\
spaces at several levels. Symmetrical\\
lives linked by design. Interior/\\
exterior mirror images like Spagy's\\
Grocery Store with shelves stocked\\
and living quarters up above,

fully furnished. Modern ranches on the edge\\
of town: brick, wood and aluminum siding,\\
all shades. Builders or owners painted\\
masonry light blue, green, white,\\
yellow or left it\\
bare in bottom line buildings\\
exposed. Even cement shifted,\\
cracked, required constant\\
vigilance. Maintenance\\
of contemporary styles, showplaces/\\
showcases of new trends tracked

as closely as bets at Northville Downs\\
where people came\\
to win against the odds. They sat\\
on bleachers, leaned forward, stood\\
as ponies rounded corners, watched\\
under artificial lights. Fluorescent\\
against the night. They lost\\
or won and came back again

drawn by neon signs, fast food.\\
They hiked trails worn by feet\\
before them, played in parks\\
and streams. They wandered\\
streets, shopped at Freydl's, Schrader's\\
Home Furnishings, Lapham's Men's Wear,\\
Ely's True Value Hardware, Brader's\\
Department Store: all still run

by founding families who settled in\\
and stayed. Some came for jobs\\
at Ford or other places.\\
The water wheel turned\\
as they moved in, a gathering\\
together. \emph{The Record} headlines,\\
and photo cutlines captured past,\\
present, future: preserved in print.

{\bf 1950-1994: Soul Sketches}

Sketchers, artists all, drawing\\
what they saw. Thin lines\\
thick as their very hopes.\\
They constructed dreams, homes\\
of sleek lines, lots of glass\\
and stained wood, set back\\
amongst the many maples still\\
standing, away from gravel/\\
dirt roads. People traveled

from other places and spaces\\
then and now, time restored.\\
Old Victorians remodeled\\
rewired, refurbished, renewed:\\
all according to plan or luck\\
or even sudden discoveries\\
of what lay beneath after\\
they scraped flaking, fading\\
colors on dormers, on gingerbread\\
trim, right down to original\\
timberwood, cut and carved.

Transitions, transformations\\
ticks of the eye. A ranch\\
redone; now a beauty\\
salon, specials listed\\
in the picture window. Restored\\
homes inhabited by attorneys,\\
doctors, antique dealers, insurance\\
companies doing business.

In the Center of town:\\
the landmark clock emerged\\
from the old crow's nest. Park\\
benches with hearts carved right\\
where people's shoulders leaned warm\\
against the wood, a resting place.

From there, they could see up/\\
down Main: the Marquis\\
Theatre where actors/\\
actresses practiced lines, listened\\
for applause, cupped hands\\
coming together, again\\
and again. Or the yeasty smell\\
of poppy seed rolls\\
and loaves of bread baked daily\\
at the Great Harvest\\
Bread Company. Or Genitti's Market\\
Place where families ate, became\\
parts in a play. A natural evolution\\
from grocery store to restaurant.\\
Singly and together people\\
window shopped, entered/\\
left stores with a range\\
of names: Fleet Feet\\
Sports two doors down

from Changing Seasons or up\\
the other way, Cobbler's Corner,\\
Preferences, Orin's Jewelers full\\
of gems, cut carefully, polished, set\\
to catch the eyes of passersby.

Northville looked bac, looked\\
forward and captured time. The hour\\
hand, the minute hand, days/\\
weeks, even years converged:\\
a single point where people\\
strolled sidewalks, all the time

knowing who they were, where\\
they wanted to be. In Northville.



\clearpage
\section{Not Knowing What}

to call those two birds\\
by name, by species,\\
indexed in field guides,\\
lost alphabetically.

In the same way,\\
your name melts\\
on my tongue, tart\\
as cinnamon candy. Red\\
dye stains\\
tongue and teeth.\\
Our talks, over touch\\
too tentative, too full

of nothing but\\
sugar, spun at high speed,\\
puffed with air, not enough\\
to grab onto. Like planes,\\
our marriage recycles oxygen,\\
dry as old arguments.

Just that once,\\
I spot them\\
at the feeder, wings tucked,\\
heads cocked like guns

before metal hits flint.\\
Detroit on New Year's Eve,\\
near dusk. Their saffron crests\\
burst brown tonight,\\
two lone crocuses above\\
snow. The whole world

upside-down, in flight.\\
I see two\\
birds, feel feathers\\
brush my eyelids, listen\\
for wings fluttering

against my lips, the sound\\
of breathing in and\\
out. Not even words,\\
hard as sunflower seeds,\\
can fill the spaces\\
between us.


\clearpage
\section{No Use for Stones}

As a kid, I never even skipped\\
one pebble part way across\\
the blue plate of Cass Lake or\\
sprinkled gravel on a mud pie or\\
dropped rocks from the edge\\
of the limestone quarry outside town and\\
watched the silver circle lariats\\
rope in anyone who dared swim there.

I will step off gravel walkways\\
at the museum of wild things\\
tucked away behind the Brookman\\
Building which is about to tumble\\
chunk by chunk into its own bowels.\\
I walk on crepe soles\\
and wear thick cotton socks although\\
concrete blocks have barred my way\\
and I have stumbled on a sidewalk pebble.

On this warm May afternoon,\\
the skyscraper where I work\\
spit me out into a blue-sky day\\
with the sun still high,\\
huge as any river boulder,\\
shiny as a hunk of fool's gold\\
I might carry for luck\\
until it burns\\
a black hole in my pocket.



\clearpage
\section{The Odyssey}

People come at her as she lies\\
woven into her narrow bed\\
by spider-lines\\
drawn tight against life's loom.\\
Transparent fibers bind\\
her fresh-spun dreams,\\
bring faces back again\\
in the city's oiled night.

It's him, above her,\\
coming back in darkness,\\
his marionette arms\\
pointing behind him\\
at vacuous woman-faces\\
sewed into her brain\\
before he left her bed\\
and chose\\
grey-eyed girls with silver bodies\\
to walk his soul past at midnight's ledge.\\
He talks to her of times they've known:\\
blackberry brandy years\\
when moon crickets\\
serenaded the sounds\\
of bodies\\
coming together.

She cannot close eyelids\\
stitched open with spider-threads\\
or move her flattened fingers\\
from the tapestry\\
to touch the flesh\\
of those who greet her.

A childhood friend returns,\\
older now,\\
too old.\\
Sunken breasts droop;\\
the paper voice recalls:\\
worn Kentucky blue grass\\
between their houses\\
where painted dolls lost smiles\\
in the rain\\
as two girls snuggled together\\
under crazy-patch quilts,\\
talked of boys\\
and flat chests growing\\
into cupped hands.

Now her mother,\\
dragging Father behind,\\
kneels over her,\\
murmurs magic chants.\\
Onion skins, mosquitos' wings,\\
Circe's voice, one black hair:\\
the midwife's cure.\\
Two voices remember:\\
cypress knees skinned\\
on gravel swamp paths,\\
summer's hair plaited\\
against autumn winds.\\
The child who was\\
before she came\\
in wedding white.

Morning light rewinds\\
tangled threads\\
and she rises\\
from deathbed dreams.


\clearpage
\section{Off Days}

Every day, I take my place, set the pace so\\
my students can teach me,\\
make me know my grammar rules,\\
mark my quizzes,\\
argue grades, write my words, read my books,\\
whisper comments, test my patience.

But then comes the day when schedules are short,\\
bells buzz in the middle, students are muddled,\\
and Alice enters Wonderland.\\
On those days, I shake them up,\\
leave the walls.\\
They hear shoes in grass,\\
find the smallest ant,\\
rub dandelions into concrete.\\
They dance in bodies\\
freed from desks\\
and speak in poems.


\clearpage
\section{On Reconnaissance}

I spot one woman's coiffed head\\
skewered on a fencepost\\
beside an empty shack.\\
No body. No extremities.\\
Just a plastic head in this war zone\\
jammed onto a rusted metal tube.

I curse whoever\\
concocted this atrocity,\\
this violent monstrosity. It could be me\\
with brown hair, flipped at the end\\
and querulous eyes exposed\\
in this mannequin's molded face.

Like me, this dummy occupies a space\\
someone else chose. Could she be\\
a warning, a woman\\
in a morality play\\
who lost her tongue\\
spying for the enemy?

I cannot decipher\\
the coded message painted\\
in pink sunset on her lips\\
turned up at the corners\\
as if she knows the secret\\
handshake for my safe passage

through this bloody world\\
where I must wear camouflage\\
and streak my face with mud.


\clearpage
\section{On Washday}

She tie-dyed her soul Monday\\
and hung it out to dry\\
on the thin, frayed rope,\\
hung between two locust trees\\
whose lacy shadows sewed themselves\\
to dripping cotton, sagging in the sun.

Twisted rubber bands, knots upon knots,\\
wrapped white spots bound tightly.\\
Bleached blackness soaked scarlet.\\
Jagged green lines wiggled the length\\
as starburst golds stretched the cloth\\
until it split, slightly, along the seams,\\
opening itself for her inspection.


\clearpage
\section{One Closet in a Clean House}

I don't like clutter, the blue sweater, too small,\\
a hole in the sleeve and missing a button,\\
the yoga mat I spread out one day and put away\\
after the downward-facing dog position threw out my back,\\
challenged my perspective about people walking bulldogs\\
or collies or even Labs big enough to ride.

Into the closet willy-nilly, I toss the yearly letters\\
from Aunt Tillie\\
who complains I don't visit or call just like\\
my mother, she says, who ignored people\\
related by blood and took up with strangers\\
she met in dry cleaning establishments, bless her\\
soul, dead now and probably lost somewhere.\\
If I search, I may find my mother there in the closet.

Imagine a mishmash of books I never finished,\\
words I can't pronounce, criticisms that pricked,\\
good-byes I never said, limp handshakes,\\
the chipped cup from my Willow ware set, the only\\
damaged piece after Grandma entrusted it to my mother\\
who gave it to me and I broke a bit of blue from that cup that fills\\
me with guilt since my daughter won't get a full set\\
of anything from me, not even the key to this closet.

This closet\\
seeps under the foundation of my home\\
and leaches joy away.\\
Still I dust, mop, rearrange furniture\\
in my tidy living room.\\
Who knows when the doorknob of that closet\\
will wiggle, when the monsters will break free.


\clearpage
\section{One Venetian Blind}

ivory as a piano key\\
stained with blues\\
on alternate Tuesdays\\
at Alvin's on Cass

where Sweet William\\
thumped his life chord,\\
his fingertips blunt,\\
swollen with sound,\\
curled down. Down

past yellowed keys\\
slick with sweat,\\
past morning and night\\
that open-and-close death\\
when his heart stopped,\\
and he tripped

through the venetian blind,\\
all slats tilted\\
toward one perfect note

blue as the blackberry brandy\\
he tossed back. Neat\\
and clean, the kicker,\\
his last kick,\\
the bucket flying up\\
and out of sight\\
as his music kissed the wind\\
down the long avenue,\\
light flying into light.


\clearpage
\section{One White Petal}

Sibilant carnation---white\\
against the preacher's black lapel.\\
Hiss as petal rubs petal. As petal\\
tilts from his wool crepe suit, drops,\\
drifts. That one white petal

descends, nestles now\\
in the crevice between\\
his parishioner's breasts.\\
She shivers having known before\\
the pleasure of a petal

translucent as her skin---hush now.\\
She is fluent in tongues,\\
in one white petal,\\
in preacher men\\
who pray carnation tunes.


\clearpage
\section{The Other Woman: Three Lives Left}

I. Cats Die by Degrees

You won't know her by her red hat,\\
brim turned down, ribbon red around\\
the crown with streamers half-way down\\
her back. Nothing so blatant. No,

she won't announce herself, won't\\
even appear in appropriate dress\\
or undress. At least\\
to you she won't. She will

wear whatever suits her\\
as long as it uncovers curves,\\
serves her very self\\
curled like a cat,\\
tail tucked and wound around\\
her paws. A furry mess

of claws, sheathed in pink\\
mounds of skin, under cover.

She will not answer\\
to any name or come\\
when you call; again\\
and again you try. She will remain\\
lost. When you least expect it,\\
she will stalk you. Watch for her. She will

arch her back.\\
Each strand of her amber hair\\
will carry electric currents,\\
measured in watts. She will

weave her seventh death, in and out\\
between your legs, shedding what\\
she is and what\\
she is not.

II. Cats Evaporate Eventually

You listen\\
for the jangle\\
of her collar bell. The jingle\\
and tinkle of that cat, her rustling\\
movement through weeds. You see

but not quite the way\\
her muscles pull her\\
hind end tight, then\\
loosen. Her whole body\\
solid as bone:\\
a subterfuge. She will

avoid you. She will\\
purr her curse. Quietly,\\
she will lie still\\
as your prayers rise\\
and fall. She will wait

until goldenrod yellows. Its stalks\\
grow straight up, ruffled\\
by Michigan winds\\
carrying the scent\\
of winter: that acrid, musty smell

of deceit. She will wind her way\\
down walkways in daylight, climb\\
the Red Delicious apple tree,\\
drag belly against bark. She will\\
crush apples into cider, discard\\
peels in the compost heap\\
to ferment there\\
behind your clapboard house.

She will rise\\
from frosted ground\\
like steam, a mere vapor,\\
insubstantial as the mist, a fog\\
that settles into your soul\\
softly, cold as snow. You know

her by her fragrance, by the way\\
your nostrils press toward bone,\\
sucking in her breath. A battle\\
for her eighth death: and even number.

III. Cats Count Number Nine

You pour\\
her past life\\
across the floor like white,\\
thick, condensed milk.

She will lap it up. She will\\
lick her own eyes closed,\\
tongue rough\\
against the truth:

you alone whistle\\
knowing full well\\
your lover, your mate\\
left you both

before any of you came\\
together. You came apart.\\
She will

wander your yard in January,\\
February: these hard months\\
crammed at the beginning\\
and ending of it all. In time

after the earth softens, you will\\
dig her grave and bury her\\
with all her lives, all\\
her bells and a blue saucer\\
filled with cream:\\
thick and rich\\
as your lover's lies.


\clearpage
\section{Our Insignificance}

I capture this rectangle\\
of black and white,\\
as if it can be framed\\
by my Instamatic camera.\\
I freeze surface scenes\\
but not what's below,\\
not what the next moment holds,\\
not what develops in the acid bath.

A ship may sail on land\\
if I double expose a frame.\\
A businessman may strut the sky.\\
Once I juxtaposed a city on a sea.\\
When I blink, the city sinks\\
and lands on sand. Yes, I am\\
a magician of motion\\
when the shutter clicks.

I shoot a strand of hair, a blade of grass,\\
a handful of air. Life wiggles there.\\
I see it move.\\
I wave my wand,\\
chant an age-old spell.\\
All is frozen\\
when I push the button down.\\
I airbrush out the flaws,\\
blow up what I want.\\
Now you see it.\\
Now you don't.

No, I won't tell you secrets\\
of my tricks. I won't\\
disillusion you about clean,\\
uncluttered places I have been.\\
I have also watched the rush\\
as city feet crack concrete walks.\\
\clearpage
We embrace the magic of open space.\\
We clench our fists\\
when we recognize\\
our own insignificance.


\clearpage
\section{Performance Art}

Women loved Oliver. Oliver loved women but not himself. Oliver edited
his image, hired a PR agent. Olivia got her start pounding drums in an
all-girl grunge band every Saturday downtown.


\clearpage
\section{Piano Keys and Peacock Feathers}

\emph{---for Zena}

The peacock's beak vibrates\\
against the strings of chicken wire\\
as he pecks a tune up\\
and down the scales,\\
a piano concerto for peahens.\\
The notes rush into the crowd,\\
settle on wing feathers,\\
stir the females' grey-brown squawks.

The peacock arches his feet,\\
saunters along the keys,\\
a slow shuffle of ivory\\
and ebony as his blue-green\\
feathers glitter\\
against the dark and light.\\
He slides the sweet melody\\
into the hot August night.

A frenzy of notes now,\\
a flutter of sound loud,\\
louder. The peacock screeches,\\
rattles his tail feathers,\\
dances on each key. Sage green\\
hammerheads strike steel wires\\
as the peacock\\
sets sound free.


\clearpage
\section{Polar Bears at the Detroit Zoo}

\emph{---for Bradley who chose polar bears}

Don't miss the Polar Passage,\\
a new exhibit, a natural habitat\\
where polar bears dive off the edge\\
of Greenland into the Arctic\\
Ring of Life. Their splashes\\
streak the outer tunnel\\
of arched glass where tons\\
of cold salt water and white\\
fur tumble past tourists\\
whose feet shuffle along the floor\\
and whose damp hands press glass\\
curved beside them, hope\\
a paw spreads out in answer\\
as if no glass came between.

Before the zoo, these polar bears climbed\\
stairs at the circus in Puerto Rico\\
and slid down pink plastic slides\\
as humans paid and cheered\\
for more as the bears dwindled,\\
lost fur, lost heart,\\
and grew sick on heat-soaked days.

Before the circus, these polar bears romped\\
in Greenland or maybe Alaska, in some faraway\\
snowdrift where black-padded paws\\
left footprints on glaciers. These bears sniffed\\
seals below ice and kept vigil at the breathing holes.

If the major and minor bears exploded tonight,\\
if the north star swam south,\\
if the polar bears escaped unbreakable glass,\\
who would tell them ice is melting everywhere?


\clearpage
\section{Petals}

Four separate seedlings bloom their beginning---\\
one purple sweet alyssum close to the ground,\\
a squat marmalade marigold, a hot pink impatiens\\
with a white throat, and a viola dressed\\
in five loose petals above heart-shaped leaves.

Above these flowers my sweet-faced,\\
green-limbed grandson tilts his watering can,\\
wants more leaves, more flowers, more\\
life as water pours from the spout.

Five viola petals fall ever so slowly\\
onto the black earth before he bends\\
and picks up each yellow-orange petal\\
and puts it oh so carefully in his pocket.


\clearpage
\section{The Photographer and Her Camera}

After completing a fine arts curriculum at an elite college in New York
and purchasing the very finest camera and equipment money can buy, the
photographer returns to her home state to create documentary essays:
those mini-seconds of perfection when a person, place or object reflects
back its very essence and everyone will applaud her work, buy it for
their walls and make her rich and famous (unlike poets' shadowed images
buried in measured horizontal lines across pages with white margins in
unopened books stacked on shelves in entertainment centers, part of
living spaces, everywhere).

Not that the photographer has anything against poets. No indeed. In fact
she admires their subterfuge, the way they blur real life with similes,
as if any two objects can possibly overlap. She aims for crisp lines
that delineate where one person starts and stops and remains contained
within those lines.

She loads only fast-speed, black-and-white film and lets the viewfinder
see. Piles up shots of herself, of places she goes, people she knows (or
doesn't know). Houses, cities, blue cornflowers in bud, weddings,
deaths, newsbreaking crimes. Only she, the photographer, controls the
shutter, the amount of light she lets in or filters out.

In her basement darkroom, she resurrects the images and washes them in
chemical baths, saving the fixer for last.

She turns negatives to positives. Flips and crops what the eye sees.
Enlarges. Shrinks. Finally, the photos emerge from their baths. Close-up
portraits: her documentation. Each fine wrinkle, each split end, each
pockmark, each patch of oily skin. In the name of truth.

No one buys her work. No one attends her gallery shows. She offers
champagne and a toast: \emph{mea culpa, mea culpa}, when my camera
deceives and I believe.

The photographer and camera wait, its eye covered. No one comes. No one
sees but she.

Later. Alone in her darkroom. Where no one knows: she superimposes,
under-exposes, airbrushes flaws, tints black and white with color,
erases any spaces. A master of disguise. Razzle, dazzle, presto. Ask
them now.

They come. They buy. Their charge cards flash and twinkle. Secret codes
and holograms. Plastic purchases. A clamor for tints that match their
sofas: soft greens and golds and dusky blues. The photographer with
camera draped around her neck never poses poets. She is rich now---and
famous, recognized in every magazine.


\clearpage
\section{Poems Read Under Duress\\ (Japanese Puzzle Box)}

I look into the semi-circle and see\\
sketched on foreheads, chins, and cheeks\\
plump golden babies like cherry blossoms\\
overflowing mothers' arms,\\
green-robed fathers picking yellow mums\\
and geisha girls with downcast eyes.

So many Japanese puzzle boxes\\
painted and lacquered\\
daring me to find the crack,\\
open the box\\
without a key\\
slip out the drawer\\
hidden underneath.

I probe, poke, and pry until\\
a drawer springs open to my left\\
a lid flips up in the middle\\
and I stuff in verbs of being and doing.

Auden, Sexton, and Plath leap\\
into velvet-lined boxes;\\
neatly stacked ideas\\
slide through cracks\\
expanding the dimensions,\\
exploding the sides,\\
and class begins.


\clearpage
\section{Poets}

like plumbers aren't blinded\\
by porcelain and veneer.\\
They get to the guts\\
wrench in hand\\
twist copper screws\\
and rubber washers\\
pluck out\\
crimped hair and bits of fluff;\\
even the metal matchbox car\\
is snaked through curves\\
as cesspools drain\\
and clean voices\\
gurgle upward.


\clearpage
\section{Pond Dipper}

One drop of pond water\\
unveils the world\\
rippling on a glass slide.

My microscopic eye magnifies\\
a star-shaped colony of green\\
algae floating free.

Under fluorescent light, one\\
cell releases oxygen I inhale,\\
lifts me from my land-locked life.

I blink as crimson, emerald, yellow\\
spheres and lines paint my slide,\\
a portrait of my life I cannot refute.

Come look, my love, we can begin\\
again. Our dreams wiggle, float\\
us back in water to our other selves.

We survive as larvae, move to nymphs\\
and buzz now like dragonflies released\\
to air, hovering in our scientific world.

Let's land on a lily pad, rest\\
as buds form, burst to pink.

We need not measure days\\
in numbered slides preserved. We can\\
file them away, forgotten in metal drawers.

Let's track our childhood to the pond,\\
dip our nets in murky water,\\
scoop tadpoles into our wet hands.

We will start again as we did back then\\
while we watch with naked eyes\\
as water bugs skim the surface.


\clearpage
\section{Power Lines}

Mostly I looked up\\
tongue out, ready\\
for the slurp\\
of snow or sun,\\
the turn and churn.

All us kids knew\\
how power lines\\
could break and fizz,\\
strike like snakes. I crossed\\
under the one over Oak Street\\
for four grades and in fifth,\\
a gull's wings flapped it up,\\
up, straight up.\\
I wanted\\
the frazzle of light,\\
the burn,\\
the drop of voices falling\\
from phones, the loss of light\\
and power,\\
the dark threat\\
the gull would drop\\
but the gull was\\
some boy's white sneakers,\\
hanging from the power line.

Wind blasted those sneakers\\
dust brown, nudged them\\
a bit down the line.\\
I never saw them drop.


\clearpage
\section{Power Outage at Christ Cathedral}

before my friend's funeral\\
in the midst of zigzag fires\\
doused by thunder boom

bang and shock\\
a pronouncement of another\\
visible fire in the sky and another

the vestibule door closed out\\
a heavy oak slam soft candle\\
light, flares of sangria flames

burned down filaments of finely\\
spun fibers encased\\
in tapered white beeswax

the stained glass window\\
arched with the Ascension\\
bits of burgundy, saffron glass

blue and white sky whisper\\
my friend's whiskey breath\\
coiled invisible in my ear


\clearpage
\section{Proof of Existence}

I notified everyone\\
that you are\\
gone from our house.

I told them you don't live\\
here or anywhere anymore,\\
asked them to remove your name.

Still you get offers\\
for golf vacations, credit cards, sweepstakes\\
numbers that will change your life.

I have given up hope\\
that customer service will correct the records\\
or certify your legitimate absence from life.

You fit their demographics of a senior man\\
retired with money you haven't spent\\
on leisure activities they advertise.

You have traveled somewhere out of reach\\
and left me to deal with all your mail\\
that has no return address listed anywhere.

Envelopes slide through the slot every day,\\
spread across the slate floor, flaunt your name\\
as if I don't know what I'm talking about.


\clearpage
\section{The Psychiatrist and His Coffee Bean}

The psychiatrist is a sensitive soul who wants more than anything to
cure his patients of the poetic delusions, contusions, confusions,
illusions they confess only to him. He listens, observes, distances
himself: a scientific man sifting fact from fiction.

Nonetheless, the psychiatrist weeps at night in dreams of those who
enter his office, pay his fees faithfully as they kneel before him and
offer words and stories on silver trays like cups of coffee he drinks
black to get them through their days.

One night the psychiatrist cannot sleep, even with his down comforter
pulled to his chin, even when he recites the text of Freud and Jung and
even Skinner, word by word. He turns on the bedside light, gets up and
paces back and forth in his locked bedroom. He knows the signs of too
much caffeine: the thumping heart, unnatural alertness, restlessness,
sleeplessness. Back and forth, he paces, faster and faster. Leaping,
almost galloping, and cursing his addiction that wakes him. The coffee
bean.

The coffee bean. That's it, the psychiatrist knows the reason, the tool
of any scientific man anywhere. He brings in the tall tropical shrubs.
Frames the hypothesis. Inspired by logic, he works fast and soon over
three thousand coffee shrubs grow from the dirt he dumps on all his
floors. Everywhere. Coffee beans and more coffee beans. He fertilizes
with fish bones. All flourish in his light. Beans ripen and burst.

The psychiatrist's patients advise him not to roast or grind or drink
these coffee beans and he listens. Patients visit him at home, now,
cured of what they never had. They fill pails full of the dark brown
beans, bathe in them, sniff the pungent smell, lie down on beds of beans
and fish bones that shift beneath them.

The psychiatrist and his patients sleep through the longest nights as
poets string rosaries of coffee beans and count their musing prayers.


\clearpage
\section{Pushed by Passion}

Aunt Violet loved the smell of dirt,\\
the way it felt, the way it crumbled in her hand.\\
She said it reminded her of sex.

Aunt Violet raked her flowerbeds\\
clean of rotted leaves and fallen twigs\\
while she inhaled the fertile, woodsy odor.

On her knees, Aunt Violet planted marigold seeds\\
plucked from the frosted stems last fall\\
just before the earth turned white as her bed sheet.

Back near the corner of her blue house, Aunt Violet\\
inhaled that first bit of green: a row of slender leaves\\
in sunlight, open to catch the aqua sky and rain to come.

Aunt Violet knew how those leaves had burst from bulbs,\\
how they had licked nutrients from dirt. Those leaves\\
wiggled upward, pushed by passion to reach the light.

Aunt Violet rose from the ground, her knees brown\\
with dirt, her flesh hot and sweaty. She had waited\\
months for the vernal equinox, for that musky smell of love.


\clearpage
\section{The Puzzler}

At three, my son chewed edges\\
of the world,\\
each continent constrained\\
by the puzzle frame he carried\\
under his arm before he dumped\\
whole countries on the living\\
room floor, crossing all boundaries.

He knew the world was flat, full\\
of colors, capital cities and land\\
surrounded by blue oceans and seas\\
stamped on paper, pasted to cardboard.\\
Each piece fit\\
in its designated space unless\\
he lost an entire continent

under the love seat where mice\\
gnawed at a peninsula,\\
left teeth marks in mountain ranges.\\
He was not afraid of walking\\
off a precipice or clutching North\\
America or South in his pudgy hand\\
for he was king of where

everything belonged. He believed\\
puzzles were nothing more\\
than toys, something he could\\
figure out, put together. Even when\\
all pieces turned upside down\\
with gray backs up, he spotted\\
shapes he could flip up at will.

At seven, he spun the globe,\\
knowing he could fall off anywhere.


\clearpage
\section{Rabbit Tracks in Snow}

In the night forest there are no shadows.\\
No stars have fallen, no moon-\\
light slips past summer oaks.\\
Maples lean into each other, shake\\
love's embrace loose and kiss\\
green. Then the blaze of red-orange,\\
a final blast, an explosion of loss\\
puffs into clouds before the drop\\
into brown. Broken on fertile ground.

Even rabbits do not mate\\
in January snow crunch.\\
Ice shivers the trees,\\
brittle as young lovers\\
with hope before they break.\\
They fall through sky---\\
two lost mallard ducks shot\\
out of season.


\clearpage
\section{The Rain Goddess Speaks}

The words of the rain goddess harden,\\
cut through the night,\\
gnaw at wood above the women\\
gathered in prayer, their heads bowed.\\
Words drip down their temples,\\
into their ears, onto their lips\\
as the rain goddess speaks:

\begin{quote}
Lie down in dry riverbeds.\\
Let hazelwood cool your breasts,\\
divine the bloodpools.\\
Burrow through sediment\\
until fish rise to meet you.
\end{quote}

The armadillo woman stirs\\
cactus flowers in gin\\
and drinks sand\\
until edible prickly pears\\
sprout from her fingertips\\
and she touches peyote visions:

\begin{quote}
Hear the coyotes howl the moon\\
from your hands\\
call forth the spirits\\
to dance with you\\
on burial grounds\\
and feast on mesquite bread.
\end{quote}

The scorpion woman\\
creeps into the badlands\\
twirls in her dress of dust\\
touches shoulders, cactus cheeks\\
of men who dance on the mesa\\
rock beneath their feet.\\
She curls in trouser cuffs\\
drinks her venom\\
until she sees:

\begin{quote}
Look for the bones\\
ivory in moonlight\\
string them on grasses\\
encircle your waist\\
give birth\\
to the west wind.
\end{quote}


\clearpage
\section{Reaching Toward the Top Shelf\\
at Sam's Used Books}

Somewhere in the last row on the top shelf in the far corner\\
Sam shelved a book. Back where he never dusts cobwebs\\
form a lattice, open spaces between spider threads.\\
Gnats, long dead, dance in silver threads brushed by wind\\
when the back door opens. I stand, stretch, reach for\\
the books, red and blue and black, stacked one against one\\
tight together, upright, titles turned outward for reading.\\
It's there, I know, somewhere, with the others buried\\
by words and phrases and sepia pages waiting for my touch.


\clearpage
\section{Reconciling the Dichotomy}

In your vegetarian phase\\
you refused\\
a fried egg for breakfast\\
with its eyeball bulging\\
as if it saw you had not done\\
geometry homework due that day,\\
had not followed your ``no sex'' pledge\\
when Tommy kissed you hard\\
in the back seat of his '99 Buick.

Someone plucked that egg from its nest.\\
Not me. No, I could not steal\\
an egg warmed by hen feathers.\\
I admit I cracked the fragile shell,\\
discarded it with a flick of my wrist.\\
You could not bear such loss.

Now you meditate, seek balance\\
of all things good and right. You resist\\
the counterweight that tugs like gravity.


\clearpage
\section{The Recovery Room Nurse}

I am the white rustle\\
who whispers your name.\\
Hear me call for your return\\
from the bottom of Lake Pleasant,\\
from your bed beneath the surface\\
where currents cool your body.

The tips of my two fingers rest\\
on the inside of your wrist\\
where your pulse beats against your skin\\
and mine. I count and chart your heartbeats\\
as you kick toward air. I am afraid\\
your lungs may burst on the way up.

I will monitor your vital signs,\\
serve as your guide back to this room\\
painted sky-water blue, the color of veins.\\
I, too, have been in that watery place,\\
that womb of forgetfulness. You cannot stay\\
in that amniotic sac you abandoned years ago.

Your eyelids flutter like white seagull wings.\\
You mumble fish words that wiggle beneath blue.\\
You have escaped the seagull's sharp beak.


\clearpage
\section{Red Lava}

The fire goddess speaks in flames,\\
flicks her tongue against our cheeks\\
warm with August sun. We hear\\
blood pulse through our temples\\
as she warns of fire days to come.

\begin{quote}
Come to my subterranean cave.\\
Lie down in my chambers.\\
Cool your fevered heads.
\end{quote}

She sears that boy's sullen lips\\
with a lover's kiss as he lip synchs\\
cuss words that pound his ears, a rap\\
song full of drugs to numb his soul.

\begin{quote}
Climb down my blue ladder.\\
Rest awhile on lavender steps.\\
Beware of the fall into red lava.
\end{quote}

She blisters that woman's ringed hand\\
with her fingertips as the woman turns\\
and punches a number into her cell phone\\
and connects with an answering machine.

\begin{quote}
Fall into the circle of my arms.\\
Let flesh touch flesh as we embrace.\\
Ride my swirl of thick, red passion.
\end{quote}

She welds together that woman and that man\\
with her blowtorch hand as they stand\\
face-to-face and argue about the cost\\
of divorce, the price of severance.

\begin{quote}
Let your furies flow with mine.\\
Whip them until they boil and burn.\\
Stir our caldron of anger and jealousy.
\end{quote}

The fire goddess raises her arms\\
as we kneel at her feet. Red lava erupts,\\
bubbles hot. Underground ovens melt rock\\
and the red mass overflows all boundaries.

\begin{quote}
Dive deep into your passion.\\
Touch bottom with your feet.\\
Brand your soles with my mark.
\end{quote}

We ride the crest of the lava flow,\\
grow giddy with heat as all burns\\
and cools to black rock. From cracks\\
in barren ground, we rise like smoke.


\clearpage
\section{Refrigerator Note}

If you were magnetized, you too\\
would stick to metal doors.\\
Spread-eagled. Stretched between

magnetic poles, positive\\
and negative pulls. Two frigid points\\
opposite, both arctic circles. Both\\
yank and tug.\\
Your soul cracked,\\
two halves of a peach

pit revealed. The very core,\\
the seed that grows like the note\\
you left on the refrigerator. Last night,

you ran out. Out into early October snow. Your way\\
illuminated by the hunter's moon full\\
of footprints that puffed up the balloon\\
above. Your tracks imprinted\\
deep with frozen edges, sharp\\
yet smooth as your green rubber boots

I could track. I could read\\
the signs along your trapline,\\
plain as your note\\
that I shredded, scattered

throughout the house. Words\\
clung to my bare feet. Last night\\
my callused soles bled\\
a trail of phrases\\
across emerald wool carpet.

In corners\\
of our rooms, your words\\
took root, germinated\\
into full sentences.

I erased vowels, soft sounds\\
round as the moon's phases. Even\\
then the curved moon cradled your consonants.

Your words rose up from paper\\
like so many tulips. Old wounds opened\\
in the day of light. Your voice\\
vibrated against walls.\\
Metal to metal.

Only the echo of you remained. You\\
who left last night, who would not\\
listen when I called your name.

I rolled your sound\\
around my tongue, savored its taste\\
when it burst, sweet\\
as a single Concord grape. You

who are no more now than a stain\\
on the refrigerator door.


\clearpage
\section{Refusing to Say Grace}

Supper hour slips through her fingers\\
slick as rosary beads. She counts\\
husband, son and daughter,\\
mouths open. Tongues ladle words,\\
dip deep into days,\\
scoop up scenes, serve\\
chewy as meatballs.\\
Well done

as if they do\\
know how\\
bread crusts crumble\\
in dry air. They swallow\\
water, flush body cavities\\
clean as the tablecloth\\
washed by hand and spread\\
before them.

Not the knife, the fork,\\
spoon. They lie\\
still, stainless,\\
their future forged\\
in heat. She sweats

opening ovens, stirring pots\\
boiling over onto electric burners\\
where bits of food flame to ash.

Dab ash on foreheads, shape the cross.\\
Their bodies, each a mass;\\
arms extend, reach\\
for rolls or meat or bowls\\
full of succotash, specked\\
with hot peppers, red\\
as a reminder. Yet,

her son will not say\\
grace or bow his head\\
for anyone or anything,\\
not even for her, anymore.

He believes\\
he dies\\
in dreams\\
every night and wakes

early, trusting only\\
what he bites off, crushes\\
between back molars\\
for sustenance.


\clearpage
\section{Rehearsal}

If I had two months of life left

would I sprout and send down roots\\
deeper and deeper in search\\
of water and divine\\
the blue, cool voice of wetness

or would I just float paper boats\\
in muddy sidewalk puddles\\
where sparrows dip their beaks?

I could assign tasks\\
of this and that\\
so friends distract themselves\\
by clipping cotton strings\\
from this rag of life\\
wrung dry of their tears.

Could I find something better\\
than driving the freeway\\
the same route every day\\
which ends when I slam the car door\\
and elevate to the 15\textsuperscript{th} floor\\
where my fingers dawdle over keys\\
reluctant before turning locks\\
needing oil. Still\\
jagged metal fits jagged metal.

I shall choose. I shall\\
scratch in cracked, brown clay.\\
With dust under my scarlet nails,\\
I shall make my mark\\
and brace for wind and water

in my garden until my soul\\
balloons from my toes and fingertips\\
and blooms a meadow of blue chicory\\
and Queen Anne's Lace\\
before scattering its seeds.


\clearpage
\section{Relativity of Roots}

There is no history of me\\
who came late to this family,\\
asking questions, begging\\
names of those who came before\\
who had not claimed their spaces\\
in empty boxes, arranged symmetrically\\
on both sides of my family tree.

The full moon of the falling leaves has come\\
and gone. I enter the sweet time\\
in this woodland where the sap runs\\
in the sugar maple trees, where strawberries ripen,\\
stain the earth red with blood,\\
a reminder of the strings, the binding\\
of all living things, the cycle of my life.

My path records the tracks of those who walk\\
the twisting trail, circling ever inward\\
to the center where we all have been\\
when we discover who we are\\
and out again to where\\
strings dangle in black space\\
as if no one knows the end.

The world cracks in half but no one hears\\
the breaking. The chasm is the proof.\\
I understand the buoyancy of emptiness.\\
my history is yours. All we need\\
are clouds of eagle feathers\\
roots thick as string\\
one green, green thread.


\clearpage
\section{Reluctant Eyewitness in the Florida Keys}

I will testify in any court,\\
tell the truth about what I've seen.

Your black eyes pin me in my place\\
like a moth you study in a locked case\\
where my arms flap against glass.

Your tented lids droop over green eyes\\
that recognize women do what they must\\
when the undertow drags them down.

You hide your eyes behind glass and skin\\
as if you can not see that I might be\\
the mother who killed her children.

Your fearful eyes fill with blue-green ocean\\
yank me in where I can drown\\
unless I take a breath and swim to rocky ground.

Your eyes smile and turn my image back to me\\
so I can see my innocence rise into the air\\
on the backs of seagulls, white feathers everywhere.

You are the madman with the onyx eyes\\
of night who labels me unworthy\\
of living in this day of light.

you do not suspect\\
I have drawn your many eyes\\
for evidence. I carry you with me\\
when I go\\
just as you have\\
etched some part of me\\
behind your eyelids\\
for safekeeping.


\clearpage
\section{The Remains of Living Things}

I have come to find the dinosaurs\\
lost under all this desert sand,\\
nestled in bedrock along the arroyo.

I hope for an egg: unbroken,\\
fossilized. A life interrupted\\
could fill my right hand.

If I could turn rock to flesh, I would\\
birth one green dinosaur\\
who would shelter with me under the Joshua tree.

Here we would weave baskets from yucca leaves,\\
sing ``whit-wheet'' with the curve-billed thrasher,\\
wish the coming and going were not permanent.

I, too, have paced this arid land,\\
heard the dark wings of the nighthawk\\
at dawn, grabbed for one black feather.


\clearpage
\section{Remains of Summer}

Brown crumple of what was once\\
red zinnia petals now crinkle\\
crunch under the gardener's boots\\
already coated with mud\\
zapped by lightening in last night's storm.

In her mind's eye, the gardener conjures up\\
a wisp of smoke, curves it\\
into her lover who left before the first frost\\
crystals froze the cutting garden\\
into shriveled stalks and dead flower heads\\
bent over puddles of fallen leaves.

She rakes her raggle-taggle garden,\\
hacks at clods of clay,\\
buries what remains.


\clearpage
\section{Remedies}

Indian summer finally comes to the mountains,\\
heralded by rain crows reading the sky's palm.\\
(Three days' sun, one rain.)\\
Their gypsy throats swell with sound\\
like her hands, large with knowing\\
tomatoes ripe on the vine, the midwife secrets.

The medicine man will come,\\
she knows,\\
bringing spring tonics in autumn,\\
drinking pokeberry wine\\
against winter's rheumatism\\
saving some for her\\
whose knotted hands reach inward\\
to count the rings,\\
set the time of coming and going.

Her healing herbs dwindle in Mason jars,\\
sons and daughters no longer hunt roots\\
or white-painted salvation at Calvary\\
where voices rise in revelation.

\begin{quote}
Cornfields lie fallow,\\
barn planks rot,\\
no one eats hominy\\
or plays gourd banjos\\
anymore.\\
But, once, wild roses circled hemlocks.\\
Comfrey roots sucked out chest pain.\\
Pennyroyal banished night chills\\
before the harvest dance\\
where her hair unbraided\\
and she walked with stars.
\end{quote}

Her hooded eyes watch for the medicine man\\
to find the path,\\
take her hand in his,\\
spoon catnip tea between her lips\\
until she sleeps.


\clearpage
\section{Reunion Sunday\\
  at the Methodist Episcopal Church}

\emph{Keweenaw Peninsula, Michigan, July 2005}

My great-grandfather's bones remain\\
at the bottom of the deepest shaft,\\
in the farthest tunnel where\\
Cornish miners once chiseled copper\\
from the black rock walls\\
of the Victoria Mine.

I have come to remember him,\\
a man I never met,\\
and other miners\\
who once walked beneath the earth\\
in hobnail boots with worn soles.

As the preacher tells\\
their stories, I listen\\
for the cadence of their lives,\\
the lilt of their picks tap-tapping\\
and the slam of hammer heads\\
breaking up copper chunks\\
that fall to the dirt like so many\\
renegade orange-red suns.

My great-grandfather lost his way,\\
some say, when he traced the purest,\\
the richest copper veins to their source,\\
a place where other men had not been.

Others claim he chose\\
the underground where fires\\
fueled his rage\\
when companies walked away\\
from the open mouths\\
of mines and bent men\\
and all foundations tumbled\\
onto hard, hard ground.

On the last Sunday in July,\\
I listen for the silence\\
of my great-grandfather's pick.


\clearpage
\section{Requebra}

The drumbeat thumps in our chests,\\
a second heart pumping rhythm\\
through arteries no longer clogged\\
as we march, black heads held up,\\
bodies twitching to the beat\\
of four hundred feet that move\\
as one in the Bahian Carnival.

Hear the cotton-candy melody,\\
the sweet sound as bodies rub together,\\
skin to skin, mouth to mouth\\
breathing in and breathing out\\
the mating call. Move your herky-\\
jerky butt in tune, grind your body\\
into mine. Strip down to basic chords.\\
The drums, guitars and keyboards hit notes\\
of red passion for who we are,\\
of gold glitter reflecting sun,\\
of green grass where life is good\\
through night as black bodies merge.

Requebra vibrates from dawn to dusk,\\
our song of celebration, our battle call.


\clearpage
\section{The Rio Grande Gorge in Summer}

This is the season of suicide\\
when young women leave their men\\
alone without any offering, without\\
even a note scribbled on lavender\\
paper. It is their words\\
echoed by the owl: who, who\\
who will find the open space\\
tucked between red rock lips?

Who has seen the stepping off,\\
heard the pebbles crumble? Not you\\
who reads the story twice. No\\
you step back from that ledge\\
from the gap-toothed grin of men\\
who have nothing to lose, not\\
even you. You would have left\\
a letter with scratched-out words.


\clearpage
\section{The River of Us}

Gaze long enough at me\\
You see yourself. You\\
Lead me to blue\\
Where pieces of sky\\
Plummet into Stony Creek.

Your arms are bare. Blue rivulets\\
Flow under your skin, warm currents.\\
You dip my finger into blue\\
And I touch liquid green.\\
We reach for white, those strands\\
Clouds leave behind\\
When they rise moist.

Look there. That heart-rock\\
Plump with age-old mystery.\\
All those river rocks in bed\\
Lulled by dragonfly buzz.\\
We stroke lichen sun drops,\\
That line of green-yellow fur\\
On the blue rock road.

Come let us soak\\
In wandering waterways,\\
Our toes gripping river rock\\
As all that wet murmurs\\
The ancient blue-green song.\\
Who we are.\\
Who we will be.


\clearpage
\section{Rock and Roll Never Forgets}

``I want to go to a Bob Seger concert.''\\
\emph{First words of Evie, a 79-year-old woman when she\\
came out of a five-year coma (from} Michigan Live, \emph{4/13/2013)}

Two baby bluebirds\\
burst from Evie's throat\\
last week. No one saw\\
their wings spread, carrying\\
globs of black tar, gasps of nothingness,\\
round neon balls, green flashes\\
of regret. Those blue feathers\\
zig-zagged past her,\\
over her sleeping aide and nothing\\
changed. Machines continued beeping.

For days, Evie scuttled across\\
the ocean floor, a crab without a shell.\\
The slush of sand shifted, tainted the silence,\\
grains nestled under her big toe. Ruffled fins\\
of orange fish tickled her neck. She stayed stuck\\
with the guitarfish wiggling over mud flats,\\
volcanoes biding time before letting loose.\\
The sea urchins waited to strike\\
with their poisonous spines. A whole world\\
she had never known. Here she breathed\\
in and out without drowning. She belonged\\
with the speechless. All living\\
creatures made space for her and\\
water held her back.

Her squinched eye spotted an RCA red plastic\\
radio on the bedside table next to a wrapped straw\\
and the cord stuck into a wall plug somewhere.\\
Her parents gave her that radio\\
when it was new and she had kept it, listened\\
to the music, especially the oldies, the rock and\\
roll of her youth. Her voice had not whistled\\
with Little Richard or Elvis for years.\\
The drone of news and no news\\
plugged her ears. She had not wanted\\
the ugliness in the world marching in.\\
Her aide turned the dial to rock `n roll.\\
Evie heard an explosion, a rush of red.\\
Blood roiled inside her like an ocean.\\
She raised herself to her elbows.\\
Bob Seger lyrics screamed,\\
``If you need a fix\\
come back baby.''


\clearpage
\section{Rock Crushes Scissors}

You always knew I would choose\\
paper, fingers splayed,\\
right hand turned down. My loss\\
as your two fingers opened\\
and closed again and again,\\
scissors cutting me in two,\\
making me go last\\
in every children's game\\
we played when we were young.

In class, my pencil swished across\\
paper. Words dipped and rose above,\\
below and on the lines of my messages\\
to you who never read or heard what\\
I said as if my hand in air spoke\\
of all I was, all I could be.

You ran with scissors, never\\
afraid of falling\\
under fist-sized rocks\\
that could poke out your eye\\
or bloody your nose for you were\\
you and you were always first.

Over the years, you honed your skills\\
into fine points,\\
yelled sharp-scissored words\\
at me, now your wife\\
during this cutting time.\\
You were promoted at work,\\
yelling all the way at everyone\\
who disagreed and you got things done\\
but not enough. You climbed faster.

You clenched your fist, a wrinkled rock.\\
With your rock, you crushed\\
scissors and people who worked for you\\
as your company went bankrupt.\\
White-knuckled, you fought the balance\\
of audits, of lawsuits, of me.

Paper was more malleable than rock.\\
You heard the crinkle as it wrinkled\\
into ball, folded into swan, floated\\
on air currents, settled to ground.\\
My paper covered your rock.


\clearpage
\section{The Roto-Rooter Guy\\
 Snakes Out Two Alligators}

Sarah next door tells\\
Mary across the street\\
and she hollers\\
to Tom on the corner\\
who warns me\\
since I'm on my own\\
on the hottest day in June\\
on this street of married couples.\\
As I run with the rest\\
to Bobbie Jean's house,\\
sweat pools between\\
my breasts as they bobble\\
and sway and settle\\
only when I stop and stare\\
dumbfounded as the Roto-\\
Rooter guy sucks up\\
the second albino alligator,\\
drops it onto the brittle brown grass\\
while the first alligator opens its jaws,\\
yawns and rolls its red eyes\\
at our squeals, our yelps\\
when we jump back,\\
snap photos on our cell\\
phones and punch in TV news tip\\
numbers. All of us\\
want to tell our story in headlines.\\
We dream of our brave faces\\
plastered across plasma TV screens\\
in every living room, on every street\\
in this nation. We want\\
viewers to frantically flush their toilets,\\
search wash tubs and yank up sewer covers\\
as they hunt albino alligators everywhere.


\clearpage
\section{Round Sounds}

I'm squeezed somewhere between here and yesterday\\
with a mother's pinched lips and dilated pupils.\\
Standing in the bathroom, I\\
brush my teeth\\
to stop the screams --\\
round sounds magnified by the mirror\\
as I pluck my eyebrows and draw the lines.\\
Filling in spaces, I\\
define shapes\\
to make my face,\\
a connect-the-dots picture in a coloring book\\
that my daughter has not yet completed.\\
Clutching the crayon, I\\
press hard,\\
retrace the lines.

But she's drawn her own circle-lines\\
with blues and yellows on cheap manilla paper.\\
She resists my attempts\\
to add her dots\\
completing my self.


\clearpage
\section{Rub Out}

``Bring me the floor plan,'' Gina's boss shrieks, ``and all the gum
erasers you can find. Right now.''

``Call me despot of the high-rise, will he? That puny accountant.'' A
flick of an eraser and his name disappears. ``I own this building.''

``Don't,'' Gina pleads. She loves that puny guy who is no more.

``This guy, too,'' the boss chuckles, rubbing holes in the plan. ``All
the men gone.'' The boss points blood-red, tapered nails at the
survivors' names: all females.

Gina dumps the empty eraser box. She draws the eraser hidden in her bra.

Rub-a-dub, her boss's name is gone.


\clearpage
\section{Run, Brotha, Run}

Friend been snitching\\
to the big guys\\
how you study them books,\\
how you ace them tests\\
too hard to pass\\
and most of all\\
how you stays to yo'self.\\
Yank on those Nikes.\\
Run, Brotha, run.

Baby sista past all savin'\\
now she done been sniffin'\\
stuff up her nose, payin'\\
for it and buyin' time\\
walkin' Woodward Avenue\\
while her no-good man's been\\
beatin' heads and he's a lookin'\\
for you who is nowhere here.\\
Run, Brotha, run.

Ain't no salvation at Calvary\\
or nowhere else in this dead\\
city jammed with emptiness\\
while all good mens outta jobs\\
and down right, downtown mean\\
as pit bulls rarin' for a fight\\
every night. Just put yo' money\\
in the man's hand playin' the game.\\
Run, Brotha, run.

Hey, Brotha, Somewhere out there\\
you be safe as Mama plots escape\\
from windows, roofs, summer\\
heat you be carryin' all this time\\
on yo' back 'cause you, too, dream\\
\clearpage
Louisiana days, long and sweet\\
hours of eatin' crawfish, laughin'\\
at fools don't know no better.\\
Run, Brotha, run for your life.


\clearpage
\section{Sally Heads for the Night}

At rush hour, Sally high-steps off the curb,\\
glides between the Fords and Chevies\\
taps the hood of a pick-up truck\\
as cars honk. A Sears delivery van rolls\\
closer before the light turns the ghastly\\
green of go ahead and run her over.

Machines stampede,\\
a herd of chromed racers.\\
Nowhere to go but over her,\\
a woman in high spirits\\
dressed in black, ready for a starry night.

Sally bows, waves, pirouettes, throws\\
kisses, promises more than she can deliver\\
in her world of dangerous cars and high hopes.


\clearpage
\section{Searching for an Entrance to a Woods}

I mistake you for my lover who grew roots\\
in my garden, stood with me those long hours\\
during the drought year. Even I shriveled.\\
My skin dried to dust as your leaves curled\\
and fell for lack of rain, lack of green\\
plump dreams swaying in the wind. I hear\\
the rustle of death on gray concrete streets\\
where my lover scraped his soul one dark night\\
and abandoned me in sunless days.

You stand contorted, your twisted limbs\\
fall apart at the edges. I lose\\
pieces of myself. I could break off a twig,\\
light your firewood, listen to the crackle,\\
the sizzle of the burn as flames snap at me,\\
a reminder of the passion, those woody days\\
when I stumbled into a hidden passage,\\
into the green forest where my life began.

I discover another way into where I have been\\
when I wore Queen Anne's Lace in my hair\\
and held a fistful of forget-me-nots.\\
Now the sparrow hawk circles above me,\\
waiting for me to drop on this barren ground\\
where others have chopped down my trees,\\
left ragged stumps shaped like humans\\
that raise their arms in agony and protest\\
the loss. I bind twigs in blue sisal, hope\\
the center will hold against greed.


\clearpage
\section{The Season of Fireworks and Naked Trees}

Fireworks blossom into blue\\
and red-gold peonies\\
floating in the burst of a moment\\
on aqua ripples\\
contained in the cut-crystal bowl of sky\\
above the lovers' heads.

Sizzle, whiz, crack and whistle\\
stir the night air, whip the single stars\\
into new constellations, into tails of froth\\
these lovers grab and climb\\
hand-over-hand through smoke\\
and ashes, into the orange explosion.

Such lovers have seen their own birth\\
and death, the black-stick trees\\
that time will tip eventually.\\
The darkened sky drops\\
and the gods smile at the magnificence\\
of love, of naked trees gone to ground.


\clearpage
\section{The Season of the Hummingbird}

\emph{For Bradley}

Cup your small hand in mine.\\
I will fill it with words\\
unspoken, unbroken\\
as if some ruby-throated hummingbird\\
plucked them like thistle,\\
like dandelion down, soft\\
enough to line its nest\\
woven with spider silk, a pattern\\
as delicate as your lifetime.

Your tongue moves sounds around\\
darting up, down -- testing\\
their taste like the hummingbird\\
dipping its long bill into scarlet\\
sage, into the nectar of this herb\\
as its wings beat the humid summer air.\\
You hum the same rapid tune\\
when feathers brush your lips,\\
when your fingertips quiver\\
as if you knew the song.

I hold you in my palm, warm\\
with summer, with wanting\\
all there is. You speak to me\\
in the season of the hummingbird\\
when the Rose of Sharon blooms.

Tonight the moon rises\\
earlier than yesterday.\\
Its white light illumines your hair\\
drops a yellow cap on your head\\
as if pollen had settled there.


\clearpage
\section{Secret Garden}

Neighbors watch\\
your hybrid roses rot\\
in rain. Weeds flourish.\\
Each rose bush strangles\\
on Virginia creeper. Vines\\
take over, cover everything.\\
Rosebuds wither on broken stems.\\
Fungus polka dots green leaves with white.\\
Aphids thrive, suck out the life juice.\\
The thorns still draw blood.

Only you can mulch the beds,\\
coax roses to bloom again.\\
Your gardening gloves hang\\
from the hook by the door\\
in the shed. There is a hole\\
in the thumb of the right hand.\\
Your trowel rusted over winter,\\
an orange coating of neglect.\\
The bottle of fungicide is more than\\
half empty. I hear fluid swish inside\\
when I shake it as if there is hope,\\
some medicine you can take\\
and cure yourself of this awful disease\\
we both battle daily as if it were a rose.

Come with me now into my secret garden.\\
Slip your hand in mine as we stroll.\\
Each day from sunrise to sunset,\\
I collect daylilies that bloom one day\\
and then another blooms. I cut the stems,\\
gather whole bouquets of these perennials.\\
They grow on leafless stalks,\\
with bud after bud opening and closing.\\
Arrange them with me, the red, the orange,\\
yellow and mahogany. Each trumpets a day\\
we had together, a memory as sharp and smooth\\
as the strappy leaves that bind us.\\
I stitch them on quilts, capture a window of sky,\\
a square of dirt where they once grew\\
and spread naturally. I set daylilies everywhere.\\
We call on their medicinal power: a relief\\
from pain and a lessening of grief\\
as we walk toward the tallest pine\\
in my secret garden where no rose survives.


\clearpage
\section{Sexual Division}

The man grew breasts.\\
Straps bit his shoulder flesh\\
like a cur's enameled fangs\\
sinking past curly chest hair,\\
cutting through bones\\
right to the heart\\
where the man-woman\\
flexed muscles,\\
dried tears\\
for what he was before

she came\\
into his house,\\
dusted untouched corners\\
with feather brooms,\\
scattered pain everywhere,\\
polluted his rivers\\
with night fears and blood.\\
He knew, intuitively,\\
his Easter soul\\
would embrace her\\
halfway.


\clearpage
\section{She Dreams}

She dreams rooms filled with books.\\
Dr. Seuss in corners, yaks climb the wall.\\
Betsy meets the Bobsey twins and Huckleberry Finn.\\
Dog-eared books flow from faucets,\\
pages turn themselves.\\
Nancy Drew mysteries are mountain ranges.\\
Superwoman stands in snow, reciting Tennyson.\\
The Pigman eats mean hamburgers\\
on a table of paperbacks.\\
He drinks sentences and paragraphs.\\
Words dive from leather boards,\\
swim in her eyes, until she dries them,\\
wraps them in silk,\\
and pins them in her hair.


\clearpage
\section{She Makes Her Bed}

All night the rain falls.\\
Oak limbs bend,\\
scratch at the glass\\
of her window. Each wet drop\\
wiggles down the pane,\\
blown by wind,\\
pulled by gravity\\
into cursive lines of memories\\
she slips over her head\\
like a well-worn nightgown.

Tonight, the crazy quilt\\
she sewed holds her down.\\
She fingers each piece.\\
The blue bowling shirt triangle\\
blazes with her husband's number,\\
holds the smoke he chose.\\
She cut up his jeans worn thin\\
on the thigh where he wiped\\
his palms when he got mad,\\
before he walked out the door once\\
and for all. Her daughter's graduation\\
gown gaped back at her, its black\\
hole holding the fights, the late nights,\\
the boyfriends who hid outside\\
taking her where her mother\\
did not know and could not go.

The antique ivory silk with lace\\
swished at her wedding so long ago.\\
She can't even vouch for what she vowed,\\
only how she clutched a bouquet\\
of pink larkspur she plucked\\
from the meadow, a wild flower\\
that attracted her and butterflies.\\
The honey bees buzzed danger.

In the morning, she makes her bed\\
as she does every day. Her muscles remember.\\
She flutters and smooths the bottom sheet,\\
then yanks the corners tight and folds.\\
The top sheet billows before she turns\\
it down and pushes it in all around.\\
The crazy quilt patterns itself on top,\\
edges loose. She will slip\\
between the sheets tonight as if she were water.


\clearpage
\section{Shoe Store Romeo}

She only bought high\\
heels at night\\
after the shoe salesman\\
cupped her ankle,\\
let her nyloned flesh settle,\\
seek the warmth of his palm\\
for a second,\\
a flash of a moment\\
nothing more,\\
nobody could accuse him\\
of anything.

Still on one knee,\\
he stared into her blue eyes\\
as he slipped on the shoe\\
with the highest heel\\
and rolled his hand\\
lightly up her shin\\
and back again,\\
just a gesture\\
lifting his white cuff\\
above his tattoo,\\
a small black skull.

The swipe of her blonde hair\\
in the air, the sink of heels\\
in beige carpet, the look-back\\
smile, the hiss of fabric against shoe,\\
the rustle of tissue, of shoe\\
against shoe, of cardboard on top\\
against cardboard bottom\\
against bag,\\
against flesh.\\
The worn and wrinkled bills\\
in this high-heeled transaction.

In school, she daydreamed\\
novels of him and her.


\clearpage
\section{The Shopping Bag Lady}

Coiled in the doorway,\\
waiting for the flutist to cover the holes,\\
charm the noises from the steel tube.

\begin{quote}
City sounds assault her.\\
Woeful wails ride the autumn breeze,\\
bang against rotted wooden walls,\\
fall through broken windowpanes,\\
explode into screams.
\end{quote}

She hears the flute and unwinds slowly.\\
Clear chants in the Arabian night,\\
slivers of sound,\\
silver crystals in the darkness,\\
a litany of sweet notes.

\begin{quote}
Street people pass the old lady\\
huddled in the corner\\
with a brown, torn shopping bag,\\
its handles wrapped\\
around her wrist.
\end{quote}

She moves,\\
rippling skin in tune,\\
sensuous in the cold morning light.\\
Musky smells ooze onto desert sands\\
as she sheds fabric layers\\
from her mind and body,\\
hearing, remembering soft voices,\\
hands throwing golden coins\\
when the flutist played\\
and she danced too long.

\begin{quote}
Pulling down the felt hat,\\
she drops her soiled chin\\
into her sunken chest,\\
coughs quietly into her bag,\\
\clearpage
rummages for a cigarette stub\\
found on yesterday's sidewalk,\\
next to the garbage can\\
where she eats and talks.
\end{quote}


\clearpage
\section{Sign of the Times}

The company's name crumbles\\
as rusted paint chips flake\\
from the green metal sign,\\
dribble to the sidewalk.

A fleck of red, part of the first\\
letter of the owner's name\\
sticks to a stranger's crepe-soled shoe\\
as if he can walk it someplace safe.

Consonants disappear from the beginning\\
and end as rain licks the hard sounds.\\
Vowels stick, the last\\
remnants of the owner's howl.


\clearpage
\section{The Sky is Falling}

The flap of skin under Mrs. Little's chin\\
waggled as she warned our third grade class\\
about bombs and flash burns during\\
our monthly air-raid drills while we stood behind\\
cement pillars in the school basement with others\\
whose legs trembled and eyes scrunched as if\\
they might cry or sink to the ground from the weight\\
of the whole school on top of them.

Back in class,\\
when I messed up the multiplication tables\\
and erased numbers so hard the paper ripped,\\
Mrs. Little rapped my head\\
hard with her sharp fingernail, polished\\
blood red. Just my whisper at her back\\
whipped her around. She aimed\\
the eraser at my head. White chalk\\
fluffed the air, settled in my hair.

Mrs. Little drilled us in disasters:\\
how fast Newton's apple fell from the tree\\
of knowledge, how the sky might freeze\\
and break into silver hatchets\\
that could split a head in half.\\
A crow could plunge through air,\\
snip off a bit of our years. She could\\
hold us chickens back a year.

I pinned the sky in place with my eyes,\\
opened my arms wide to whatever might fall.


\clearpage
\section{Skylight}

Today we snack on sky,\\
pinch off a bit\\
of its blue hem\\
and chew a piece\\
of air. All that holds\\
us here. We spot

a woman in black taffeta,\\
her gown shredding against\\
concrete. Back and forth\\
she paces, against night, against\\
the hiss and swish of navy\\
anonymity as sky canopies her turf

just as sky jackets our bodies.\\
Hems unravel up and down the avenue\\
like lines on a map, marking\\
major freeways, route numbers,\\
constellations on the ground. Sky turns\\
upside down. We follow threads, not knowing

which strand dead ends at curbside,\\
which leads to our blue-roofed home.


\clearpage
\section{Sleepless}

This day perches heavy on my head\\
like some green parakeet repeating again\\
and again words that spiral in my ear\\
and screech until I cannot sleep.

Such trivial notions harden into seeds\\
that bird plucks from air and crunches in its beak.\\
It feeds on my misfortune. It sharpens every claw\\
on my mistakes. I turn onto my left side, away from it.

My eyes remain open, dry as this Torrid Zone\\
where I reside tonight without relief.\\
Only this nervous bird keeps me company\\
as it pecks at the chalky shell of a cuttlebone.

I need my sleep, my dreams of flying straight.\\
I roll to my right side, away from the hallway light,\\
away from the green, green bird.\\
It wants to spread its wings in flight.

This parakeet belongs to me. No one else will own it.\\
It grooms its green feathers. I plump my pillow\\
and burrow in as if I can escape\\
the sound of feathers when they fall.


\clearpage
\section{Sleight of Hand}

A swoosh of your hand\\
spreads cards\\
face up,\\
yours and mine:\\
pale squares

full of uncut diamonds. All\\
red numbers. Add, subtract,\\
divide and multiply. Even\\
odd numbers come up black\\
sometimes. Repeat themselves\\
if you count

cards. Too many\\
duplicates drift\\
into your trouser cuff\\
lined with hearts

you claim as trump.\\
You fan them\\
in your hand,\\
collapse them\\
into a slender neck

that fits your fist\\
full of cardboard, corners\\
marked and bent. A flick

of your fingers discards\\
the lowest heart, the two.

A single curve.\\
Thick as our beginning.\\
Thin as our final end.


\clearpage
\section{Snow Carnival}

All day weather advisories\\
skitter across the bottom of the television screen:\\
\emph{Winter storm\\
Stay off the roads}

Aunt Jane stumbles\\
over a bridge of frozen slush\\
snowplows tossed up at the bottom\\
of the driveway she just shoveled.\\
Josh, only seven, believes in days\\
off school and the fort he builds\\
against cold winds, a silent place\\
where he eats his lunch\\
and pretends he alone\\
survives this snow white day.

Cars, trucks, delivery vans skid, slip-slide,\\
smash into someone else's fender\\
like bumper cars. Like a Tilt-a-Whirl\\
these vehicles fall to earth. Like snowflakes\\
fall to earth\\
one ice crystal at a time,

I snuggle into the top seat\\
on the Ferris wheel\\
where I hoard bits of snow\\
in my red-gloved hand,\\
where this shimmer of white glitter\\
glints its dazzle of destruction\\
and I marvel at the beauty of it all.


\clearpage
\section{Soaring}

Inside the House of Fabrics\\
I find myself shuffling up and down\\
aisles of cotton, wool, lace, satin\\
wrapped on bolts. Like me they wait\\
for the first cut of the scissors,\\
for what they will become.

I brush against purple silk,\\
more like violet with lavender designs\\
in a language I don't know. It speaks\\
of Japanese simplicity, of silk worms\\
spinning this web around me, pinning me\\
in place so I don't fall down.

Although I have no pattern, no thought\\
of what I will make, I buy three yards\\
and rush to work no longer afraid. I can\\
not stop death of the old people I assist\\
in living. They swallow medicine, forget\\
who they are. Still, they keep moving.

These delicate folks want to know what\\
I carry from the outside world. I tell\\
about the silk, hold it this way and that\\
as light shimmers, turns lavender to blue\\
and back again. Silk rustles. One woman\\
exclaims she holds twilight in her hand.

On my knees, I cut my silk into a kite.\\
I gather four fallen twigs from outside,\\
fasten silk edges tight at each corner. I find a ball of twine as we go
together\\
outside where the wind swirls. We each hold\\
a string as the kite catches a wind current.

Our kite rides the waves of pink-tinged sky;\\
soars, glides, rises, trails a blue silk tail,\\
chants some sweet language of clouds and air.\\
\clearpage
We cut the strings\\
at dusk when colors fade,\\
before night envelops us.


\clearpage
\section{Some Pinecones}

I listen to crunch.\\
My boots squash paper\\
cups, one tin can,\\
three bottle caps\\
and oak leaves in the fall.

I hear it all\\
as I walk. Murmur of voices\\
screened by mesh, a shout\\
metallic with gunshot fear. A dog\\
yaps, wind shakes maple\\
leaves that cling to wood. Every\\
star clusters above me. Around\\
the moon and over roofs, stars shoot,\\
burn hot, drop at my feet. One\\
wish I'm granted in all this noise.

Under a streetlight, scotch pines needle\\
me, rub branch against branch. I am\\
alone on suburban streets. My subdivision\\
waits a mile up. I know its name,\\
the house with my number, the place\\
I bring back what I have found.

I set together\\
two pinecones waxed with moonlight,\\
stuffed with seeds.\\
One balances upright against darkness.\\
One tilts on its side, damaged.\\
One slender twig cradles them.

Two broken scales. Love spirals\\
out of control. The red squirrel feeds.


\clearpage
\section{Somewhere in New England}

Above the old woman's head\\
squirrels fly at night,\\
legs spread.\\
Folds of skin catch the wind\\
and hold the woman's song,\\
carry it across the spaces\\
from limb to limb,\\
turn it in forepaws.

The high note's a clasp,\\
the words, a chain,\\
a scrimshaw to hang on the fir.\\
The mast is rigged.

She sails on salt visions:\\
men larger than waves,\\
whalebone in moonlight.\\
Her hair flows like seaweed\\
as she anchors and waits\\
for her man\\
to rise from the wet,\\
push back the storm,\\
walk to the rocks\\
where she sits now\\
and harvests stonecrop.


\clearpage
\section{Squirrel Control}

A flash of red-gray fur.\\
The swish of a bushy tail.\\
Scritch-scratch. Claws scrape\\
glass behind the fireplace door.

A black-eyed rodent glares\\
as I stare back, watch\\
four paws squirl the pile of ash,\\
all that's left of last night's fire.

I am a city girl afraid of sharp teeth\\
and squirrels that slip-slide down\\
my chimney, spray my home with chitter-chatter\\
fear of dark red blood or some stark brain disease.

I punch in the number of Critter Control.\\
The man comes right out. He chuckles as\\
the squirrel triggers the ash dump spring,\\
tumbles fast into a black-soot void.

The man tempts the hungry creature\\
with peanut butter bait. I must wait\\
until late tonight, strain to hear\\
the metallic clink as the trap door closes.

All night I expect squirrel teeth at my neck,\\
a furry tail blindfold, a smudge of ash on my forehead,\\
the burst of a flame in my brain, the loss\\
of hope as the critter skitters over me.

In the peanut butter morning, the squirrel licks\\
his paws in the cage. With gloves on, I shake\\
my invader, exile it to the back porch where Critter\\
Control will snatch it, release it in green fields.

Later I visit my father in the common area\\
of the Alzheimer's Ward. Squirrels flit in his mind,\\
a twittering taunt. He recalls his Ohio farm days,\\
the sharp retort of his shotgun in humid summer air.


\clearpage
\section{Starling in My Fireplace}

Mystic carrier of messages,\\
diviner of star talk and dark spirits,\\
you fold your black iridescence\\
close against your body,\\
drop past charred chimney bricks,\\
land on my fireplace floor\\
where you scratch grey ash,\\
claw prophesies in squiggles\\
while night shadow clings in corners\\
and even the television remains silent.

Bundle of feathers,\\
curved bill of truth,\\
you peck a secret code against the glass\\
between us. You can snap a beetle's\\
back in your beak. The flurry of your wings\\
frightens me as they flap\\
soot from the chimney shaft.

With harsh gurgles and clicks,\\
you call your flock form the evergreens\\
as I cover mirrors and windows with sheets.\\
You with your purple-green hood\\
slick with oil spout a litany of black threats:\\
famines, fires, ice, hurricanes, plagues,\\
volcanoes, wars, guns, what rises\\
from the earth, what falls from the sky\\
and all that turns in between.

Bird of the occult,\\
friend of the ancients,\\
you would carry me on your back,\\
translate the tongues you speak\\
until I pull open the glass\\
and you glide past my oak door,\\
into the light stream speckled with dust.


\clearpage
\section{The Statue of Liberty Wears Next to Nothing}

In the land of the free, men may lose\\
their shirts. Pink bunnies stand upright,\\
dangle baskets crammed with yellow plastic\\
eggs. Such ragged rabbits shed fake fur everywhere.\\
Paws jab the clouded sky with printed signs,\\
crooked words of spring sales,\\
a clearance of everything.

Just down the street, an old man holds\\
his Mickey Mouse head against his waist,\\
out of character for minimum wage.\\
No one minds, really, that his head wobbles\\
on top of Mickey Mouse's body as long as\\
he wears the costume, hands out red flyers\\
saying buy one get one free of almost anything.

Outside Liberty Tax Service, the statue paces\\
without her torch and wears a green, ragged crown\\
with no windows anyone can look out and spikes\\
drooping. No Independence Day tablet, no\\
tarnished gown. Just a sandwich board saying:\\
\emph{Early refund.} But she knows she has\\
nothing coming just like most everyone else.


\clearpage
\section{Still Life}

Today I captured stillness\\
in the palm of my hand. I held\\
the hush, listened to silence\\
brush my eardrum\\
soft as an egret's feather.\\
The sound of absence,\\
of white, of waiting\\
for you to surface\\
and walk to where I was

last year when we fished\\
together, when you taught me\\
to bait a hook, dangle a lline\\
from this dock I can still see\\
from the hill where I sit now\\
alone with all this quiet.

I came here today with my paints,\\
my camera as witness. I cannot\\
mix blue to match the placid lake\\
or the green of wilderness. Only\\
this snapshot emerged from my darkroom\\
where I stood still as the snowy egret.\\
Only my eyes scanned shadows, someplace\\
where you might appear in black and white.

But egrets do not fly until they feed\\
and I weep buckets of minnows for you.


\clearpage
\section{A Still Shot}

When theater lights dim\\
during Sunday matinees,\\
widows pull sweater shrouds\\
around their drooping shoulders\\
against the air-conditioned night,\\
muffling the popcorn shouts\\
and coke kisses of the young\\
whose budded hands\\
flower against pink angora knits\\
in the reflected light\\
of someone else's fantasy.


\clearpage
\section{Stone Chronicles}

In the hunt for history,\\
consider an ocean stone,\\
a slender, black, volcanic disk\\
worn thin from waves,\\
still full of fire, the story\\
waiting for release.

That grey stone holds\\
the moist breath of a dinosaur\\
who wandered here before us,\\
chomped on horsetail rushes\\
and left its bones compressed\\
in stones, its legacy of wings.

One gull feather tipped in black\\
spirals from the navy sky,\\
drifts onto one white glacial stone\\
chilled from the ocean floor, broken\\
from bedrock and carrying tales of what\\
happens down deep where few go.


\clearpage
\section{Strawberry Girl with Heart}

Three strawberry hearts ripen\\
on the bush, dangle their perfection,\\
their promise of plump flesh and red juice.

The girl sits by the berry bush, scribbles in her diary.\\
Seven years ago, she buried in this very spot\\
her last baby tooth tipped with scarlet at its root.

She tops her letter ``i" with a heart,\\
a flash of her red-glitter self. On the page, she curves\\
a berry, adds one black blotch, the mark of rot.

Changes turn on a breath of wind or the wiggle of a worm.\\
Nothing stays still. Nothing waits.\\
Her own eye blinks in her heart-shaped face.

A quick twist and two berries nestle in her palm,\\
sweeten her tongue with the taste of loss. A cardinal\\
snatches the blemished berry. The heart is a beating wing.


\clearpage
\section{Students}

I look into the semi-circle and see\\
sketched on foreheads, chins and cheeks\\
plump golden babies like cherry blossoms\\
overflowing mothers' arms,\\
green-robed fathers picking yellow mums\\
and geisha girls with downcast eyes.\\
So many Japanese puzzle boxes\\
painted and lacquered\\
daring me to\\
find the crack,\\
open the box\\
without a key\\
slip out the drawer\\
hidden underneath.

I probe, poke and pry until\\
a drawer springs open to my left\\
a lid flips up in the middle\\
and I stuff in verbs of being and doing.\\
Hemingway, Faulkner, and Oates leap\\
into velvet-lined boxes;\\
neatly stacked ideas\\
slide through cracks\\
expanding the dimensions,\\
exploding the sides,\\
and class begins.


\clearpage
\section{Tattooed Saint}

You're the real thing.\\
None of this cyber-saint stuff for you.\\
No sainthood.com with flashing green lights,\\
a good deed and a plastic master card\\
over the limit. But \$24.95 plus more\\
will buy you a shrine and a halo\\
you already have. Your miracles,\\
your heart is right where it ought to be

upfront with two chambers, both full\\
with spots of white space. Room\\
for a poetic squiggle, even a blue note.\\
You, patron of arts, are known everywhere.\\
Your pale face topped with blue, black,\\
red tam draping your ear. The tassel swings\\
both ways: the tragedy and comedy\\
of saints who grin at pain. The release.

We, the abandoned, yearn\\
for sustenance, something crunchy\\
enough to wake our souls. Cornflakes,\\
a burst of tattoos, every hue, all lines\\
and geometric shapes. The stuff of life.\\
You know the way your body shivers\\
as needles ink your deeds over ribcage,\\
collarbone and move to face.

On your Saint's Feast Day, we live\\
your pain, your sacrifice permanent,\\
marked on every bit of flesh, on us.\\
We pray you will guide our hands.\\
Our hearts clutch, paralyzed\\
in white space begging for our mark.

Oh, Tattooed Saint, we bear witness.\\
Summon the miracle of art.


\clearpage
\section{A Teacher Answers}

Questions curve your cheeks\\
cling to bone\\
rise like meat hooks\\
to peel cataracts.\\
You see\\
taste of blood\\
of whys\\
and grow dumb

like me\\
who scrapes chalk against skin\\
knowing all the while\\
answers are gauze\\
wrapping the wounds\\
and scars ache on\\
winter days.


\clearpage
\section{Teachers}

like plumbers aren't blinded\\
by porcelain and veneer.\\
They get to the guts\\
wrench in hand\\
twist copper screws\\
and rubber washers\\
pluck out\\
crimped hair and bits of fluff;\\
even the metal matchbox car\\
is snaked through curves\\
as cesspools drain\\
and clean voices\\
gurgle upward.


\clearpage
\section{Tea Party}

Silence brews between them,\\
a pot of tea.\\
The bags, heavy with water, hang\\
suspended by cellophane.\\
Leaves leak from the paper sack.\\
Specks float, settle,\\
foretell the future if she\\
or he can see the bottom\\
or even drink from the cup.

They put their finest china on the tray,\\
serve themselves the senna tea.\\
No cakes with pink frosting or wine,\\
just porcelain between their lips.\\
No Alice comes to take a place.

``Senna purges dreams,'' the woman says,\\
``too dark for gypsy eyes.''

He shuts his lids, asleep it seems,\\
until he starts to speak,\\
``No riddles today to twist the brain\\
or make things seem like what they aren't.''

She hasn't heard. Her clock has stopped\\
and no one comes to clean.

He takes a sip of tea and grows.\\
His voice is louder now.\\
``March hares arrive in April.\\
My hatters are all sane.''

She balances her saucer on her knee.\\
``If I were in the well, I'd hear\\
water in my ears.''

She sinks,\\
a teabag cut loose,\\
sprawled at the bottom,\\
ready for a reading.


\clearpage
\section{Tear off a Hunk of Bread and Chew}

The little girl who ate words\\
chewed them, savored them\\
on her pink tongue as she curled\\
against her father's side, her head\\
balanced on his arm, her eyes on whatever\\
book he held palms up, tilted toward her.

Each letter on each page\\
jumped into her mouth, tasted\\
new, sharp as cheddar cheese.\\
The whole alphabet jumbled,\\
melted together in a casserole\\
of spaghetti noodles and red,\\
round tomatoes with an acid\\
sting, a sit-up-and-take-notice\\
zing when Heidi, an orphan,\\
met Peter, the goat herder, and shook\\
hands with the hermit in the hut\\
who was her grandfather,\\
his voice cold as the snow-\\
crusted Alps Heidi had climbed.

The little girl snuggled closer\\
as she swallowed Heidi's story,\\
nourished by each chapter, warm\\
as dough baked into loaves,\\
the crusty bread\\
dipped into goat milk,\\
her daily sustenance.


\clearpage
\section{The Test}

The boy in the back is\\
a verb like walk, talk,\\
think of\\
tonight's dance:\\
black discs turn\\
other voices into\\
his calloused ear\\
out his mouth\\
as he taps\\
disco repeats in five\\
paragraph essays\\
with no conclusion.

The girl in front is\\
the noun-\/-\/-the person,\\
place,\\
thing that anchors\\
its weight in chairs\\
and sits\\
and rocks\\
solidly\\
a wadded dream:\\
flutes drown in\\
drum sounds\\
as she picks up\\
sticks\\
and beats out\\
paperball beginnings\\
pile up\\
on her\\
on her cluttered desk

But they pass the test\\
eventually\\
when grades\\
are plotted\\
on the curve.


\clearpage
\section{Their Marriage Album}

They speak in snapshots.\\
His voice, yellow smoke,\\
puffs of black and whites\\
held tight with corners.

``I said,'' he says,\\
``you used to listen.''\\
Close-ups blurred\\
he turns blank pages.

``What's that?'' she asks,\\
dusting in another room.

She turns the towel.\\
More strings unravel\\
like spider legs dangling,\\
skittering across wax,\\
wet as newly spun webs\\
above the cornice. Dust,\\
in the air, on her tongue,\\
settles softly,\\
dray as days past.

``Come look,'' he says.

He recalls retirement. His\\
day stuck in the ground\\
like a stake marks tomato seeds\\
covered lightly with soil. He\\
yearns for red, a sign\\
of ripeness, color of his blood\\
in daylight. Even pink\\
streaks the evening sky, delicate\\
as blown glass,\\
fragile as an insect wing.

``Yes, look,'' she laughs,\\
``A missing head,\\
you shot too low.\\
Our eyes were closed.\\
And there's a photo missing.''

In photos, they look\\
away. From each other,\\
from the camera, each\\
hides as if the shutter\\
may open too wide,\\
an empty space full\\
of light to tantalize.\\
Then, shut fast,\\
sharp as a guillotine.


\clearpage
\section{Therapist Treats Agoraphobic at Her Home}

Elderberry bushes rooted,\\
flourished in my fertile brain.\\
White flowers, every perfect petal\\
shriveled, replaced\\
by black berries\\
I turned into wine.\\
Sip a bit with me.

You, my therapist,\\
tracked me for years.\\
Your footprints\\
meandered along shores of my thoughts,\\
paused where you plucked cattails,\\
fuzzy in autumn wind, fluffy\\
and white as my fear.

I saw you toss a stone\\
in my river, feared\\
for my life\\
as concentric circles rippled\\
outward from center\\
as that rough-edged river\\
stone sunk in silt.

Such a small disturbance\\
could tug me under water,\\
could hold me there\\
until I grew gills\\
and minnow fins. My mind\\
huddled behind skull bone,\\
its home, right\\
where it belonged. Only I\\
should walk there.

Here, let me fill your glass.\\
As you drink\\
elderberry spirits,\\
you will understand the way\\
my neurons flickered,\\
lighted gray ghosts\\
outside my window. Yesterday,

I swam on dry land. Moist\\
with sweat, rivulets of me\\
dripped onto roads\\
red with Georgia clay,\\
with shimmering heat\\
mirages where I rested\\
on sand under weeping willows.\\
I wiggled my toes in the blue pond,\\
curled my skin around wetness.

Don't interrupt the telling\\
for I know where I go. You can\\
traipse after me, mark my route.\\
If you stare through my window,\\
you may glimpse one leaf\\
on that red maple, unfurling late\\
this spring as if it weren't sure\\
it should show itself at all.\\
I watched every day\\
waiting for that leaf to open.

Down the street\\
by the stoplight,\\
one diseased ash tree\\
draped its limb\\
against the green light\\
before the trimming.\\
I listened to\\
that awful whizzing saw\\
you hear\\
before something tumbles\\
to ground. Yes, I heard\\
the fall. I certainly did.

I could show you\\
ways to leave this state\\
I'm in. But then\\
you have seen\\
all seasons, anticipated\\
changes when snow melted\\
and wild daffodils dotted\\
grassy banks on the freeway.\\
I rode the backs of pigeons,\\
saw orange barrels closing lanes\\
under construction\\
for a better tomorrow I thought\\
the signs said. Hard to see\\
clearly when a pigeon feather

twirled toward earth, toward gray\\
concrete that would not melt\\
ever. Like my mind, the lines\\
led to other states, even\\
the Arctic Circle, a frigid zone

where all was frozen,\\
where I straddled an ice floe,\\
drank elderberry wine,\\
and toasted you who lived\\
my life outside this house.


\clearpage
\section{The First Day of April}

The day\\
all the daffodils tumbled\\
from their green beds\\
at the same time,\\
I knew\\
winter had given up its darkness.

On this first day of April:\\
the town jokers turned\\
street signs topsy-turvy.

Outside,\\
I swam the butterfly\\
stroke in the aqua sky\\
where the dampness spoke of spring\\
rain and ferns that uncurled their tongues.

I heard the trumpet of a daffodil,\\
a blast of air so sweet\\
the note centered in my throat\\
as if I could release it to the wind\\
and like a yellow warbler\\
it would take flight.

Red balloons floated the message:\\
the carnival was on its way.

Pranksters rode the Ferris wheel\\
as high as the sun,\\
rocked in its yellow glow\\
knowing their wax wings would not melt

until their feet touched solid ground.\\
Week after week the carnival stayed:\\
Icicles dangled\\
from the Ferris wheel\\
and snow nestled in the seats.\\
The townsfolk waited.

They watched for:\\
the sun that would coax life from the underground,\\
daffodils that would fall from the cloudless sky,\\
the turning of the wheel.


\clearpage
\section{A Timed Exercise}

If I had two months of life left\\
would I sprout and send down roots\\
deeper and deeper searching,\\
divining the blue\\
cool voice of wetness

or would I just float paper boats\\
in muddy sidewalk puddles\\
where sparrows dip their beaks?

I could assign tasks\\
of this and that\\
so friends\\
clip cotton strings\\
from this rag of life\\
wrung dry of tears.

Could I find something better\\
than driving the freeway\\
the same route every day\\
which ends when I slam the car door\\
and elevate to the 15\textsuperscript{th} floor\\
where my fingers dawdle over keys\\
reluctant before turning locks\\
needing oil. Still\\
jagged metal fits jagged metal.

I shall choose. I shall\\
scratch in cracked, brown clay.\\
With dust up under scarlet nails,\\
I shall make my mark\\
and brace for whirlwinds

in my garden\\
until my soul balloons\\
and blooms a meadow of blue chickory\\
and Queen Anne's Lace\\
before scattering its seeds.


\clearpage
\section{The Time Will Be}

She---\\
walking across waxed linoleum floors.\\
Reaching for the wall phone.\\
Talking long distance through wires.\\
Yelling at the boy tracking in mud.\\
Hanging up.\\
Wondering what to fix for dinner.\\
Pushing the phone buttons again.\\
Juggling the skillet in her left hand.\\
Listening to the busy sound.\\
Breaking the connection.\\
Punching now another number.\\
Hearing another woman starting dinner.\\
Saying nothing.

She---\\
Waiting for the sound of his car.\\
Waiting for the phone to ring.\\
Waiting for someone to contact her\\
About something she should have done\\
Yesterday, or can do tomorrow or next week.

She---\\
Knows she must stop waiting.\\
Knows she is forty now.\\
Knows the mortgage is paid.\\
She sits still\\
Waiting for\\
Something to happen.


\clearpage
\section{Tinkertoys}

When my grandson tilts the box,\\
105 Tinkertoys roll out. Orange connectors,\\
plastic pieces, small, shiny, slick,\\
notched at the center. Perfect\\
for holding things together\\
when they're about to fall apart\\
at the edges. Round wooden wheels

unfinished, until my grandson\\
crams one blue rod into the center hole\\
in one wheel, then joins the other end\\
to a different wheel. One flick of his\\
three-year-old finger launches this two-\\
wheeled car across the tiled floor\\
where it hits the wall, zooms back\\
once more into his agile fingers

content with his car, fresh from the assembly\\
line of his imagination. That blue car\\
without hubcaps, seats, engine, even headlights\\
still does what he wants, goes where\\
he aims it, covers the ground\\
between his warm fingers and the cold

basement wall and bounces back\\
in one piece. Safe as my grandson\\
who spends the whole morning building\\
towers where he can live, buses, trains,\\
airplanes, birds, the sun. A dictionary\\
of connections crackle along the synapses\\
in his mind, a place where he pokes around\\
and carries back adventures he constructs\\
with bits of wooden shapes, all joined

together. He cradles his world in his own quick hands\\
afraid a piece might break off,\\
fall to the floor and roll under the couch\\
where he can't stretch far enough to reach it.\\
He has placed every tipsy piece\\
just so. A pirate ship with ladders,\\
and a plank to walk over a sea\\
swarming with sharks and whales that\\
can eat him, he says, if he falls in

when I turn my head.\\
Not even for lunch\\
will my grandson dismantle\\
the pirate ship, put it back\\
in the box where it belongs.\\
He wants to tell his scary pirate stories\\
again and again.

Only I know\\
nothing\\
goes back the same way\\
once you've taken it apart.


\clearpage
\section{To Dylan at the Beginning}

May you discover\\
the whole world\\
a piece at a time.

Scoop a handful\\
of earth damp with promise; inhale\\
the vision of one brown acorn\\
alone sprouting into an oak\\
tree strong enough to hold you\\
embraced by the blue, blue sky\\
up high where your dreams perch\\
on the edge of the sun.

Catch one raindrop\\
on your fingertip; count\\
each speck of life, each\\
microscopic molecule floating\\
free as you swimming under\\
water, choosing to see\\
what is and\\
what can be.

Bite into one\\
plump, red apple; savor\\
the sweet juice, sun-warmed,\\
dribbling down your throat,\\
and in that moment, recall\\
that first nibble,\\
that crunchy taste\\
of truth.

Blow on one blade\\
of green grass; whistle\\
long and loud your own tune\\
\clearpage
of happiness, of friends forever\\
true, of books you read,\\
of what you do along the way\\
to make this world\\
a good and better place.

\begin{quote}
\emph{With Much Love from Grandma Carol\\
October 2, 2006}
\end{quote}


\clearpage
\section{To Each}

Sullen talks.\\
The couch and car,\\
divided dividends,\\
no animosity,\\
or so they said,\\
to each other.

The children, too, split\\
between Saturdays and Sundays\\
springs and summers;\\
even the cold death\\
months considered\\
and ignored, until later\\
when they discussed\\
other matters that required\\
more words,\\
their complete attention\\
to all details.

She and he got\\
through it all,\\
and were glad,\\
or so they said,\\
to each other.


\clearpage
\section{Tomatoes}

The poems I do not write\\
are crammed in my head\\
like so many overripe tomatoes\\
bleeding into each other,\\
dripping, unsynchronized,\\
out the bottom of the wooden crate\\
that warps and swells and rots,\\
giving birth to white worms\\
who lap up the red juices\\
of the words I cannot write.


\clearpage
\section{To Men Who Want Mothers}

She stands by the bed and captures\\
nightmares in the folds of her gown.\\
Fingers perch on her palms,\\
ready to swoop down,\\
pick off his moans in mid-air\\
and swallow them whole.\\
Her arms are red cedars;\\
incense oozes from her pores,\\
perfumes his phantoms\\
caged in a flannel fog.\\
Like a mountain\\
she holds her ground and grows.\\
Her head pierces clouds.

From the peak,\\
she rains April lullabies\\
against his closed eyelids.\\
She knows:\\
he must not look at her mouth\\
open like the tulip.\\
She pulls him to her heart,\\
feeds his dreams.


\clearpage
\section{To My Daughter on a Fine Fall Day}

I stand behind you and push,\\
harder and harder.\\
You pump higher.\\
Feet kick the air.\\
You swing back,\\
just miss my arms\\
waiting to receive you.\\
Hands fall to my sides,\\
hide themselves in coat pockets.

You're over the sandbox\\
filled now with maple leaves.\\
Shoes skim the hedge,\\
reach past this boundary line.\\
You lie in your bed of air,\\
no pillows for your head.

I watch you wake,\\
hanging there above me,\\
a sparrow in its nest.\\
Your fingers clutch the chains,\\
probe the links,\\
let go the hold.\\
You jump, splitting the noonday sun.\\
I cannot stop your feet from touching ground.


\clearpage
\section{To the Boy Who Rode a Skateboard}

You zipped from behind that bush,\\
right in front of me and I nearly had\\
a heart attack the way you didn't stop,\\
just flipped that skateboard up\\
and twirled it twice as if you were in a circus\\
and crowds paid to see you do what you did.

Let me tell you, you're lucky I stopped\\
in time. You could have been squished\\
under my wheels like a squirrel,\\
left with two broken legs and blood in your mouth,\\
your right arm trapped under my tire. I shudder\\
to think what might have been.\\
I could have had your death\\
on my conscience for the rest\\
of my life and I wouldn't dare drive again.\\
That would have killed me.\\
It would have been all your fault.

I wish\\
I had said all that\\
to that boy\\
who didn't even see me.


\clearpage
\section{Tonight the Round Woman Moon\\
  Swallows the Dark}

The moon glides across black silk,\\
bright, plump with promise. A silver puff\\
polishes the spoon's curve\\
slick as a woman's lips. Shiny\\
quicksilver strokes her throat.\\
A disc of glitter, a pendant of moon\\
rocks slides between her breasts, glistens.

A girl of fourteen sits on a grey boulder, her\\
long dark hair still, straight, almost touches\\
the cold rock as she leans back on soft hands,\\
considers the moon a woman, strong,\\
steady enough to pull the tides as she sweeps\\
the world with her white-gold gown. She illuminates\\
the girl, then shadows her in a grey cashmere shawl.

The girl can almost hear moon words,\\
whistled whispers in her ear before the silver circle\\
plops into the pond right in front of the girl\\
who opens her hands to catch slippery minnows.\\
Their scales flash white translucence as they jump,\\
leave only air and concentric circles on the mirror.\\
One moon floats on water. One moon moves across the sky.

The young girl wades into the pond as the water moon\\
gathers its shimmer and rises. The girl stands submerged,\\
moondust and minnows curl her hair as her arms reach down\\
for a fractured reflection and up to the round woman moon\\
who lingers atop a blue spruce over her right shoulder.\\
The girl bursts through the wet blue barrier, head above water.\\
The moon swallows the darkness, dips from sight.


\clearpage
\section{Too Dark}

for opaque mirrors,\\
your dread. Look into silver\\
mirrors instead. Or

ride the river rapids. Flash\\
emerald against blue. Rise\\
like a leaf in a whirlpool.

Your face can break\\
the surface. Wet,\\
wreathed with foam,\\
you drink the cream.\\
Breathe out\\
again.


\clearpage
\section{Transgression at Hospice}

\emph{For the woman with Alzheimer's}

Her brain cells pop like plastic\\
air bubbles shippers use\\
for packing, for sending on\\
breakables to other states.

Reminds her of a jack-in-the-box.\\
She had one, once; she hears\\
``Pop Goes the Weasel'' playing\\
now, waits for the lid to snap up\\
as if clowns pounce\\
from cramped places\\
and clean, unlighted spaces, dark

pockets like pills. Round and hard\\
they stick in her throat, dissolve\\
eventually. She untangles\\
her shoelaces, strips off support hose.\\
Varicose veins bulge against skin,\\
blue ink, the indelible kind.

She shrinks daily, a sliver\\
of herself peeled to the core\\
of apple seeds, moist\\
with juices. Tipsy and cider-drunk,\\
she steals out windows, out doors, out\\
any open crack, a contortionist\\
bending low under the limbo stick.

Her body a drum, skin stretched tight,\\
reverberates with memories, too loud.\\
Hammers pound, play her bones\\
like a xylophone, up and down the scale:\\
the mating dance. Back and forth,\\
her legs straddle the stick, this side

then that. Silence, sudden as white\\
elephants. Smooth as ivory tusks.

She scales mountains,\\
a solitary climber, lost\\
in the blizzard, pale as a papercut\\
snowflake unfolded, pasted\\
on a schoolroom window somewhere\\
in the north. She searches toeholds\\
in granite chinks. Igneous rocks\\
forged by fire, cold now, coated\\
in ice, hold her weight.

She pounds spikes into crevices\\
above her, as far as she can reach.\\
She uncoils cord from her waist,\\
flings the lasso, pulls the slipknot\\
tight. Up, fighting gravity, she\\
ropedances. Eyes water in the wind.

Ends fraying, genetic ropes\\
dangle down, twist through centuries.

When will he come? she begs nurses\\
repeatedly, the suitor with hair slicked\\
back, still damp. A dandy, her father\\
warned. She parts curtains,\\
watches for him. Today is

Tuesday, the nurses say again.\\
At dinner time, your husband comes,\\
feeds you, clips your nails. Must\\
remember that. They yank her back.


\clearpage
\section{Travelogue: Dad's Garage}

Look in the southwest corner, Dad's own\\
painted desert. Can upon can\\
dripped gobs of color.

See the specks:\\
our marigold kitchen wall, circa 1972.\\
Avocado green for the living\\
room. Sunset purple, royal blue,\\
lavender, the aster pink all\\
mixed by Harry's Hardware back then\\
when Dad skipped up Grandville Street\\
with us, warning: Jump the cracks. Hold\\
upright the aster, forget-me-not, violet,\\
even the celery stalk for Harry.\\
He matched all colors, stirred\\
again and again. Our experiments\\
exact: recreating life.\\
A swirl of green added for father,\\
who painted our world on April days\\
each year, windows wide open\\
was we held his ladder.

Check out the southeast corner. Dad's own\\
everglades. Hoses coiled into hummocks\\
drizzled his days, his 80 Midwest years.

Count the summer sections:\\
our sea wings, those brown\\
inner tubes there at the bottom. Ragged\\
as our toddler steps, lurching,\\
twirling under Dad's hand-held waterfall.\\
That green tube with brass nozzle\\
turned full blast during our teens.\\
Dad misted seedlings, perennials,\\
yews until delphiniums spiked blue,\\
sunflowers swiveled toward light,\\
furry faces dark\\
as our own bodies,\\
slick with baby oil and iodine.\\
My sister and I\\
grew into ourselves,\\
our bellies against bluegrass\\
knowing all the while our father\\
would warn against burning, turn on\\
the sprinkler. He, the water magician\\
coaxed life from dry seeds.

Glance into the northeast\\
and northwest corners, Dad's own\\
wilderness, full of wood and work.

That vise gripped broken chair legs\\
until glue held the splinters\\
together, healed\\
his daughter's hearts.\\
This was his work.\\
His shovel tilted there\\
against the rake.\\
He waited for Detroit snow,\\
predicted by a lone forecaster,\\
a record-breaking blizzard,\\
below zero. When Dad wasn't ready\\
for cold or death, it came.

We two daughters zip our jackets\\
against winter as we linger\\
there in that unheated garage.\\
We have come to clean,\\
bundle up our youth in Hefty bags.\\
We climb onto two blue bikes,\\
wait for his hand\\
snug against our backs.

Our feet pedal,\\
round and round once more.\\
We wobble forward on used Schwinns.\\
Ahead and on our won,\\
we pump the north wind home.


\clearpage
\section{Two Pine Cones}

I listen to crunch.\\
My boots squash paper\\
cups, one tin can,\\
three bottle caps\\
and oak leaves in the fall.

I hear it all\\
as I walk. Murmur of voices\\
screened by mesh, a shout\\
metallic with gunshot fear. A dog\\
yaps, wind shakes maple\\
leaves that cling to wood. Every\\
star clusters above me. Around\\
the moon and over roofs, stars shoot,\\
burn hot, drop at my feet. One\\
wish I'm granted in all this noise.

Under a streetlight, scotch pines\\
needle me, rub branch against branch. I am\\
alone on suburban streets. My subdivision\\
waits a mile up. I know its name,\\
the house with my number, the place\\
I bring back what I have found.

I set together\\
two pinecones waxed with moonlight,\\
stuffed with seeds.\\
One balances upright against darkness.\\
One tilts on its side, damaged.\\
One slender twig cradles them.

Two broken scales. Love spirals\\
out of control. The red squirrel feeds.


\clearpage
\section{Two-Way Mirror}

Then:\\
You called me Calliope\\
and heard with your lips.\\
Your hands\\
touched circus epics\\
until clowns tap danced\\
across the bridge\\
of your nose\\
in five seconds flat\\
and your quiet laugh\\
shattered crystal lies.

Now:\\
You smile through windows\\
as your concave eyes film\\
moon craters\\
of your soul's surface,\\
wait for volcanic lava\\
to harden\\
to petrify\\
before the steamy words\\
from droplets\\
on the glass\\
I cannot break.


\clearpage
\section{Uncle Amos}

Solid as a brick\\
the rich folks on the ridge said\\
whenever anyone asked. Someone\\
they could count on to make things right\\
when they were wrong

like when the carburetor plugged up\\
on their new Oldsmobile 88s\\
from the dust on the old county highway\\
and no air could get through\\
or when their silos rotted\\
and crumbled plank by plank,\\
full of carpenter ants still chewing\\
and spitting sawdust. Wasn't anything\\
Uncle Amos couldn't fix or rebuild\\
better than new, most folks admitted.

But it was tougher jobs\\
he hung his reputation on.\\
Cases like that No Good Johnny Ray\\
who could pluck a grandmother's garnet ring\\
from her finger without her even taking notice\\
or eat a blueberry pie still in the oven\\
without a trace of crust left in the tin\\
or stain on his lips. Uncle Amos believed\\
anyone, even Johnny Ray, had redeeming qualities\\
even if they didn't surface much. Before anyone knew\\
what was missing, it was back in place.

Uncle Amos never let on how he did it,\\
just hunched his shoulders,\\
rubbed the back of his neck\\
like he was working out some stiffness\\
settling in his joints.

Even today, when ridge folks talk\\
of Uncle Amos, they wish he was still around,\\
bless his soul, they say, a solid person.\\
As if he was\\
nothing more than a red brick\\
from the old Methodist church,\\
part of the design built into their back patios\\
beside the swimming pools, all Olympic size.

Wonder if they ever noticed\\
how Uncle Amos' eyes\\
were the grey-brown color\\
of a sparrow's feather\\
and his eyelids fluttered\\
like wings in open space.


\clearpage
\section{Under the Microscope}

I was a child again\\
last night scouting\\
the perfect place where\\
I knelt, scooped pondwater\\
until my empty jelly jar was\\
full of wetness,\\
my own piece of pond I carried\\
home with the tadpole swimming\\
side to side. Overnight, sediment\\
dropped to the bottom and I saw\\
more clearly all my barefoot days.

Squeeze just one transparent drop\\
from my pond jar I urged students,\\
onto your sterile glass slide.\\
Now gaze at life\\
magnified. Stare long enough\\
for green dots, strands, squiggles\\
of algae. Watch them multiply\\
before your eyes. Ah, you see the yellow hairs\\
sway like your own as you toss\\
your head, lift your eyes confused.


\clearpage
\section{Untitled}

I have never seen\\
the big brass, the faceless one,\\
all angular arms. His pinchers\\
grip our round: a few countries\\
or maybe even blue oceans\\
wet enough to drop. No safety net.

Centuries of wars, famine, plagues,\\
tarnish our ball, pile weight on one side,\\
then the other. We all scurry,\\
here to there. We slide past\\
his brass forearm, past his wrist bone,\\
right into his palm. We dangle there.

The big brass can not\\
be bribed with gold. He will protest\\
declarations, explanations, proclamations,\\
tallies of wrong that may be right. NO,\\
he commands balance up, down, across\\
and around. Resist and go to jail.

The strategy of brass is white button.\\
A fingertip temptation.


\clearpage
\section{Upon Being Asked for Advice}

You mapped your face in mirrors\\
for fifty years. Traced\\
every open pore, closed\\
mouth bracketed by worry\\
lines. Two furrows

cut deep between your brows\\
cradled your babies, held them\\
close. Safe behind your eyes. You\\
swaddled them in your own\\
imperfections, reversed reflections.\\
Skin to skin, they knew

you inside out and\\
outside in

as I did, my friend. Do you\\
want to chart, again,\\
your whole geography? Change\\
our ways of seeing: who you are\\
and where you've been? All

this medical sorcery\\
the price for skin pulled tight\\
as a girdle, holding in\\
what should be let out? Do you

choose a pattern\\
in your size? Does it fit\\
with seams, stitches, tucks\\
like a child's outgrown smock?\\
Will it

make your skin transparent,\\
luminescent, hold your death\\
day in abeyance: a distant,\\
fluorescent green glow?

Too dark for opaque mirrors,\\
your dread. Look into silver\\
mirrors, instead. Or

ride the river rapids. Flash\\
green against blue and rise\\
like a leaf in a whirlpool.

Your face can break the surface.\\
Wet and wreathed with foam,\\
you drink the cream\\
and breathe again.


\clearpage
\section{Uptown Jazz}

The jazzman's fingers skitter\\
over the naturals, jounce off\\
sharps and flats. He plays\\
in that place where no one\\
can live for long,\\
where the first star's bright\\
wish light sucks him up,\\
up where storms brew inside\\
white clouds, where lightning licks\\
the licorice sky before his rhythm\\
rattles hickory roots, grumbles,\\
tumbles into the heartwood,\\
slides under the ebony and ivory.

Musical notes rumble\\
in the jazzman's belly,\\
primitive sounds he can't\\
write down, won't remember\\
tomorrow. His throat grunts\\
with each chord, each key\\
his fingers stroke, his song\\
from the start. Only him\\
and wind and fire\\
and water and dirt.


\clearpage
\section{Vapor Rising}

November is the cold month\\
when black nights fall into black\\
Lake Superior which rocks its dead\\
whose souls rise to white,\\
curl into luminescent wisps of loss,\\
of memories strong enough to hold\\
the hull, pregnant with cargo

like me rounded and full of life,\\
my belly an aquarium\\
where one fish swims safe\\
in dark water without sight.

I stand watch, legs braced\\
against pain in this pilot house\\
as I navigate our passage\\
through each contraction,\\
each breath I take for you,\\
my child, who knows nothing\\
of life or death or reasons\\
why one freighter splits a path\\
in this opaque night you breach.

Water breaks warmer than air tonight.\\
Vapor rises white as cumulus clouds\\
that float this vessel, this child of mine.


\clearpage
\section{View from My Blue Bungalow}

The woman next door lost her job last year\\
and can't find hope anywhere,\\
not even on her couch\\
behind the yellow beaded pillow\\
from India where the elephant is frozen in time,\\
its trunk raised for good luck when in truth\\
no one has any luck at all these days.

What with the cost of coffee, my friend\\
down the street is tempted to give up\\
the drink altogether but his brain fogs\\
from lack of caffeine. He begs\\
the neighborhood watch group\\
for day-old coffee, even discarded grounds\\
good enough for one more perk, one more drip,\\
enough to sip while he searches\\
for some job a man might slip into\\
as if he had never slipped out.

Even the for-sale signs on my dusty street\\
have faded and tipped onto dry dirt\\
where nothing takes root.


\clearpage
\section{The Vigil of the Man with One Good Eye}

I knew the man, his one good eye\\
turned sideways toward the northeast,\\
full of storm clouds that blotted\\
the ink-stained sky. He watched

for you, prayed for your passage\\
while he called home the swallows\\
lost in air, buffeted by wind currents\\
like you. He charted your ragged course

toward shore, warned you with a single light\\
that you would crash on jagged rocks\\
submerged, lurking there to slice your hull\\
sink you in your ignorance of how to navigate

past the siren who sings her sweet song\\
of death, of promises that will not be\\
if you abandon course. The man with\\
one good eye sees past the danger of what is,\\
straight to your vision of what could be

if only you wait for clear weather. Listen!\\
Hear the metallic sound of the bell? Taste\\
that sharp bite of metal? Swallow dark air.\\
See the one-eyed man above the waves?

He rings you out to the center, the deepest place\\
where I once pumped his hand, wet with foam.


\clearpage
\section{Vineyards}

Hummingbird words buzz in my ear,\\
hover about my eyes,\\
lose me when I was found\\
unconscious on city streets,\\
awaken me in vineyards\\
with ruby-throated sounds,\\
the nectar sweet on my tongue.\\
Slender bills draw juices,\\
drip wine into my parted lips.


\clearpage
\section{Waiting for You at Morels Restaurant}

A pink-suited woman asked\\
about the empty seat\\
next to me. That oak chair\\
full of air, standing there\\
behind the place set for you

next to me. From the other side\\
she leaned over her gold-rimmed plate,\\
over chicken, two half breasts\\
browned and simmered in wine\\
sauce flecked with morels.

For flavoring, she said. The black\\
morels can be gathered. Look\\
north near wet elm leaves in shade.\\
Like men who stake an empty chair\\
you will find fungus growing there.


\clearpage
\section{The Waitress at Big Boy}

For two years, the same\\
plump waitress wobbles over\\
to my red vinyl booth.\\
Her eyes rest on the plate\\
glass window behind me.

Color her face red\\
like a check in Big Boy's pants,\\
like a limp tomato slice\\
tucked under the lettuce\\
in the green, green salad bar.

She chants her litany,\\
all the specials she can offer:\\
a Big Boy with its secret sauce,\\
the Slim Jim with French fries,\\
a fresh strawberry pie with whipped cream.

I do not know what to call this waitress\\
who wears no name tag,\\
who licks the tip of her pencil\\
ready to record whatever\\
I may order. She writes the exact price.

No-name turns away with my choices\\
clutched in her fist as if she may forget\\
everything if she doesn't hold on tight.\\
Her shoelaces drag on the tiled floor,\\
pulling crumbs of bread along behind.

Grey hair falls over her right eye.\\
Her feet slip-slide to my table.\\
She looks at no one, just sets\\
dishes in front of me. Her no-polish\\
no-shape, frayed nails betray her.
\clearpage
She walks many miles\\
every year and arrives\\
nowhere she ever wants to go,\\
and she has no cat at home,\\
not even a grey kitten.

I imagine her name is Ruth.


\clearpage
\section{The Way in Is the Way Out}

I lost me somewhere\\
among the tire tracks\\
on a rutted winter road.

Spring cycled me around.\\
I walked alone into the toothless\\
mouth of the sun's labyrinth.

Here, I gave myself up,\\
stumbled over hot stones,\\
resisted magnetic forces.

Green hedges grew toward me.\\
Such stiff, dense, unyielding\\
walls nudged me ever inward.

I spiraled toward the center,\\
toward that open space where I would\\
melt into a bubbling yellow puddle.

I feared evaporation,\\
the permanent loss of substance,\\
nothing left of me in this clearing.

Imagine how each yellow bubble rose\\
on air currents, broke into nothingness.\\
From such emptiness, I formed my meaning.

This is my story of transformation.\\
As solar flares blazed, I could see\\
I was where I was supposed to be.


\clearpage
\section{Weaving Her Tapestry}

People come at her as she lies\\
woven into her narrow bed\\
by spider-lines\\
drawn tight against life's loom.\\
Transparent fibers bind\\
her fresh-spun dreams,\\
bring faces back again\\
in the city's oiled night.

It's him, above her,\\
coming back in blackness,\\
his marionette arms\\
pointing behind him\\
at vacuous woman-faces\\
sewed into her brain\\
before he left her bed\\
and chose\\
grey-eyed girls with silver bodies\\
to walk his soul past midnight's ledge.\\
He talks to her of times they've known---\\
blackberry brandy years\\
when moon crickets\\
serenaded the sound\\
of bodies\\
coming together.

She cannot close eyelids\\
stitched open with spider-threads\\
or move her flattened fingers\\
from the tapestry\\
to touch the flesh\\
of those who greet her.

A childhood friend returns,\\
older now,\\
too old.\\
Her breasts droop;\\
the paper voice recalls---\\
worn Kentucky bluegrass\\
between their houses\\
where painted dolls lost smiles\\
in the rain\\
as two girls snuggled together\\
under crazy-patch quilts,\\
talked of boys\\
and flat chests growing\\
into cupped hands.

Now her mother,\\
dragging Father behind,\\
kneels over her,\\
murmurs magic chants.\\
Onion skins, mosquitos' wings,\\
Circe's voice, one black hair,\\
the midwife's cure.\\
Two voices remember---\\
cypress knees skinned\\
on gravel swamp paths,\\
summer's hair plaited\\
against autumn winds.\\
The child who was\\
before she came\\
is wedding white.

Morning light rewinds\\
tangled threads\\
and she rises\\
from her deathbed,\\
dressed in a sheet of white.


\clearpage
\section{The Weight of Water}

The flash and dazzle of this lake of giggling children disappear under
the wet, lost in blue-green shades of in-between. My son dangles upside
down from a fishing line sharp enough to half the sky if it were black,
a stain against the very air he breathes. No more than a glint of skin
covers his skeleton, my son's white bones concealed in secret.

Confusion reigns when my son bursts into flame. Beside him, a lake trout
leaps and dives as its flank and tail flip, flaunt red spots with blue
halos. Scales weigh the distance between water and air. Waves of light
wiggle through the sky. The sun explodes in an orange fireball crunched
between the scorched molars of noon. One drop of water on a lily pad
illuminates forest green under royal blue. Children frost the waves with
glitter.

Freeze the moment like ice before it melts.


\clearpage
\section{What Cold Cell is This?}

The police arrested Tillie Mae\\
when she roamed Oakwood Cemetery\\
after midnight. Even the Big Dipper hid\\
under the black blanket stretched\\
across the sky. No constellations

claimed their brilliance above Tillie Mae\\
whose radiance flared from a flash-\\
light wobble and old batteries.\\
With one stiff finger she traced the names\\
carved on her family's headstones.

Last in the line in this small town\\
on this November night, Tillie Mae\\
spread her best lace tablecloth, set the chicken\\
wings just right before she asked her mother,\\
``Who would care? Who would

summon the police\\
to this ghostly place\\
all because of a failing light\\
slicing the night?'' Tillie Mae's frail voice\\
cracked with such terrible loneliness.


\clearpage
\section{What Happened at Loon Lake}

No one swims today.\\
Cool fall breezes ripple my skin,\\
skitter across Loon Lake, hurry\\
water skippers over its surface,\\
rustle the heart-shaped lily pads\\
still green after 30 years as if\\
the whole scene graces a post card\\
with only white space on the back\\
empty of memories anyone\\
would dare write\\
about that day\\
I must remember.

I make myself\\
listen for the splash of wet\\
drops against my skin, the watery\\
memory of laughter, the gurgle\\
of adolescent girls, my four friends\\
as they dive under water\\
and split the surface back then\\
when their swim caps bloomed\\
fluorescent flowers all the rage.\\
Those four young heads crowned\\
with lime rubber petals, hot\\
pink, even orange, float\\
right there among white water\\
lilies. So much life full of fragrance.

I lie\\
sun-soaked on shore. My toes\\
curl over grains of sand, measure\\
seconds the sun remains\\
behind clouds, the very moment\\
I did not answer\\
my friends' calls to come on in\\
the water's fine I did not\\
want to wear my swim cap\\
without flowers of any kind\\
without the iridescence of even\\
one loon feather or a bluegill's fin.\\
I hunch today\\
on shore, unable to name\\
the exact moment one\\
friend slipped under the skin\\
of Loon Lake, her lime green\\
flower returned to seed.

I know now\\
death occurs between\\
the opening\\
and closing of one\\
white water lily\\
luminescent before night\\
drops over Loon Lake.


\clearpage
\section{What I Remembered}

I was only six,\\
old enough to never forget.

We were in a hurry that day,\\
late to my ballet lesson.

Mama was driving, talking out the side\\
of her mouth, \emph{You never, ever listen.}

And she smacked my bare leg. I think\\
I remember the sound and the red mark.

I knew she wished I would cry, hoped I would\\
say I was sorry as if that would make it all right.

She was looking at me when it happened.\\
I saw it coming. \emph{Go back, stop} I screamed.

I saw the blur of red fur, heard the horrible\\
thump against my door. Mama stopped the car.

She never came all the way around or kneeled\\
down or stroked the red fur wet with blood. I saw.

\emph{God damn dog,} she muttered. I opened my mouth.\\
\emph{Shut up. It's all your fault. And don't you ever forget.}


\clearpage
\section{What Might Have Been}

My ears have seen\\
the very moment\\
when dawn breaks\\
into a thousand shards.

My eyes have heard\\
the shatter, the scatter\\
the deafening scrunch\\
of rose-colored lens under foot.

After the petals of pink tulips\\
drift from their grass-green stems,\\
after the puzzle of pieces jumble,\\
no glue can hold.


\clearpage
\section{What Robins and Starlings Have to Say}

I just found a dead robin\\
on the basement floor of my new house.\\
No, I did not leave\\
front doors open. I slammed\\
every window, twisted every lock\\
against intruders. I swear I did. No,\\
I have never owned\\
a cat or dog. I would never accept\\
a pet whose teeth could crunch fragile bones,\\
could force me down on hands and knees\\
in grief. Do not accuse me\\
of forgetfulness or negligence.\\
Do not blame me\\
for this corpse. No.

How did this robin sneak\\
past me, sidle through\\
some invisible crack\\
in the foundation? How?

I only know\\
the dead robin on the basement floor\\
of my new house does not sing.\\
Such silence speaks of red feathers,\\
a throat swollen with unshed noise.\\
Do not warn me of death ahead.\\
Such tittle-tattle tangles me\\
in dreams where robins peck\\
me bald, weave my hair into nests\\
where baby robins rest and chirp\\
as if nothing can invade their space.

I will bury\\
my bundle of brown and red,\\
in a shoebox under the cherry tree.\\
Let this be the end. I refuse\\
to see this as an omen.\\
\clearpage
Death would only dare\\
present itself as a blue-black starling,\\
strut across frozen ground\\
straight to the front door\\
scream \emph{caw, caw, caw}\\
until it got in. Its three-pronged feet\\
would sink into my white carpet,\\
present another puzzle I can not solve.


\clearpage
\section{What the Old Woman Sees\\
  from Her Frosted Window}

The old woman wanders when dusk drops,\\
glints on the frosted window\\
cracked with age and a rock\\
thrown at her second floor flat\\
yesterday by some boy who stared at her\\
long and hard before\\
he tossed that rock, before\\
he split her window\\
in half. She still sees\\
the rock as it rises and falls,\\
the boy as he struts off,\\
his Detroit Tigers cap twisted backwards.

On this night of the first snow,\\
the old woman changes into her\\
white ruffled nightdress, her very best.\\
She rests her forehead\\
cold against the window\\
frosted with her breath.\\
She counts dormant trees\\
those big, dark trunks of sleep\\
across the street. Five maples\\
rooted there. Their branches hold,\\
then break in wind sway.\\
Twigs snap, twirl to white,\\
One trunk moves with two legs\\
across the crunch, leans\\
into the blizzard, raises two arms\\
as if any human can\\
hold back the sodden dark,\\
keep this night from falling.


\clearpage
\section{What the Photographer Found
on the Abandoned Street}

Brown pill bottles crammed with dirt,\\
clay, sand and concrete bits lined the window sills\\
of one tipsy house without a roof where air\\
and crows could perch. NO doors or window panes\\
slowed entry into vacant space unadvertised,\\
unwatched even when fire swallowed one whole wall\\
last week and spit black-red embers\\
onto parched ground, hard as hope.

The photographer focused on the torn,\\
white labels, the names of pills faded, the dosages\\
smeared, the patient's name illegible,\\
the pharmacist's address gone like whoever lived\\
in this house with brown pill bottles. The noon sun\\
burned into the bottles as he unscrewed\\
white, plastic caps, poured out bits of earth,\\
photographed the sediment of one man's life.


\clearpage
\section{What the Thief Took}

At eighty-eight, Sam did not know what he looked like,\\
the man who stole his wallet at the Dollar General store\\
and became him.\\
That no-good, money-grubbing guy\\
overdosed on The Suicide Machines'\\
\emph{Destruction by Definition} paid for\\
with Sam's credit card. The next bill listed\\
a fur-lined leather coat, a room at the casino\\
and other stuff Sam never bought.\\
He conjured up a young crook,\\
with black hair combed straight back\\
but always falling forward. Nothing stayed\\
where it was. Even Sam's wallet took a hike.

No matter what anyone said, Sam lost who he was\\
in the aisle of orange plastic Easter lilies at the Dollar\\
General store and he would never get himself back.\\
Even when he looked through the plate glass window,\\
no reflections took his shape. He lost his way home\\
more frequently as if he were wandering about\\
on the last half inch of his ruled life\\
searching for his name and finding only\\
streets he could not recognize and wearing a red-striped\\
wool hat he would never pull down over his ears\\
like some goofy kid with no name,\\
shuddering as he wobbled through December,\\
snow piling up at his feet.


\clearpage
\section{What We Did the Last Night of the Year}

Even my dog let loose howls\\
my neighbors pretended they didn't hear\\
tonight. Too many one-note party horns blasted the air.\\
Too many feet slinked up the stairs and around the corner\\
and tap-tapped into the bedroom on the second floor\\
where men snatched winter coats\\
worn by those women who slipped out the back door\\
and later snuck back to tip whiskey down their throats\\
thrilled at the risks they dared before the last\\
tick of midnight faced them forward.

Maybe twenty neighbors crammed together\\
under blue, crepe paper twists in the basement\\
where the whole neighborhood could fit\\
if they didn't settle for too long\\
under aluminum stars dangled from silver threads.\\
Arms reached out helter skelter, fingertips slid along\\
cheek bones, lips pressed lips and feet slip-slided\\
on the tile floor. Wiggly bodies dripped sweat.\\
people yelled for their spouses, aroused\\
when the end moved nearer and a last kiss\\
became the first, hovering\\
between lost and found.

One man scurried home\\
alone before the final, fatal tick.\\
The changing of the years,\\
the tick tock of clocks\\
echoed in his head.\\
He stroked his gun\\
three times, tilted the barrel toward the sky.\\
When he squeezed the trigger,\\
the bullet split the air above his head.\\
He did not see\\
where it lodged or if\\
blood flowed in the new year.


\clearpage
\section{What Women Know}

The fire moon ignites,\\
burns to red lace\\
passion in my scarlet room.

With my eyes shuttered,\\
my fingers stroll\\
across my flesh.

I read each bump, each\\
pink point of your message\\
coded in braille on my skin.

Listen for the shush,\\
the slide of skin against skin\\
as we speak in hot, red tongues.

Flames blind me and I cannot hear\\
the words you do not say aloud\\
over the hissing sound of heat.

Women know the lesson of red lace,\\
the way the embers cool to black.


\clearpage
\section{What You Throw Away}

I am the insomniac\\
dragging sleep behind me\\
like an empty red wagon\\
full of night air. My house

slippers slide over concrete walks\\
past your house of dreams\\
where you rest not knowing\\
how I steal what

you discard of yourself every week\\
on trash day. I mine bits of you:\\
letters from your lover, leftover\\
chicken bones, mostly wings and thighs

reminders of flesh against flesh, whiskey\\
warmth drained from the bottle, red roses\\
crumbled to dust, razor blades too dull\\
for blood or skin and I carry these trinkets

in my wagon for nights.\\
I dare not close my eyes.


\clearpage
\section{When Birds Fly South}

Even the birds have left me alone\\
with an empty bird house and frozen seeds\\
like my husband who drove away Saturday.


\clearpage
\section{When I Was Unlearned}

My fifth grade science teacher\\
stuffed students into formaldehyde jars,\\
studied the unborn embryos\\
curled in the dark liquid,\\
labelled on sterile white squares,\\
their various stages of unbirth.

In my dream-soaked bed,\\
his bald head glistened\\
like a fish's belly\\
on the dissecting table\\
as I raised my steel knife\\
and cleanly slit a line\\
to probe his printer's-drawer mind\\
with its letters, numbers, signs\\
neatly categorized, a definite\\
order. Letters without words.\\
With jagged stitches, I\\
sewed the gap\\
and closed my eyes.


\clearpage
\section{When I Speak Out}

I shall uncross my blue-veined legs,\\
unclench my hands\\
folded against my skirt of forget-me-nots.

Fingers lace like shoestrings;\\
I shall pick at my knots\\
until my fingernails fray\\
and white moons at the base rise up\\
toward the crest of a snow-capped mountain\\
where my toes tip my feet on end

as if I were the only Wall Street stockbroker,\\
the clear-sighted woman above the crowd,\\
my arms bare to my elbows, naked skin\\
exposed as I cradle international markets\\
with products and companies bought, sold\\
and rocked against my flesh.

As if these stocks, commodities and bonds\\
were more than mere paper pulp and ink,\\
more than glaciers\\
slipping into valleys, melting\\
into salty seas and carving out

the space\\
where I shall stand erect\\
and drink my toast\\
to red cedars and blue spruce\\
that split night skies between themselves\\
just as I, nourished on brine,\\
shall grow taller than these trees

until my head pokes through cumulous clouds\\
and seeds the darkness\\
with words of rain.


\clearpage
\section{When the Narrator Speaks}

If leaves\\
curl from aphids thick on blaze roses,

release five hundred ladybugs,\\
a plague of open mouths, a flurry of red-beetle wings.

When the white moon rises\\
before the pink sun recoils from the day,

turn toward the moon, a mere sliver tonight,\\
a curve of promise towing the silver sea.

Do not be angry at the red lips of the forest fire\\
as they gobble green, spit out ash.

Brush aside dust on long, wet days\\
until pine seedlings stitch scars and fireweed flowers.

And when the creek vanishes in dry air,\\
when fish flounder on sand and ghosts rustle the grass,

lie down in the bed where nothing can sleep.\\
Hear the story of stones, the hoarse tale of leaves.


\clearpage
\section{When the Petals Drop}

Four separate seedlings gloom their beginning---\\
one purple sweet alyssum close to the ground,\\
a squat marmalade marigold, a hot pink impatiens\\
with a white throat, and a viola dressed\\
in five loose petals above heart-shaped leaves.

Above these flowers my sweet-faced,\\
green-limbed grandson tilts his watering can,\\
wants more leaves, more flowers, more\\
life as water pours from the spout.

Five viola petals fall ever so slowly\\
onto the black earth before he bends\\
and picks up each yellow-orange petal\\
and puts it oh so carefully in his pocket.


\clearpage
\section{When You Hear the Story}

Think of Goldilocks who tested\\
her existence by the rule of threes\\
and settled for the middle,\\
for the just-right porridge. She\\
dropped into dreamland\\
only when she laid herself out\\
on the not-too-hard, not-too-soft bed,\\
exposed on top of the quilt, not under.

All three bears gaped and grumbled\\
when they discovered this uncovered\\
intruder who had gobbled\\
the mother's bowl of porridge\\
after only a nibble of the two extremes,\\
thank goodness. For all Goldilocks knew,\\
father bear might have clutched her tight,\\
squeezed her dead\\
if she had preferred hot or cold\\
instead of choosing just-right.

A more gifted storyteller would have\\
pawed the dirt, then stood up\\
on two hind legs, would have\\
gone for the drama of too hot\\
or too cold. Forget the facts.\\
Hear what you want.\\
The truth lingers in the telling.


\clearpage
\section{Where Our Enemy Lived}

My best friend and I wrote\\
directions to our enemy's house.

Turn left at the end of Oak onto Elm,\\
streets named before Dutch elm disease

devastated blocks of trees. DDT spray almost\\
killed us even though we hardly breathed

that Detroit summer in 1955 when we were\\
twelve and skinny as sticks. Our enemy's fat mouth

called us ``a carpenter's dream,'' as if\\
our chests would never swell.

Must have been a million times\\
we rang his bell and hid behind his dead

trees near the street. He could not see\\
us crouched there, hunched

in our own soft shells, satisfied and dangerous,\\
beautiful beneath our thin skin.


\clearpage
\section{Wicker Chair and Coreopsis}

My mother left this white wicker chair\\
to me who knew the intricacy of the weave,\\
the way each strand overlapped\\
and wrapped over and under itself,\\
creating the pattern of us.

There in her private place I listened\\
when she hummed like the wind rustling leaves\\
in the plum tree. She whistled sparrows to the ground:\\
t-weet, t-weet, t-weet. She could make me hear\\
the rise and fall of butterfly wings in air.

Mother kept her sewing basket full of thread\\
on the white wicker table. All colors\\
moved through the eye of her needle: the deep blue\\
of a jay's wing feather, the red of her blazing roses\\
climbing the fence, the white of clouds shifting shape.

From pieces of fabric, she designed my clothes,\\
knowing what would fit, what would bring out\\
my coloring, what would move with me like skin\\
when I was ten and growing faster than grass.\\
She sewed me into myself, leaving seams to let out.

I watched her fingers move, the needle flash in\\
and out of cloth as she turned flat flowered cotton\\
into a full skirt buttoned at the waist. She was\\
a magician who taught me how to sit quietly and wait\\
for coreopsis buds to open into gold suns at our feet.

Every year the coreopsis bloom on the same day\\
and I place the white wicker chair and table just right.\\
Light weaves my mother's shape in that quiet space\\
and I hum along with mother's songs, hear leaves fall\\
as mother performs her magic tricks and all is an illusion.


\clearpage
\section{Wild Field of Men}

I name the vegetation\\
at my house: crabgrass, bent grass,\\
ryegrass, Kentucky bluegrass and one weeping\\
willow down at the creek where I toss\\
memories of men. Like stones, they sink.\\
I'm reminded of Tim when I smooth\\
a blade of grass between my fingertips,\\
slit it with my tongue and whistle\\
a shrill blast in the August heat, a warning\\
call to those randy dandy slivers of men\\
who invaded my lawn on summer days.\\
I still recall their pale flesh\\
feel their hands like insects crawling up my leg.

I know the Latin names of flowers\\
as well as my neighbor's maiden name\\
embraced after her divorce. Since then,\\
she won't answer her door when salesmen\\
knock or petition bearers wave their clipboards.\\
She wants nothing and no cause is good enough.\\
Yesterday, I watched her cut a daylily\\
doomed to die that night.\\
She poked the fragile flower through\\
the top buttonhole of her black blouse.\\
Orange pollen bloomed on her right breast.

I plant sentries, yellow marigolds, brilliant\\
suns firing messages into the sky.\\
These guards are women who run\\
for life and tell our stories in dance and song.\\
Their names nourish me and my neighbor\\
as we toil in our gardens.


\clearpage
\section{Window of the Black Moon}

\emph{``The days just before and after February 26 when the moon is
  dark may offer the ideal window for US forces to launch a nighttime
  attack on Iraq.'' \emph{The Detroit News}, 02/12/1998.}

February is the fast month,\\
full of shackled hearts and lunacy,\\
that cold, cold month when cousin Mark\\
limped off the plane at Metro,\\
thrown out of Vietnam, no longer in uniform,\\
with silver crutches\\
and one pinned-up pant leg.

His mind a savaged moon\\
puckered with unmapped craters\\
where the air strikes whistled\\
a catchy refrain to an old, old tune:\\
Persian Gulf, Iraq, Kosovo.

He alone faced the TV news,\\
with red-streaked eyes,\\
with tributaries of red, red blood\\
flowing pure from the source,\\
his heart.

Even today,\\
he longs for war, a bloody battle\\
against the black, black moon\\
that bumps against his chest.


\clearpage
\section{The Woman's Field Guide to Survival}

I have seen the chrysalis\\
of the painted lady. Her\\
brown house is upside-down\\
before walls crack and fall away\\
from the green, globe thistle leaf,\\
from her butterfly body slick as mine.\\
Motionless\\
we hang onto the damp hours before dawn

and count the high cost\\
of a single wrong choice,\\
a split-second indiscretion

before our colors blossom: our camouflage complete.\\
Our wings tipped with eyespots expand, a burst\\
of pink and purple woven lace across our backs\\
strong as any grocery sack that holds our sun.

I have seen the painted lady\\
perch on goldenrod\\
and heard her wings rustle against my lips\\
as she wrangles the wind with me.

We hear the whisk and woosh\\
as the blue jay tilts his beak\\
and dives for us.\\
He may get one

eyespot but we still see\\
the everlastings, white puffs\\
of baby's breath and moon-red\\
crests of cockscomb. I have seen

the painted lady\\
lay her froth of eggs and leave me\\
alone in wetlands

\clearpage
where I stand on one leg\\
in the warm shallows\\
like a blue heron. I choose

fish words that streak\\
through waves and I wait\\
for one perfect trout.


\clearpage
\section{Women Past and Present}

See the darkness in my eyes\\
inherited from my grandmother\\
who came to this country\\
from another place\\
where women plucked tomatoes\\
from tangled vines, bit through\\
skin, chewed chunks of red,\\
let juices dry on their chin\\
in remembrance of green words\\
unsaid, growling in their throats.

Here she settled on farmland,\\
knowing soil could be tilled,\\
planted with wheat and corn\\
or lay fallow for a season.\\
She picked corn, shucked ear\\
after ear, tossed tasseled silk\\
in the compost heap where it decayed\\
with green husks and table scraps,\\
a fertilizer for next year's crop\\
which rooted in her eyes\\
and withered in the drought.

On this land, she waited for a sign,\\
gave birth to my mother\\
who gave life to me on city streets\\
one November day when snow\\
froze its crust of ice\\
over my mother's eyes shut in pain\\
until she cradled me and saw\\
my grandmother's eyes, her own eyes\\
staring back unblinking, black as screams\\
echoed by generations of women\\
who want their voices heard\\
through the ragged drapes of time.


\clearpage
\section{Women's Poems}

Women's poems are personal.\\
Politely phrased participles modify, clarify\\
identify sauerkraut boiled on back burners\\
and onion rings fried round their fingers.\\
Odors crest like waves,\\
crash against typewriter keys,\\
cramped fingers snap like lima beans.\\
Their words encased in skin,\\
stripped now to kernels.

Women serve stanzas on stoneware.\\
Iris lie in sinks. Cut stems\\
leak milk onto porcelain\\
and husbands, mothers children even\\
lovers wash their hands\\
in poems, read at conferences\\
to other women who wrap their hearts\\
in IBM ribbons and paint\\
their faces with liquid paper.


\clearpage
\section{Words from the Showerhead\\ on My Birthday}

Clearing of the throat:\\
Nagged and pounded,\\
straight down wet.\\
You glistened.\\
You forgot:\\
the meat thawed\\
and red-run\\
spoiled. Hot\\
water leashed,\\
treated on schedule,\\
a city decree,\\
a birthday trust in purity.\\
You remembered:\\
white, the coconut cake,\\
blue-tipped orange,\\
the candles' flames\\
extinguished that familiar song.\\
You raked:\\
red maple leaves\\
cupping rain. Little\\
birthday boats afloat,\\
sunk under by\\
tire treads rolled over\\
November.\\
You lit:\\
the arson spark,\\
the ark of pairs,\\
the hope in dreams\\
beyond despair.\\
Liquid birth\\
day. Celebrate\\
the miracle of waterfalls,\\
the day's beginning.


\clearpage
\section{Writing a Poem}

Think of it this way:\\
a tangled ball\\
of words\\
wrapped loosely.

Fling out the yo-yo string.\\
Make it sleep, walk,\\
go around-the-world\\
and back to the cradle.\\
Humming, purring, pulling\\
the metallic plastic moon\\
toward white harvest pages.

Or tug red-ribbon mornings\\
from the center of the sphere.\\
Iron them into party bows\\
perched on packages\\
full of silk images\\
and velvet phrases\\
slithering across paper,\\
shedding dead skin.

Yank the dangling sisal.\\
Shake those jumbled words\\
until frayed fiber ends\\
cover erasures\\
and empty space.\\
Twist stanzas\\
into one taut rope.

Walk that poem\\
around the block\\
and back\\
again.


\clearpage
\section{Yellow Pencil \#2\\
  with Pink Eraser}

Do you\\
still believe a pencil can kill\\
you or someone you love\\
or some strange man standing\\
in aisle six at the giant Kroger\\
store who licks his pencil lead\\
(you know how you do this too\\
out of habit)? He prints \emph{bread}\\
in tiny, damp letters\\
making sure he doesn't forget\\
anything, especially not\\
the pencil he tucks back\\
into his pocket where it lurks,\\
a concealed weapon\\
stashed next to his gun.

Back in first grade, your mother told you\\
her teacher had warned her whole class\\
(and more than once so it must be\\
true your mother said) not to\\
suck on the yellow pencil lead\\
you just sharpened or your teeth\\
would turn black from the poison\\
and the yellow wood shavings\\
would slip down your throat\\
ride your blood to your heart\\
where your own wet words pile up\\
sharp and shiny as meat cleavers.

And you all believed what you heard\\
back then about the dirty place\\
where bad kids stayed behind bars\\
(and where you would be sent if\\
you didn't behave and you did,\\
sort of) and these boys only got pencils\\
from the guards who wouldn't shave\\
the wood or the lead tip\\
so the boys couldn't write their stories\\
couldn't even stab their dark points\\
or erase themselves without a smudge.

Everything is something else. Be certain\\
where death lurks (check behind your eyes).\\
Yesterday yellow pencil paint lost its lead.


\clearpage
\section{Yin Yang --- Tibet}

I make my pilgrimage to Lhasa,\\
cradled like a lotus flower\\
in the Himalayan mountains.

Yak butter candles burn\\
behind doors of monasteries,\\
the smell of resurrection.

One red-orange door fuels my passion\\
for the yin-yang balance\\
of Tibet's life after death.

That closed door calls forth brown earth,\\
blue water, yellow fire, strong winds,\\
thin air and the consciousness of creation.

My hand yearns to tug the leather strap,\\
enter into centuries of spirits who linger\\
behind weathered wood, waiting for me.

Still, I hear shadow stories on wind,\\
the way life was back then when Dalai Lama\\
and all Buddhists prayed for enlightenment.

The knot of eternity is never-ending,\\
even for me who journeyed far enough\\
and high enough through Tibetan doors.

I circle this monastery three times,\\
prostrate myself on the threshold,\\
plant a prayer flag for freedom.

\clearpage
\pagestyle{empty}
\cleardoublepage
\mbox{ }

\clearpage

\mbox{ }

\end{document}
